---
title: "readme.rmd"
author: "_jas"
date: "`r format(Sys.Date())`"
output: 
    html_document:  
        keep_md: true  
always_allow_html: yes
---


# Tidy tuesday challenge 2018-09-18: Hypoxia & us-airports


## keep it simple:

### Objectives: 

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate

### Define your goal to do:

added at last, my goal here was to keep the week streak. just that.
after starting i came out with some other stuff, and cool ideas, that have to wait.

## 1st leson: Import the new data from github master 

update my fork with master repo: https://help.github.com/articles/merging-an-upstream-repository-into-your-fork/


* Open Git Bash.

* Change the current working directory to your local project.

* Check out the branch you wish to merge to. Usually, you will merge into master.

* git checkout master

* Pull the desired branch from the upstream repository. This method will retain the commit history without modification.

* git pull https://github.com/ORIGINAL_OWNER/ORIGINAL_REPOSITORY.git BRANCH_NAME

* If there are conflicts, resolve them. For more information, see "Addressing merge conflicts".

* Commit the merge.

* Review the changes and ensure they are satisfactory.

* Push the merge to your GitHub repository.

* git push origin master

## 2nd lesson: github html publication

publishing last week seem odd as images weren't keep
now i got the same issue, also with html cool stuff as data tables. (DT)

* http://stat545.com/block007_first-use-rmarkdown.html#step-6-publish-your-report

* http://stat545.com/bit006_github-browsability-wins.html#html

* https://yihui.name/knitr/options/


as got other stuff to do got the links, one step closer :p


```{r echo=FALSE,message=FALSE,warning=FALSE}

library(tibble)
library(dplyr)
library(tidyr)
library(stringr)
library(stringi)#; install.packages("stringi")
library(DT)
```

# Import

```{r}

hypoxia_file_location <- '../../data/2018-09-18/hypoxia.csv'
us_airports_file_location <- '../../data/2018-09-18/us-airports.csv'

hypoxia_raw_data <- read.csv(hypoxia_file_location,stringsAsFactors = FALSE)
us_airports_raw_data <- read.csv(us_airports_file_location,stringsAsFactors = FALSE)

glimpse(hypoxia_raw_data) 
glimpse(us_airports_raw_data) 

```

# Tidy

as seen both need some work on the naming.
hypoxia needs a little more work.

## Tidy Airports

for now just will ad an "ID" columns instead of X

```{r}
us_airports_data <- us_airports_raw_data %>% rename(id = X)
DT::datatable(us_airports_data,rownames = FALSE,options = list(dom = 't',initComplete = JS(
        "function(settings, json) {",
        "$(this.api().table().header()).css({'font-size': '50%'});",
        "}")))  %>% 
    formatStyle(names(us_airports_data), fontSize = "8pt")
```


## Tidy Hypoxia

as seen in file, it has 2 rows that are for colnames.
more on this on: https://github.com/rfordatascience/tidytuesday/files/2343596/Hypoxia.Article.proof.pdf

so ill reload the hipoxia data and get the 1st and 2nd rows as headers.

ill preserve the headers just in case are needed.

```{r}
#
hypoxia_headers_raw <- read.csv(hypoxia_file_location,stringsAsFactors = FALSE,nrows = 2,header = FALSE)
hypoxia_headers <- t(hypoxia_headers_raw) %>% as_tibble() %>% 
    rename(main = V1,measure =V2) %>%
    mutate(original_main=main,original_measure=measure)

# hypoxia_headers <- hypoxia_headers[-1,]
hypoxia_headers[1,1] <- "id"
hypoxia_headers[1,2] <- "id"

hypoxia_headers <- hypoxia_headers %>% 
    mutate(main = str_to_lower(main)) %>% # lowercase
    mutate(main = str_replace(main,"with o2=100%","")) %>%
    mutate(main = str_trim(main)) %>% # remove leading/trailing spaces
    mutate(measure = str_to_lower(measure)) %>% # lowercase
    mutate(measure = str_trim(measure)) %>% # remove leading/trailing spaces
    mutate(measure = if_else(measure == "","100%_o2",measure)) %>%
    mutate(main = str_replace_all(main," ","_")) %>% 
    mutate(main = str_replace_all(main,"%","perc")) %>%
    mutate(main = str_replace_all(main,"_1","")) %>%
    mutate(measure = str_replace_all(measure," ","_")) %>% 
    mutate(measure = str_replace_all(measure,"%","")) %>% 
    mutate(measure = str_replace_all(measure,">90_","")) %>% 
    mutate(measure = str_replace_all(measure,">35_","")) %>% 
    mutate(measure = str_replace_all(measure,"=","")) %>% 
    mutate(measure = str_replace_all(measure,"mmhg_","")) %>% 
    mutate(measure = str_replace_all(measure,"ft/m","ft_m")) %>% 
    mutate(header_end=paste0(main,"_",measure)) %>%
    mutate(header_end=if_else(header_end=="id_id","id",header_end))
    

#set colnames manually:
#colnames_hypoxia <- c("id","altitude","air_press","pp_O2","alv_p_O2","perc_sat_O2","alv_p_CO2","alv_p_O2_with_O2_100_perc","perc_sat_ _O2_1","alv_p_CO2_1")
# skips 2 rows

colnames_hypoxia <- hypoxia_headers %>% pull(header_end)
hypoxia_raw_data <- read.csv(hypoxia_file_location,
                             stringsAsFactors = FALSE,
                             skip = 2,
                             header = FALSE,
                             col.names = colnames_hypoxia)

hypoxia_data <- hypoxia_raw_data %>% 
    mutate(alv_po2_on_air=if_else( is.nan(alv_po2_on_air), as.numeric(NA),alv_po2_on_air )) %>% 
    mutate(perc_sat_o2_desired=if_else( is.nan(perc_sat_o2_desired),  as.numeric(NA),perc_sat_o2_desired )) %>% 
    mutate(alv_pco2_best=if_else( is.nan(alv_pco2_best),  as.numeric(NA),alv_pco2_best )) %>%
    mutate(altitude_ft_m=str_replace_all(altitude_ft_m,"k","")) %>%
    mutate(altitude_ft_m=if_else( str_detect(altitude_ft_m,"Sea level"),  "0/0", altitude_ft_m)) %>%
    separate(col = altitude_ft_m,into = c("altitude_ft","altitude_mts"),sep='/') %>%
    mutate(altitude_ft = as.numeric(altitude_ft)*1000,altitude_mts =as.numeric(altitude_mts)*1000 ) %>%
    select(-id) %>% # because id its wrong numbered.
    rownames_to_column("id")

DT::datatable(hypoxia_data,rownames = FALSE,options = list(dom = 't',initComplete = JS(
        "function(settings, json) {",
        "$(this.api().table().header()).css({'font-size': '50%'});",
        "}"))) %>% 
    formatStyle(names(hypoxia_data), fontSize = "8pt")

```

# Understand

## about airports

as spent much time on hypoxia , i fall short on airports.

## about hypoxia

seems to have several domain things that are on the article, also some extra data that is not here for other outputs.

other things that figure on the table and not so clear whate they are are colored fonts, on red, green and lightblue

# Comunicate

## conclussion 

* it has been a nice practice to clean small data.
* dealt with some workflow issues ( git )
* it still have juice to pull out :D

## future

* translate the work to spanish.
* can make a better univariated analysis
* add similar analysis to argentina dataset.

### future airports

* can me made a map similar to previous week ( hextile )
* can try other map types: as expressed on: http://ryanhafen.com/blog/geofacet
* can make simple stuff as rank histogram  , like the cat/dog.

### future hypoxia

* can be transform to long format and then used in ggplot to have some plots.
* ex: something showing the heigths and values related there.




