---
title: "week 2018-09-25"
author: "_jas"
date: "`r format(Sys.Date())`"
output: html_document
---

# Week 26 - Invasive Species in Africa

## keep it simple:

### Objectives: 

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate

# workflow

* Open Git Bash.

* Change the current working directory to your local project.

* Check out the branch you wish to merge to. Usually, you will merge into master.

* git checkout master

* Pull the desired branch from the upstream repository. This method will retain the commit history without modification.

* git pull https://github.com/rfordatascience/tidytuesday.git master

* If there are conflicts, resolve them. For more information, see "Addressing merge conflicts".

* Commit the merge.

* Start the week challenge !

# Import

To have an idea of how the data is presented gotta check: 

http://www.pnas.org/content/pnas/suppl/2016/06/16/1602205113.DCSupplemental/pnas.1602205113.sapp.pdf

```{r}
library(dplyr)
library(tidyr)
library(here)#;install.packages("here") # 1st time using this , its quite handy :D
library(janitor)#;install.packages("janitor") janitor::clean_names() # 1st time using janitor, really usefull ! 

```


```{r}

table_1_raw <- read.csv(here("data", "2018-09-25","raw","table1.csv"),stringsAsFactors = FALSE)
table_2_raw <- read.csv(here("data", "2018-09-25","raw","table2.csv"),stringsAsFactors = FALSE)
table_3_raw <- read.csv(here("data", "2018-09-25","raw","table3.csv"),stringsAsFactors = FALSE)
table_4_raw <- read.csv(here("data", "2018-09-25","raw","table4.csv"),stringsAsFactors = FALSE)
table_6_raw <- read.csv(here("data", "2018-09-25","raw","table6.csv"),stringsAsFactors = FALSE)

```


# Tidy

```{r}
table_1 <- table_1_raw %>% janitor::clean_names() %>% as_tibble()
table_2 <- table_2_raw %>% janitor::clean_names() %>% as_tibble()
table_3 <- table_3_raw %>% janitor::clean_names() %>% as_tibble()
table_4 <- table_4_raw %>% janitor::clean_names() %>% as_tibble()
table_6 <- table_6_raw %>% janitor::clean_names() %>% as_tibble()

```



# Understand

will do an iteration for each file, and later on more if needed / time helps.

## table 1: iteration 1

### transform

as seen on the original PDF , is a ranking with rank / country / ott
but its broken into 3 columns to use the pdf space.

#### 3 columns

1st step its to get the dataset in the correct form , that is 3 columns

```{r}

table_1_part_1 <-  table_1 %>% select(rank,country,o_tt)
table_1_part_2 <-  table_1 %>% select(rank_1,country_1,o_tt_1) %>% rename(rank=rank_1,
                                                                          country=country_1,
                                                                          o_tt=o_tt_1)

table_1_part_3 <-  table_1 %>% select(rank_2,country_2,o_tt_2) %>% rename(rank=rank_2,
                                                                          country=country_2,
                                                                          o_tt=o_tt_2)

table_1_final <- table_1_part_1 %>% union(table_1_part_2) %>% union(table_1_part_3) %>% arrange(rank)


```

#### see data

```{r}
library(skimr)#install.packages('skimr')
skim(table_1_final)
```

as seen on summary; 

* country: have 125 unique values, that are complete
* rank: are 124 values with a missing value
* o_tt: are 124 values with a missing value

checking the missing value

```{r}
table_1_final %>% filter(!complete.cases(rank))
```

as can see is a full blank row, we can delete that safely.

```{r}
table_1_final <- table_1_final %>% filter(complete.cases(rank))
skim(table_1_final)
```

there we go , 124 full .


### visualize

a simple visualization

```{r}
library(ggplot2)

ggplot(table_1_final,aes(y=o_tt)) + 
    geom_boxplot() + 
    geom_jitter(aes(y=o_tt,x=0),
               position=position_jitter(width=0.1,height=0),
               alpha=0.6,
               size=3) 

```

a simple nice plot , could have the dots representing each country continent. for future expansion.


### model

--

## iteration 1: table 2

have the same issue as table 1, all data is spread into 3 columns

```{r}

table_2 %>% head()

```


after looking at the data; 
we see that 
- the 1st row its part of the header !
- the 1st and 2nd column got a problem. 
- 1st column got both columns thogheter 
- 2nd column its filled with NA

todo: 

- fix header
- remove NA columns
- split rank country  in 2 columns
- get the data into 3 columns.
- remove the money sign + parse as number

### transform

```{r}
library(readr)#;install.packages("readr") # 1st time using parse number :D
library(stringr)


table_2_headers <- c('rank','country','TICt_M_USD')
table_2_new <- table_2[ -1,] # remove the 1st row
table_2_part_1 <- table_2_new %>% select(x,ti_ct_millions) 
table_2_part_2 <- table_2_new %>% select(x_2,ti_ct_millions_1) %>% rename(x=x_2,
                                                            ti_ct_millions=ti_ct_millions_1)
table_2_part_3 <- table_2_new %>% select(x_4,ti_ct_millions_2) %>% rename(x=x_4,
                                                            ti_ct_millions=ti_ct_millions_2)

table_2_final <- table_2_part_1 %>% union(table_2_part_2) %>% union(table_2_part_3)

table_2_final <- table_2_final %>% 
    mutate(rank = readr:::parse_number(x),
           country=stringr:::str_extract(x, "[:alpha:]+"),
           TICt_USD =  readr:::parse_number(ti_ct_millions) * 1000000 ) %>% 
    select(rank,country,TICt_USD)


# missing 1 all values in row
table_2_final %>% filter(!complete.cases(rank))

# just keep the complete cases
table_2_final <- table_2_final %>% filter(complete.cases(rank))
skim(table_2_final)

```

### visualize

```{r}
ggplot(table_2_final,aes(y=TICt_USD)) + 
    geom_boxplot() + 
    geom_jitter(aes(y=TICt_USD,x=0),
               position=position_jitter(width=0.1,height=0),
               alpha=0.6,
               size=3) 
```


well the boxplot doesnt help 

having a map will be cooler: 

```{r}

# https://stackoverflow.com/questions/47502716/creating-a-choropleth-in-r-measuring-violent-crimes-by-county-in-north-carolina
# library(ggthemes)#;install.packages("ggthemes")
# 
# world_map <- tbl_df(map_data("world"))
# 
# df_world_countries <- data_frame(country = unique(world_map$region)) %>% as_tibble()
# 
# df_world_countries <- df_world_countries %>% left_join(table_2_final)
# 
# df_world_countries <- df_world_countries %>% mutate(TICt_USD=if_else(is.na(TICt_USD),0,TICt_USD))
# 
# data_world_join <- world_map %>% left_join(df_world_countries,by=c("region"="country"))
# 
# ggplot() +
#   geom_polygon(data=data_world_join, color="black",
#                aes(x=long, y=lat, group=region, fill=TICt_USD)) +
#   viridis::scale_fill_viridis(name="TICt_USD ", direction=-1) +
#   coord_map("polyconic") +
#   ggthemes::theme_map() +
#   theme(legend.position="bottom")
# 
# 
# ggplot() +
#   geom_polygon(data=world_map, color="black",
#                aes(x=long, y=lat, group=region, fill=1)) +
#   viridis::scale_fill_viridis(name="1 ", direction=-1) +
#   coord_map("polyconic") +
#   ggthemes::theme_map() +
#   theme(legend.position="bottom")


```

well nothing came good from this map trial , was trying to get a more short way to get the map
of the world data but cant get to it.


### model


## iteration 1: table 3

### transform

### visualize

### model

--

## iteration 1: table 4

### transform

### visualize

### model

--

## iteration 1: table 6

### transform

### visualize

### model

--


# Communicate


# future work: 

## future work: branch pulses ? 

not related to this week challenge, but to answer " how many branches of the main repo are still active ?"
may be doing a shiny app ?


# Conclusion. 

1.30 hs. 2 / 5 datasets

ton of banging head looking for easy maps.


learnt: 

here: for easy locations across a project ; ( self ) 
janitor: for quick colnames;  ( invasive species r)
readr: for easy number parsing; ( invasive species r)
stringr: extract with regex the alpha numeric with country ( invasive species r)


thats all for this shot. 
might come back later.
