
---
title: "Tidy tuesday challenge: Week 2019-06-11 meteorites !"
author: "julio"
date: "2019-06-11"
output: html_document
---

# Tidy tuesday challenge: Week 2019-06-11 meteorites !

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

### Data:

meteorites


### objectives:

test cross it with population of places.
so you can have places that have not been explored for meteorites inpacts

- population can be extrated from gapminder

## details:


## import data
```{r echo=FALSE,message=FALSE,warning=FALSE}
library(tidyverse)

library(stringr)#;
library(rebus)#; install.packages('rebus')
library(tidytext)

```


```{r echo=FALSE,message=FALSE,warning=FALSE}
meteorites <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-11/meteorites.csv")
```

# explore data

```{r echo=FALSE,message=FALSE,warning=FALSE}
glimpse(meteorites)

```


```{r echo=FALSE,message=FALSE,warning=FALSE}
meteorites %>% skimr::skim()

# nobel_winner_all_pubs$
```

can see that several missing vaues exist, as want to get this quickly done, ill just ignore them.

```{r echo=FALSE,message=FALSE,warning=FALSE}
meteorites_complete <- meteorites %>% filter(complete.cases(meteorites))
glimpse(meteorites_complete)
skimr::skim(meteorites_complete)
```

ok now go tto check countries in lat longs.

```{r echo=FALSE,message=FALSE,warning=FALSE}

library(rgdal)
library(geojsonsf)#;install.packages("geojsonsf")
library(sf)
library(sp)
library(spatialEco)


# world_ggplot <- ggplot2::map_data("world")
# https://gis.stackexchange.com/questions/230608/creating-an-sf-object-from-the-maps-package
world_sf <- sf::st_as_sf(maps::map("world", plot = FALSE, fill = TRUE))

meteor_points <- st_as_sf(meteorites_complete, coords = c("long", "lat"))
st_crs(meteor_points) <- st_crs(world_sf)

#https://stackoverflow.com/questions/3647744/intersecting-points-and-polygons-in-r
# library(spatialEco)#; install.packages("spatialEco")
meteors_by_country <- point.in.poly(meteor_points, world_sf)


meteors_country_sf <- st_as_sf(meteors_by_country) %>% 
    rename(country=ID,
           meteor_name=name)

meteors_country_summary <- meteors_country_sf %>% 
    group_by(country) %>% 
    summarise(total_meteors=n()) %>% 
    mutate(binify=case_when(total_meteors >= 15000 ~ "+15k",
                            total_meteors >= 5000 & total_meteors < 15000 ~ "5k - 15k",
                            total_meteors >= 1000 & total_meteors < 5000 ~ "1k - 5k",
                            total_meteors >= 500 & total_meteors < 1000 ~ "500 - 1k",
                            total_meteors >= 100 & total_meteors < 500 ~ "100 - 500",
                            total_meteors < 100 ~ "< 100")) %>% 
    st_set_geometry(NULL)

meteor_country_map <- world_sf %>% 
    left_join(meteors_country_summary, 
               by = c("ID"="country"))

```



```{r}

meteor_variability_bin <- meteors_country_summary

        # filter(total_meteors<1000) %>% 
meteor_variability_bin %>% 
    ggplot(aes(y=total_meteors,x=binify,label=country))+
    geom_boxplot()
    # ggrepel::geom_label_repel()
    # geom_polygon(aes(long, lat, group = group),
    #              fill = "#FFFFFF", color = "#505050", data = mundo,
    #              alpha=0.5) 
meteor_variability
```

```{r}

# ggplto colors
# https://www.datanovia.com/en/blog/ggplot-theme-background-color-and-grids/
# chorolplet na values
# https://stackoverflow.com/questions/16220812/how-do-i-change-the-na-color-from-gray-to-white-in-a-ggplot-choropleth-map

# meteor_country_map$binify <- factor(meteor_country_map$binify,levels = c("+15k","1k - 5k","500 - 1k","100 - 500","< 100"))
meteor_country_map$binify <- fct_rev(factor(meteor_country_map$binify,levels = c("+15k","1k - 5k","500 - 1k","100 - 500","< 100")))




n_meteors_country <- meteor_country_map %>% 
    # filter(total_meteors<5000) %>% 
    ggplot()+
    geom_sf(aes(fill=binify),color=NA)+#+
    # scale_fill_brewer(palette = "YlGnBu",na.value="grey")+
    scale_fill_brewer(palette = "PuBu",na.value="grey")+
    # scale_fill_viridis_d(na.value="grey")+
    theme( panel.border = element_blank(),
      panel.background = element_blank())+
    labs(title="Amount of meteors by country",
         fill="# meteors",
         caption="#tidytuesday")

n_meteors_country

ggsave(n_meteors_country,filename = "n_meteors_country.png",width = 10,height = 5)

```

# population

as is accumulated values, just get the last as last meteor year near to gapminder data

```{r}
# 2007
# gapminder::gapminder %>% select(country,pop,year,) %>%  pull(year) %>% max()

# 2013
# meteorites_complete %>% arrange(desc(year))  pull(year)  %>% 
library(wbstats)
# install.packages("wbstats")
new_cache <- wbcache()
str(wb_cachelist, max.level = 1)
unemploy_vars <- wbsearch(pattern = "Total population")

```





# bivariate population / meteors


things got better as we get more reviews per item, nevertheless there are a lot of reviews spread.

#tweet: 

2019-06-11 #TidyTuesday #rstats Meteors! Made meteors by country/region. Can see that antartica has more impacts, 
would be interesting crossing with region size, and population. As more people might find easily more meteors.
may be some bivarate color palette map ... next time :p 

https://twitter.com/jspairani/status/1137894754220466177

# communicate

well i was aiming to cross with gapminder or other population / size data, too much time spent. 

places to ge tthe data: 
- gapminder: ended 2007
- other: 
https://cran.r-project.org/web/packages/wbstats/vignettes/Using_the_wbstats_package.html
https://population.un.org/wpp/Download/Standard/Population/



