---
title: "Tidy tuesday challenge: Week 2019-04-09 Grand Slam"
author: "julio"
date: "02-04-2019"
output: html_document
---

# Tidy tuesday challenge: Week 2019-04-03 Grand Slam

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

### Data:

this week data its related to Grand Slam
https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-04-09

### objectives:

- download and d osomething ...

## details:


## import data
```{r echo=FALSE,message=FALSE,warning=FALSE}
library(dplyr)
library(tidyr)
library(skimr)
library(readr) 
library(ggplot2)
library(scales)
library(forcats)
library(lubridate)
library(gganimate)
library(animation)

```


```{r echo=FALSE,message=FALSE,warning=FALSE}
player_dob <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-04-09/player_dob.csv")

grand_slams <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-04-09/grand_slams.csv")

grand_slam_timeline <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-04-09/grand_slam_timeline.csv")
```

# explore data

```{r echo=FALSE,message=FALSE,warning=FALSE}

player_dob %>% skimr::skim()
player_dob %>% glimpse()

grand_slams %>% skimr::skim()
grand_slams %>% glimpse()



grand_slams_procs <- grand_slams %>% 
    mutate(court=case_when(grand_slam=='australian_open'~'hard',
                           grand_slam=='french_open'~'clay',
                           grand_slam=='us_open'~'hard',
                           grand_slam=='wimbledon'~'grass')) %>% 
    # https://github.com/emilykuehler/tidytuesday/blob/master/grand-slam-tennis/tennis.R
    mutate(name = ifelse(name == 'Evonne Goolagong Cawley', 'Evonne Goolagong', name)) %>% 
    mutate(tournament_order = case_when(grand_slam=='australian_open' ~ 1,
                                          grand_slam=='french_open' ~ 2,
                                          grand_slam=='wimbledon' ~ 3,
    grand_slam=='us_open' ~ 4)) %>% 
arrange(tournament_date)

# grand_slams %>% count(grand_slam)
# 
# grand_slam_timeline %>% skimr::skim()
# grand_slam_timeline %>% glimpse()
```


# Players by court type; are there players specialized , and players multi surface ?



```{r echo=FALSE,message=FALSE,warning=FALSE}
player_gs_plot <- grand_slams_procs %>% 
    filter(year < 2019) %>% 
    count(name,court) %>% 
    mutate(name=fct_reorder(name,n)) %>%
    arrange(desc(n)) %>% 
    # head(20) %>%
    ggplot(aes(x=name,y=n,fill=court))+
    geom_col()+
    coord_flip()+
    theme_light()+
    theme(axis.text.y = element_text(size=5) )+
    labs(title="Who won in which court type ? ",
         subtitle="50 of Tennis Grandslams: winners: 1968 - 2018",
         caption="#TidyTuesday week 2019-04-09",
         x="",y="# of grand slams",fill="court type")

ggsave(player_gs_plot,filename = "player_gl_all_time_catchy.png")
   
```

```{r echo=FALSE,message=FALSE,warning=FALSE}
# make a cumulative plot
grand_slams_procs %>% count(year)
grand_slams_procs %>% 
    count(name,court,gender,year) %>% 
    mutate(name=fct_reorder(name,n)) %>% 
    ggplot(aes(x=name,y=n,fill=court))+
    geom_col()+
    coord_flip() +
    transition_time(year) +
    ease_aes('linear')+
    labs(title="Grandslam players by year & court",
        subtitle = 'Year: {frame_time}', 
         x = '', 
         y = '# grand slams') +#+
    # facet_grid(~court)
    
grand_slams_procs %>% 
    count(court,year) %>% 
    ggplot(aes(x=year,y=n, color = court))+
    geom_line()
    
```


```{r echo=FALSE,message=FALSE,warning=FALSE}
grand_slams_procs %>% 
    count(court,year) %>% 
    ggplot(aes(x=year,y=n, color = court))+
    geom_line(alpha=0.5)+
    facet_grid(court~.) +
    # scale_color_discrete() +
    labs("Amount of winners by Court types by year",
         x="",y="#amount of ")+
    theme_light()
```

# Player Race

```{r echo=FALSE,message=FALSE,warning=FALSE}
# https://github.com/emilykuehler/tidytuesday/blob/master/grand-slam-tennis/tennis.R
init_df <- grand_slams_procs %>% 
  filter(year <= 1975) %>%
  group_by(name) %>% 
  filter(rolling_win_count==max(rolling_win_count)) %>% 
  ungroup() %>% 
  top_n(10, wt=rolling_win_count) %>% 
  arrange(desc(rolling_win_count)) %>% 
  select(name,gender, rolling_win_count) %>% 
  mutate(curr_year = 1975,
ordering = as.double(rev(seq(10:1))) * 1.0)


grand_slams_procs %>% count(year)

for (i in 1976:2019) {
  for (j in 1:4) {
    tmp_df <- grand_slams_procs %>% 
      filter(year < i | (year==i & tournament_order <= j)) %>% 
      group_by(name) %>% 
      filter(rolling_win_count==max(rolling_win_count)) %>% 
      ungroup() %>% 
      top_n(10, wt=rolling_win_count) %>%
      select(name, gender, rolling_win_count) %>%
      arrange(desc(rolling_win_count)) %>%
      slice(1:10) %>% 
      mutate(curr_year = i,
             tournament_num = j,
             ordering = as.double(rev(seq(10:1))) * 1.0) 
    
    init_df <- init_df %>%
      bind_rows(tmp_df)
  }
}

final_df <- init_df %>% 
  group_by(curr_year, tournament_num) %>% 
  mutate(num = group_indices()) %>% 
  ungroup()

# plot race
#make plot
barplot_race <- ggplot(aes(ordering, group = name), data = final_df) +
  geom_tile(aes(y = rolling_win_count / 2, 
                height = rolling_win_count,
                width = 0.9, fill=gender), alpha = 0.9) +
  scale_fill_manual(values = c('#F8AFA8','#74A089')) +
  geom_text(aes(y = rolling_win_count, label = name), 
            # family=my_font, 
            nudge_y = -2, size = 3) +
  geom_text(aes(y = rolling_win_count, label = as.character(rolling_win_count)), 
            # family=my_font, 
            nudge_y = 0.5) +
  geom_text(aes(x=1,y=18.75, label=paste0(curr_year)), 
            # family=my_font, 
            size=8, color = 'gray45') +
  coord_cartesian(clip = "off", expand = FALSE) +
  coord_flip() +
  labs(title = 'Most Grand Slam Singles Championships',
       subtitle = 'Open Era Only',
       caption = 'data source: Wikipedia | plot by @emilykuehler',
       x = '',
       y = '') +
  transition_states(num, 
                    transition_length = 4, state_length = 3) +
  ease_aes('cubic-in-out')

animate(barplot_race, nframes = 1000, fps = 30, width = 600, height = 400, res=80, detail = 3)

anim_save("barplot_race.gif")

```
