---
title: "Tidy tuesday challenge: Week 2019-05-05 Chicago Bird Collisions"
author: "julio"
date: "5 de mayo de 2019"
output: html_document
---

# Tidy tuesday challenge: Week 2019-05-05 Chicago Bird Collisions

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

### Data:

anime data, as seen on readme.

### objectives:

- do some text mining

## details:


## import data
```{r echo=FALSE,message=FALSE,warning=FALSE}
library(dplyr)
library(tidyr)
library(skimr)
library(readr) 
library(ggplot2)
library(scales)
library(forcats)
library(lubridate)
library(gganimate)
library(animation)

library(tm)#; install.packages('tm')
library(topicmodels)#; install.packages('topicmodels')
library(stringr)#;
library(rebus)#; install.packages('rebus')
library(tidytext)

library(widyr) # for pairwise core
library(ggraph) # for graph drawing
library(igraph)  # for graph managing

```


```{r echo=FALSE,message=FALSE,warning=FALSE}
bird_collisions <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-04-30/bird_collisions.csv")
mp_light <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-04-30/mp_light.csv")
```

# explore data

```{r echo=FALSE,message=FALSE,warning=FALSE}
glimpse(bird_collisions)
glimpse(mp_light)

# bird_collisions %>% count(family,genus,species,sort = TRUE)

```

```{r echo=FALSE,message=FALSE,warning=FALSE}
skimr::skim(bird_collisions)
skimr::skim(mp_light)

```


```{r echo=FALSE,message=FALSE,warning=FALSE}
# # Clustering by numeric vars, not id and not missing
# library(dummies)
# 
# dummified_bird_collisions <- bird_collisions %>% 
#     mutate(stratum=as.factor(stratum),
#            habitat=as.factor(habitat),
#            flight_call=as.factor(flight_call)) %>% 
#     dummies::dummy.data.frame()
# 
# # Calculate the Distance
# dist_survey <- dist(dummified_bird_collisions,method='binary')
# 
# # Print the Original Data
# print(job_survey)
# 
# # Print the Distance Matrix
# print(dist_survey)

# glimpse(dummified_bird_collisions)
# bird_collisions %>% 
#     # mutate(year=year(date),month=month(date)) %>% 
#     # count(year,date,stratum,habitat,flight_call,sort = TRUE) %>% 
#     count(habitat,flight_call,sort = TRUE) %>% 
#     ggplot(aes(x=habitat,y = flight_call,fill=n))+
#     geom_tile()+
#     labs(title="Chicago Bird Collisions")
# 
# 
# bc_tmp <- bird_collisions %>% 
#     mutate(stratum=as.factor(stratum),
#            habitat=as.factor(habitat),
#            flight_call=as.factor(flight_call)) %>% 
#     select(stratum,habitat,flight_call) %>%
#     mutate(stratum=as.integer(stratum),
#            habitat=as.integer(habitat),
#            flight_call=as.integer(flight_call)) #%>%
#     
#     
# 
# bc_tmp[] <-  lapply(bc_tmp,as.integer)
# head(bc_tmp)
# corrplot::corrplot(cor(bc_tmp[1:3]))
# not happy with result, throw away idea
```

```{r}
# ton of random stuff
# 
# bc_tmp <- bird_collisions %>% 
#     mutate(year=year(date)) %>% 
#     # genus,family,species,
#     count(year,flight_call) %>% 
#     mutate(flight_call=fct_reorder(flight_call,n)) %>% 
#     mutate(percent=n/sum(n))
# 
# 
# 
#      autoplot()
#     labs()
# 
# bc_tmp %>% 
#     ggplot(aes(x=year,y=percent,color=flight_call,fill=flight_call)) +
#     geom_line()
#         # geom_density()
#     # geom_line()+
#     # geom_point()+
#     # geom_density(aes(y=percent),adjust=1.5)
#     # geom_area(alpha=0.5, aes(fill=flight_call,group = flight_call), position='stack')
# library(forecast)
# tsdisplay(diff(WWWusage))
# ggtsdisplay(USAccDeaths, plot.type="scatter")
# 
# install.packages("ggseas")
# library(ggseas)
# 
# filtered_bc_tmp <- bc_tmp %>% filter(flight_call=="Yes") %>% 
#     select(year,n) 
# 
# filtered_bc_tmp %>% 
#     ggplot(aes(x = year, y = n))+
#     geom_line(colour = "grey75") +
#     stat_rollapplyr(width = 12, align = "center") 
# 
# 
# range(bird_collisions$date)

# finally an animated plot to close the week :P
bc_bar_plot <- bird_collisions %>% 
    mutate(year=year(date)) %>% 
    count(year,family) %>% 
    mutate(family=fct_reorder(family,n)) 



plot_anim <- bc_bar_plot %>% 
    ggplot(aes(x=family,y=n,fill=family))+
    geom_col()+
    geom_text(aes(label = as.character(year)),
            x = (length(levels(bc_bar_plot$family))/2),
            y = (max(bc_bar_plot$n)/2),
            color = '#cdcdcd',
            size = 20)+
    coord_flip()+
    theme_light()+
    theme(legend.position = "none")+
    labs(title="Chicago Bird collisions", subtitle="yearly by bird family",
         caption="#TidyTuesday",
         x="",y="")+
    transition_states(year,transition_length = 1,state_length = 0.1,wrap = FALSE) +
    ease_aes('linear')
 

# animando final
animate(plot_anim,
        fps = 6)

# guardando animacion como gif
anim_save('tidy_tuesday_bird_collisions_chicago.gif')
```

## communicate

well , this week still working on animations, 
got to see how to made better transitions with texts.


