

---
title: "Tidy tuesday challenge: Week 2019-08-20 Nuclear Explosions"
author: "julio"
date: "2019-08-20"
output: html_document
---

# Tidy tuesday challenge:  Week 2019-08-20 Nuclear Explosions

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

### Data:
 
 Week 2019-08-20 Nuclear Explosions
 
### objectives:

explore & do something

## details:


## import data
```{r echo=FALSE,message=FALSE,warning=FALSE}
library(magrittr) # para el %T>%
library(tidyverse)
# library(sf)
library(dplyr)
library(stringr)#;
# library(rebus)#; install.packages('rebus')
# library(tidytext)
# library(prophet)


# install.packages("Rcpp")
# remotes::install_github("tylermorganwall/rayshader")
# library(rayshader)
library(lubridate)
library(ggforce)
library(ggrepel)

# library(arules)#install.packages('arules')
# library(arulesViz)#install.packages('arulesViz')
# library(igraph)

```


```{r echo=FALSE,message=FALSE,warning=FALSE}
nuclear_explosions <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-08-20/nuclear_explosions.csv")
```

# explore data


```{r echo=FALSE,message=FALSE,warning=FALSE}
nuclear_explosions %>% head()

```

```{r echo=FALSE,message=FALSE,warning=FALSE}
nuclear_explosions %>% glimpse()

```

```{r echo=FALSE,message=FALSE,warning=FALSE}
nuclear_explosions %>% skimr::skim()


```


purpose & type migth get cryptic so , lets be more descriptive

```{r echo=FALSE,message=FALSE,warning=FALSE}
nuclear_explosions_proc <- nuclear_explosions %>% 
    separate(col=purpose,into = c("purpose","purpose_subtype"),sep = ":") %>% # subtype
    separate_rows(col=purpose,sep = "/") %>% # multitype
    mutate(purpose_desc=case_when(purpose=="COMBAT"~"WWII bombs dropped over Japan",
                               purpose=="FMS"~"Soviet test,\nstudy phenomenon of nuclear explosion",
                               purpose=="ME"~"Military Exercise",
                               purpose=="PNE"~"Peaceful nuclear explosion",
                               purpose=="SAM"~"Soviet test,\naccidental mode/emergency",
                               purpose=="SE"~"French/US tests - testing safety\nof nuclear weapons in case of accident",
                               purpose=="TRANSP"~"Transportation-storage purposes",
                               purpose=="WE"~"British, French, US, evaluate effects\n of nuclear detonation on various targets", 
                               purpose=="WR"~"Weapons development program")) %>% 
    mutate(type_desc=case_when(type=="ATMOSPH" ~ "Atmospheric",
type=="UG" ~ "underground",
type=="BALLOON" ~ "Balloon drop",
type=="AIRDROP" ~ "Airplane deployed",
type=="ROCKET" ~ "Rocket deployed",
type=="TOWER" ~ "deplyed at top of constructed tower",
type=="WATERSURFACE" ~ "on surface of body of water",
type=="BARGE" ~ "on barge boat",
type=="SURFACE" ~ "on surface or in shallow crater",
type=="UW" ~ "Underwater",
type=="SHAFT" ~ "Vertical Shaft underground",
type=="TUNNEL/GALLERY" ~ "Horizontal tunnel")) %>% 
    mutate(purpose_subtype_desc=case_when(purpose_subtype=="PLO"~"Plowshare programme",
                                          purpose_subtype=="V"~"Vela uniform tests"))
```

lets check descriptions and types / purposes

```{r echo=FALSE,message=FALSE,warning=FALSE}
nuclear_explosions_proc %>% 
    count(purpose,purpose_desc)
```

can see can have multiple purposes.

also can see that the  "PNE" type can have other types not specified, as : plo and v

so plo & v subtypes all send em to other colum ( : separator )
and the multitype ones ( " / " separator) will be treated as a bomb with 2 types.

well for more detail can chech the pdf. 
its already late so ill go for a simple one.

wich countries done more tests over time, and wich types / purposes

or wich purposes over time / type of bomb



```{r echo=FALSE,message=FALSE,warning=FALSE}

label_events <- data.frame(year=c(1963,1996),desc=c("Partial test Ban\n(no more ATMO)",
                                                 "Comprehensive\nnuclear test ban"),
                           nudge_y_value=c(1205,1500),
                           nudge_x_value=c(20,-20),
                           stringsAsFactors = FALSE)

data_count <- nuclear_explosions_proc %>% 
    count(year,purpose_desc, sort=TRUE)

line_simple_plot <-  data_count %>% 
    ggplot(aes(x=year, y=n,color=purpose_desc))+
    geom_line()+
    geom_vline(xintercept = label_events$year,linetype=2)+
    geom_label_repel(data = label_events, aes(x=year,
                                              y=100,
                                              label=paste0(desc,"\n",year)
                                              # nudge_y = nudge_y_value,
                                              # nudge_x=nudge_x_value),
                                              ),
                     color="#333333",
                     nudge_y=c(0,50),
                     nudge_x=c(15,-15),
                     size=3
                     )+
    scale_x_continuous(breaks = seq(from=min(nuclear_explosions_proc$year),
                                    to=max(nuclear_explosions_proc$year),
                                    by=5))+
    scale_y_continuous(breaks = seq(from=0,
                                    to=max(data_count$n)+15,
                                    by=15))+
    theme_light()+
    theme(axis.text.x = element_text(angle=90))+
    
    
    labs(title="Amount of nuclear explosions by purpose over the years",
         subtitle=paste0("From ",min(nuclear_explosions_proc$year),
                         " to ",max(nuclear_explosions_proc$year)),
         x="",y="",color="Purpose",
         caption = "#tidyTuesday")

ggsave(plot=line_simple_plot,filename = "line_simple_plot.png",width = 7,height = 5)

```

```{r echo=FALSE,message=FALSE,warning=FALSE}

label_events <- data.frame(year=c(1963,1996),desc=c("Partial test Ban\n(no more ATMO)",
                                                 "Comprehensive\nnuclear test ban"),
                           nudge_y_value=c(1205,1500),
                           nudge_x_value=c(20,-20),
                           stringsAsFactors = FALSE)

years_cum_sum <- data_count %>% group_by(year) %>% summarise(total=sum(n))

col_simple_plot <- data_count %>% 
    ggplot(aes(x=year, y=n,fill=paste0(purpose_desc,"\n")))+
    geom_col()+
    geom_vline(xintercept = label_events$year,linetype=2)+
    geom_label_repel(data = label_events, aes(x=year,
                                              y=125,
                                              label=paste0(desc,"\n",year)
                                              # nudge_y = nudge_y_value,
                                              # nudge_x=nudge_x_value),
                                              ),
                     color="#333333",
                     fill="#FFFFFF",
                     nudge_y=c(5,5),
                     nudge_x=c(-14,-15),
                     size=2.5
                     )+
    scale_x_continuous(breaks = seq(from=min(nuclear_explosions_proc$year),
                                    to=max(nuclear_explosions_proc$year),
                                    by=5))+
    scale_y_continuous(breaks = seq(from=0,
                                    to=max(years_cum_sum$total)+15,
                                    by=15))+
    theme_light()+
    theme(axis.text.x = element_text(angle=90))+
    
    
    labs(title="Acumulated amount of nuclear explosions by purpose over the years",
         subtitle=paste0("From ",min(nuclear_explosions_proc$year),
                         " to ",max(nuclear_explosions_proc$year)),
         x="",y="",fill="Purpose",
         caption = "#tidyTuesday")

ggsave(plot=col_simple_plot,filename = "col_simple_plot.png",width = 7,height = 5)


```


```{r echo=FALSE,message=FALSE,warning=FALSE}

label_events <- data.frame(year=c(1963,1996),desc=c("Partial test Ban\n(no more ATMO)",
                                                 "Comprehensive\nnuclear test ban"),
                           nudge_y_value=c(1205,1500),
                           nudge_x_value=c(20,-20),
                           stringsAsFactors = FALSE)

log_col_plot <- data_count%>% 
    ggplot(aes(x=year, y=log(n),fill=paste0(purpose_desc,"\n")))+
    geom_col()+
    geom_vline(xintercept = label_events$year,linetype=2)+
    geom_label_repel(data = label_events, aes(x=year,
                                              y=10,
                                              label=paste0(desc,"\n",year)
                                              # nudge_y = nudge_y_value,
                                              # nudge_x=nudge_x_value),
                                              ),
                     color="#333333",
                     fill="#FFFFFF",
                     nudge_y=c(5,5),
                     nudge_x=c(-14,-15),
                     size=2.5
                     )+
    scale_x_continuous(breaks = seq(from=min(nuclear_explosions_proc$year),
                                    to=max(nuclear_explosions_proc$year),
                                    by=5))+
    scale_y_continuous(breaks = seq(from=0,
                                    to=15,
                                    by=1))+
    theme_light()+
    theme(axis.text.x = element_text(angle=90))+
    
    labs(title="Amount of nuclear explosions by purpose over the years",
         subtitle=paste0("Log count explosions.",
                         "From ",min(nuclear_explosions_proc$year),
                         " to ",max(nuclear_explosions_proc$year)),
         x="",y="",fill="Purpose",
         caption = "#tidyTuesday")

ggsave(plot=log_col_plot,filename = "log_col_plot.png",width = 7,height = 5)

```




```{r echo=FALSE,message=FALSE,warning=FALSE}

label_events <- data.frame(year=c(1963,1996),desc=c("Partial test Ban\n(no more ATMO)",
                                                 "Comprehensive\nnuclear test ban"),
                           nudge_y_value=c(1205,1500),
                           nudge_x_value=c(20,-20),
                           stringsAsFactors = FALSE)

log_line_plot <- data_count %>% 
    ggplot(aes(x=year, y=log(n),color=paste0(purpose_desc,"\n")))+
    geom_line()+
    geom_vline(xintercept = label_events$year,linetype=2)+
    geom_label_repel(data = label_events, aes(x=year,
                                              y=log(100),
                                              label=paste0(desc,"\n",year)
                                              # nudge_y = nudge_y_value,
                                              # nudge_x=nudge_x_value),
                                              ),
                     color="#333333",
                     nudge_y=c(5,5),
                     nudge_x=c(-14,-15),
                     size=2.5
                     )+
    scale_x_continuous(breaks = seq(from=min(nuclear_explosions_proc$year),
                                    to=max(nuclear_explosions_proc$year),
                                    by=5))+
    scale_y_continuous(breaks = seq(from=0,
                                    to=15,
                                    by=1))+
    theme_light()+
    theme(axis.text.x = element_text(angle=90))+
    labs(title="Amount of nuclear explosions by purpose over the years",
         subtitle=paste0("Log count explosions.",
                         "From ",min(nuclear_explosions_proc$year),
                         " to ",max(nuclear_explosions_proc$year)),
         x="",y="",color="Purpose",
         caption = "#tidyTuesday")


ggsave(plot=log_line_plot,filename = "log_line_plot.png",width = 7,height = 5)

```

## items as absolute values



#tweet: 

2019-08-20 #TidyTuesday #rstats Nuclear Explosions!
lets see them by purpose over the years. enjoy :D



# communicate


## resources: 
