---
title: "Week 20181009"
author: "_jas"
date: `r format(Sys.Date())`
output: html_document
---

# Tidy tuesday challenge: Week 2018-10-09: Voter turnout data


## keep it simple:

### Objectives: 

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate

https://rviews.rstudio.com/2017/06/08/what-is-the-tidyverse/ 

# Workflow:

## Import

```{r echo=FALSE,message=FALSE,warning=FALSE}
library(here)
library(readr)
library(tibble)
library(dplyr)
library(tidyr)
library(skimr)#install.packages('skimr')
```

```{r echo=FALSE,message=FALSE,warning=FALSE}

voters_raw <- here("data","2018-10-09","voter_turnout.csv")

```

## Tidy

```{r echo=FALSE,message=FALSE,warning=FALSE}

```


## Understand ( iteration 1 )

### Transform


```{r echo=FALSE,message=FALSE,warning=FALSE}

```

### Visualize

```{r echo=FALSE,message=FALSE,warning=FALSE}

```

```{r echo=FALSE,message=FALSE,warning=FALSE}

```


```{r echo=FALSE,message=FALSE,warning=FALSE}

```


```{r echo=FALSE,message=FALSE,warning=FALSE}


```

### Model

--

## Understand ( iteration 2 )

### Transform


```{r echo=FALSE,message=FALSE,warning=FALSE}

```


### Visualize

### Model

--

## Communicate


### future work: 

