---
title: "Week 20181009"
author: "_jas"
date: 20181012
output: html_document
---

# Tidy tuesday challenge: Week 2018-10-09: Voter turnout data

## keep it simple:

### Objectives: 

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate

https://rviews.rstudio.com/2017/06/08/what-is-the-tidyverse/ 

### this week expected

well its kind of friday so ... just one thing: 

* use the faceted maps through geofacet

http://ryanhafen.com/blog/geofacet

https://hafen.github.io/geofacet/



# Workflow:

## Import

```{r echo=FALSE,message=FALSE,warning=FALSE}
library(here)
library(readr)
library(tibble)
library(dplyr)
library(tidyr)
library(skimr)#install.packages('skimr')
library(ggplot2)
library(geofacet)#install.packages('geofacet')
# library(sf)#install.packages('sf')

```

```{r echo=FALSE,message=FALSE,warning=FALSE}
voters_raw <- readr::read_csv(here::here("data","2018-10-09","voter_turnout.csv"))
```

## Tidy

```{r echo=FALSE,message=FALSE,warning=FALSE}
glimpse(voters_raw)
voters_raw <- voters_raw %>% rename(id = X1)

skimr::skim(voters_raw)

voters_raw %>% filter(!complete.cases(voters_raw))
```


## Understand ( iteration 1 )

### Transform

after a 1st peek , looks like the only problem is that some are missing.
as being voters data might be several stuff,
like that year was no election on that state
or data not gathered in that place. 

so proceeding to make a geofacet directly.
if for some practicar stuff the NA bring some errors with geofacet, 
ill impute 0 votes just to get the facet going, 
in case of a real case, the reason should be further looked out & clarified on the visualization.

as seen on the documentation of geofacet, 
time data should be transformed to something like:

```{r echo=FALSE,message=FALSE,warning=FALSE}
head(state_unemp)
```

  year rate state
1 2000  4.6    AL

```{r echo=FALSE,message=FALSE,warning=FALSE}
# head(voters_raw)
# need to add the state codes
state_df_to_join <- data.frame(state_abb=state.abb,state_name=state.name)

voters_transform_1 <- voters_raw %>% 
    select(year,state,votes,eligible_voters) %>% 
    mutate(vote_ratio=votes/eligible_voters) %>% # adding ratio votes / elegible voters
    left_join(state_df_to_join,by=c('state'='state_name')) %>% 
    rename(st_name=state, state=state_abb)
    
```
some states are NA, as are not needed , will be ignored
```{r echo=FALSE,message=FALSE,warning=FALSE}
voters_transform_1 %>% 
    filter(is.na(state)) %>% 
    group_by(state,st_name) %>% 
    tally()
```



```{r echo=FALSE,message=FALSE,warning=FALSE}
voters_transform_1_no_na <- voters_transform_1 %>% 
    filter(!is.na(state)) 
```


### Visualize

making the geofaceted map on the ratio.

```{r echo=FALSE,message=FALSE,warning=FALSE}
ggplot(voters_transform_1_no_na, aes(year, vote_ratio)) +
  geom_line() +
  facet_geo(~ state, grid = "us_state_grid2", label = "name") +
  scale_x_continuous(labels = function(x) paste0("'", substr(x, 3, 4))) +
  labs(title = "Voter turnout data 1980-2014 (bianual)",
    caption = "Tidytuesday: Week 2018-10-09: 1980-2014 (bianual)",
    x = "Year",
    y = "Ratio (votes / eligible_voters )") +
  theme(strip.text.x = element_text(size = 6))
```

### Model

--

## Communicate

its quite easy to make the map. 
also have a huge amount of differente grids + a grid designer. 
awsome to take into acount for future works.

### learnt:

geofacet depends on sf, sf depends on gdal , gdal need version 2.
https://github.com/r-spatial/sf

also : libmysqlclient-dev its needed by gdal, and not in the line of sf.
added the issue to add the lib on the redme: https://github.com/r-spatial/sf/issues/862

if something goes wrong on linux packages: broken pkg on linux: 
https://unix.stackexchange.com/questions/389156/how-to-fix-held-broken-packages
