---
title: "week 2019-01-22 incarceration USA #rstats #tidytuesday"
author: "_jas"
date: 20190127
output: html_document
---

# Tidy tuesday challenge: Week 2019-01-22

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

### Data:

this week data its related to incarceration in USA by state.

### objectives:

want to test the prohpet library.
- do a TS forecast on some subset of the data.


## import data

```{r echo=FALSE,message=FALSE,warning=FALSE}
library(here)
library(readr)
library(dplyr)
library(tidyr)
library(scales)
library(ggplot2)
library(forcats)#install.packages('forcats')
library(lubridate)
library(visdat)#install.packages('visdat')
library(stringr)
library(naniar)
library(cowplot)
library(rstan) # install.packages("rstan")
library(prophet)# install.packages("prophet")
library(ggrepel)
```



```{r echo=FALSE,message=FALSE,warning=FALSE}
prision_population_file <- here::here("data","2019","2019-01-22","prison_population.csv")
prision_population_raw <- readr::read_csv(prision_population_file)

prision_population_processed <- prision_population_raw %>% 
    mutate(decade = floor(year / 10)*10 ) %>% 
    naniar::add_any_miss()

dplyr::glimpse(prision_population_processed)
skimr::skim(prision_population_processed)
```

we can see that there are a lot of missings on population and prission populations. 
2 things to take in account: 
1) data might not be recorded on digital
2) data might not be existant.

from there we assume that as time goes on , missings will fall down.

```{r echo=FALSE,message=FALSE,warning=FALSE}
prision_population_processed %>% 
    count(decade,any_miss_all,sort = TRUE) %>% 
    ggplot(aes(decade,n,fill=any_miss_all)) +
        geom_col() +
        theme_light() +
        scale_y_continuous(labels = scales::number) +
        coord_flip()+
        labs(title="Evolution of missings by decade",
             subtitle = "1970 - 2016",
             caption = "#tidytuesday week: 2019-01-22",
             x="",y="# of registries",
             fill="is missing ?")
```

Can see that that completeness of registries evolved quite fine since 1970, with a big jump in 1990.


Lets check the smaller datasets to see if this has been handled

```{r echo=FALSE,message=FALSE,warning=FALSE}
prision_summary_file <- here::here("data","2019","2019-01-22","prison_summary.csv")
prision_summary_raw <- readr::read_csv(prision_summary_file)

prision_summary_processed <- prision_summary_raw %>% 
    mutate(decade = floor(year / 10)*10 ) %>% 
    naniar::add_any_miss()

dplyr::glimpse(prision_summary_processed)
skimr::skim(prision_summary_processed)
```

well no missings for prision summary , ill use this to try to use prophet
```{r echo=FALSE,message=FALSE,warning=FALSE}

prision_summary_plot_1 <- prision_summary_processed %>% 
    group_by(decade,urbanicity,pop_category) %>% 
    summarise(sum_decade_rate=sum(rate_per_100000))

# validation of numbers
# prision_summary_processed %>% 
#     filter(decade == 1980, urbanicity=='rural',pop_category=='Black') %>% 
#     pull(rate_per_100000) %>% sum()

# race  / gender  / total

pop_category_gender <- c("Male","Female")
pop_category_total <- c("Total")
pop_category_race_not <- c(pop_category_gender,pop_category_total)

# prision_summary_plot_1 %>% ungroup()%>%  count(pop_category)

plot_race <- prision_summary_plot_1 %>% 
    filter(!(pop_category %in% pop_category_race_not)) %>%
    ggplot(aes(decade,sum_decade_rate,fill=pop_category)) +
        geom_col() +
        # coord_flip() +
    facet_grid(urbanicity~.)+
    labs(x="",y ="",
         subtitle="Rate within a category for prison population per 100,000 people",
         fill="race",
         title="Prision summary, by urbanicity & race")

plot_gender <- prision_summary_plot_1 %>% 
    filter(pop_category %in% pop_category_gender) %>%
    ggplot(aes(decade,sum_decade_rate,fill=pop_category)) +
        geom_col() +
        # coord_flip() +
    facet_grid(urbanicity~.)+
    labs(x="",y ="",
         subtitle="Rate within a category for prison population per 100,000 people",
         fill="gender",
         title="Prision summary, by urbanicity & gender")

plot_total <- prision_summary_plot_1 %>% 
    filter(pop_category %in% pop_category_total) %>%
    ggplot(aes(decade,sum_decade_rate)) +
        geom_col() +
        # coord_flip() +
    facet_grid(urbanicity~.)+
    labs(x="",y ="",
         subtitle="Rate within a category for prison population per 100,000 people",
         title="Prision summary, Total by urbanicity")

plot_race_2<- prision_summary_plot_1 %>% 
    filter(!(pop_category %in% pop_category_race_not)) %>%
    ggplot(aes(decade,sum_decade_rate,fill=pop_category)) +
        geom_col() +
        coord_flip() +
    facet_grid(urbanicity~.)+
    labs(x="",y ="",
         subtitle="Rate within a category for prison population per 100,000 people",
         fill="race",
         title="Prision summary, by urbanicity & race")

plot_gender_2 <- prision_summary_plot_1 %>% 
    filter(pop_category %in% pop_category_gender) %>%
    ggplot(aes(decade,sum_decade_rate,fill=pop_category)) +
        geom_col() +
        coord_flip() +
    facet_grid(urbanicity~.)+
    labs(x="",y ="",
         subtitle="Rate within a category for prison population per 100,000 people",
         fill="gender",
         title="Prision summary, by urbanicity & gender")

plot_total_2 <- prision_summary_plot_1 %>% 
    filter(pop_category %in% pop_category_total) %>%
    ggplot(aes(decade,sum_decade_rate)) +
        geom_col() +
        coord_flip() +
    facet_grid(urbanicity~.)+
    labs(x="",y ="",
         subtitle="Rate within a category for prison population per 100,000 people",
         title="Prision summary, Total by urbanicity")

plot_total
plot_gender
plot_race
plot_total_2
plot_gender_2
plot_race_2
```


```{r echo=FALSE,message=FALSE,warning=FALSE}
prision_summary_prophet_1 <- prision_summary_processed %>% 
    group_by(year,pop_category) %>% 
    summarise(sum_yearly_rate=sum(rate_per_100000)) %>% 
    ungroup() %>%
    mutate(ds=ymd(paste(year,'12','31'))) %>% # for prophet
    rename(y=sum_yearly_rate)

prophet_total <- prision_summary_prophet_1 %>% 
    filter(pop_category=="Total") %>% 
    # filter(urbanicity==param_urbanicity) %>%
    ungroup() %>% 
    select(ds,y)
m <- prophet(prophet_total,seasonality.mode = 'multiplicative')
future <- make_future_dataframe(m, periods = 5,freq ='year' ) # head(future)
forecast <- predict(m, future) # tail(forecast) # plot(m, forecast)
forecast_2 <- forecast %>% mutate(ds2 = as.Date(ds))
join_w_forecast <- prophet_total %>% full_join(forecast_2,by = c("ds"="ds2"))
filtered_points <- join_w_forecast %>% filter(ds>ymd("2015-12-31"))



prision_summary_prophet_urbanicities <- prision_summary_processed %>% 
    group_by(year,urbanicity,pop_category) %>% 
    summarise(sum_yearly_rate=sum(rate_per_100000)) %>% 
    ungroup() %>%
    mutate(ds=ymd(paste(year,'12','31'))) %>% # for prophet
    rename(y=sum_yearly_rate)

make_forecast <- function(data_forecast,param_urbanicity="rural"){
    prophet_total <- data_forecast %>% 
        filter(pop_category=="Total") %>% 
        filter(urbanicity==param_urbanicity) %>%
        ungroup() %>% 
        select(ds,y)
    
    m <- prophet(prophet_total,seasonality.mode = 'multiplicative')
    future <- make_future_dataframe(m, periods = 5,freq ='year' )
    # head(future)
    forecast <- predict(m, future)
    # tail(forecast)
    # plot(m, forecast)
    forecast_2 <- forecast %>% mutate(ds2 = as.Date(ds))
    join_w_forecast <- prophet_total %>% 
        full_join(forecast_2,by = c("ds"="ds2"))
    
    filtered_points <- join_w_forecast %>% filter(ds>ymd("2015-12-31"))
    
    plot_ret <- ggplot(join_w_forecast, aes(x=ds,y=y,label=as.character(yhat) )) + 
        geom_point(aes(y=y),colour="#0072b2") +
        geom_point(data=filtered_points,
                   aes(y=yhat), col="red", size=3) +
        geom_line(aes(y=yhat),colour="#0172b2") + 
        geom_ribbon(aes(ymin=yhat_lower, ymax=yhat_upper), fill="#cce2ef", alpha=0.5) +
        geom_text(data = filtered_points,aes(label=as.character(yhat)))+
        theme_light()+
        labs(x="",y="",
             title= "Forecast of total population in prison",
             subtitle= paste0("Rate per 100000 forecast, for ",param_urbanicity),
             caption = "#tidytuesday week: 2019-01-22")
    plot_ret
}

p_total <- ggplot(join_w_forecast, aes(x=ds,y=y,label=as.character(yhat) )) + 
    geom_point(aes(y=y),colour="#0072b2") +
    geom_point(data=filtered_points,
               aes(y=yhat), col="red", size=3) +
    geom_line(aes(y=yhat),colour="#0172b2") + 
    geom_ribbon(aes(ymin=yhat_lower, ymax=yhat_upper), fill="#cce2ef", alpha=0.5) +
    geom_text(data = filtered_points,aes(label=as.character(yhat)))+
    theme_light()+
    labs(x="",y="",
         title= "Forecast of total population in prison",
         subtitle= "Rate per 100000 forecast, all urbanicities",
         caption = "#tidytuesday week: 2019-01-22")


pl_rur <- make_forecast(prision_summary_prophet_urbanicities,"rural")
pl_sm <- make_forecast(prision_summary_prophet_urbanicities,"small/mid")
pl_su <- make_forecast(prision_summary_prophet_urbanicities,"suburban")
pl_u <- make_forecast(prision_summary_prophet_urbanicities,"urban")

p_total
cowplot::plot_grid(pl_rur,pl_sm)
cowplot::plot_grid(pl_su,pl_u)
  

# prophet_plot_components(m, forecast)
```


# Communicate

- tested the library quite easy to get it going , lot experimenting to be done =) .

- usefull data on: 

- https://facebook.github.io/prophet/docs/non-daily_data.html
- https://facebook.github.io/prophet/docs/quick_start.html#r-api
- https://facebook.github.io/prophet/
- https://quantdare.com/playing-with-prophet-again/
- https://mode.com/example-gallery/forecasting_prophet_r_cookbook


