---
title: "Tidy tuesday challenge: Week 2019-01-29 Federal Research and Development Spending by Agency in USA"
author: "julio"
date: "16 de febrero de 2019"
output: html_document
---

# Tidy tuesday challenge: Week 2019-01-29 Federal Research and Development Spending by Agency in USA

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

### Data:

this week data its related to Federal Research and Development Spending by Agency in USA

### objectives:

- objective not clear, just practice some data & visualization

## import data

```{r echo=FALSE,message=FALSE,warning=FALSE}
library(here)
library(readr)
library(dplyr)
library(tidyr)
library(scales)
library(ggplot2)
library(forcats)#install.packages('forcats')
library(lubridate)
library(visdat)#install.packages('visdat')
library(stringr)
library(naniar)
library(rpart)
library(rpart.plot)
library(cowplot)
library(rstan) # install.packages("rstan")
library(prophet)# install.packages("prophet")
library(ggrepel)
library(anomalize)
```



```{r echo=FALSE,message=FALSE,warning=FALSE}
fed_rd <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-02-12/fed_r_d_spending.csv")
energy_spend <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-02-12/energy_spending.csv")
climate_spend <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-02-12/climate_spending.csv")
```


```{r echo=FALSE,message=FALSE,warning=FALSE}
glimpse(fed_rd) 
glimpse(energy_spend) 
glimpse(climate_spend) 

```

```{r echo=FALSE,message=FALSE,warning=FALSE}
skimr::skim(fed_rd) 
skimr::skim(energy_spend)
skimr::skim(climate_spend)

```

## tidy

```{r echo=FALSE,message=FALSE,warning=FALSE}

is_outlier <- function(x,extreme_only=TRUE) {
    if (extreme_only){
        return (x > quantile(x, 0.75) + 1.5 * IQR(x))
    }
  return(x < quantile(x, 0.25) - 1.5 * IQR(x) | x > quantile(x, 0.75) + 1.5 * IQR(x))
}

totals_climate <- climate_spend %>% 
    group_by(year) %>% 
    summarise(total_spending=sum(gcc_spending))

climate_spend_processed <- climate_spend %>% 
    nest(year,gcc_spending) %>% 
    mutate(db = purrr::map(data,
                           ~ mutate(.x,
                                    outlier=ifelse(is_outlier(gcc_spending,FALSE),
                                                   year, 
                                                   as.numeric(NA))))) %>% 
    unnest() %>% 
    left_join(totals_climate,by = c("year"="year")) %>% 
    mutate(gcc_spending_porc = gcc_spending / total_spending) %>% 
    mutate(department=forcats::fct_reorder(department,gcc_spending_porc)) 

```

## visualize

```{r echo=FALSE,message=FALSE,warning=FALSE}

title_paster <- paste0("Spending by department")
subtitle_paster <-  paste0("Global Climate Change Research Program: ", min(climate_spend$year),"-",max(climate_spend$year))

climate_spend_processed %>% 
    ggplot(aes(x=year,y=gcc_spending,color=department)) +
        geom_line() +
        scale_y_continuous(labels=scales::dollar_format()) +
        labs(title=title_paster,
             subtitle=subtitle_paster,
             caption = "#tidytuesday week 2019-02-12",
             x = "",
             y="spending")

climate_spend_processed %>% 
    ggplot(aes(x=year,y=gcc_spending_porc,fill=department)) +
        geom_col() +
        scale_y_continuous(labels=scales::percent_format()) +
        theme_light() +
        # theme_dark() +
        labs(title=title_paster,
             subtitle=subtitle_paster,
             caption = "#tidytuesday week 2019-02-12",
             x = "",
             y="spending")


climate_spend_processed %>% 
    ggplot(aes(x=year,y=gcc_spending_porc,fill=department)) +
        geom_area() +
        scale_y_continuous(labels=scales::percent_format()) +
        theme_light() +
        # theme_dark() +
        labs(title=title_paster,
             subtitle=subtitle_paster,
             caption = "#tidytuesday week 2019-02-12",
             x = "",
             y="spending")


climate_spend_processed %>% 
    ggplot(aes(y=(gcc_spending/1000000),x=department,fill=department)) +
        geom_boxplot(alpha=0.5) +
        ggrepel::geom_text_repel(aes(label = outlier), na.rm = TRUE, hjust = -0.3) +
        scale_y_continuous(labels=scales::dollar_format()) +
        theme_light() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
        # theme_dark() +
        labs(title=paste0("Vairability of ",title_paster),
             subtitle=subtitle_paster,
             caption = "#tidytuesday week 2019-02-12",
             x = "",
             y="spending in millions")


```




## communicate



### useful resources:

- area inspiration by :

https://youtu.be/KiqpX-gNIS4?t=2280


- generate new variables on nested data: nest > map > unnest help by:

https://stackoverflow.com/questions/46436107/how-to-add-calculated-columns-to-nested-data-frames-list-columns-using-purrr



### not used but usefull for future: 

- nest > map > unnest: help passing a not netsted variable to a nested column

https://stackoverflow.com/questions/53642589/r-add-a-column-in-nested-data-sets

- many models, purr, etc:

https://www.kaggle.com/timib1203/managing-many-models-with-tidyr-purrr-and-broom/report



