---
title: "Tidy tuesday challenge: Week 2019-03-26 Seattle Pet Names"
author: "julio"
date: "30-03-2019"
output: html_document
---

# Tidy tuesday challenge: Week 2019-03-26 Seattle Pet Names

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

### Data:

this week data its related to Seattle Pet Names
https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-03-26

### objectives:

- make the animation of more populars by year.


## details:


## import data

```{r echo=FALSE,message=FALSE,warning=FALSE}
library(dplyr)
library(tidyr)
library(skimr)
library(readr) 
library(ggplot2)
library(forcats)
library(lubridate)
library(gganimate)

```



```{r echo=FALSE,message=FALSE,warning=FALSE}
seattle_pets <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-03-26/seattle_pets.csv") %>%     filter(!is.na(animals_name)) %>% 
    mutate(license_issue_date = lubridate::mdy(license_issue_date)) %>% 
    mutate(license_year = lubridate::year(license_issue_date)) #%>% 

utils::head(seattle_pets)
dplyr::glimpse(seattle_pets)
skimr::skim(seattle_pets)

```

after peekign the data i'll go for the not missing values

```{r echo=FALSE,message=FALSE,warning=FALSE}

# filtered_for_plot <- seattle_pets %>% 
# 
#     
#     # count(animals_name,license_year,sort = TRUE) #%>% filter(n>50)
# 
# tmp <- filtered_for_plot %>% 
#     nest(-license_year) %>% 
#     mutate(top_20_yr = purrr::map(data,.f=function(data_x){
#         data_x %>% count(animals_name,sort = TRUE) %>% arrange(desc(n)) 
#     })) %>% 
#     mutate(plot_yr = purrr::map(top_20_yr,.f=function(data_x,yr=license_year){
#         data_x %>% 
#             ggplot(aes(x=animals_name,y=n,fill=animals_name)) +
#             geom_col()+
#             coord_flip()+
#             labs(title= paste0('Top 20 names in',yr))
#     }))
#     # head(20) %>% 
#     mutate(animals_name=forcats::fct_reorder(animals_name,n)) %>% 
# 
# 
#     # tmp[1,]$top_20_yr[[1]] %>% count(animals_name) %>% filter(n>1)
#     tmp[2,]$plot_yr[[1]]
    
    
    # seattle_pets %>% 
    #     filter(!is.na(animals_name)) %>% 
    #     count(species,animals_name) %>% 
    #     group_by(species,animals_name) %>%
    #     top_n(n = 5, wt = n) %>% 
    #     arrange(species)
# asd <-  seattle_pets %>%
#         count(species,animals_name,sort = TRUE) %>% 
#         arrange(species) %>% 
#         group_by(species) %>% 
#         slice(1:5 )

# asd %>% 
#     ungroup() %>% 
#     mutate(animals_name=fct_reorder(animals_name,n)) %>% 
#     mutate(species=fct_reorder(species,n)) %>% 
#     ggplot(aes(x=animals_name,y=species,fill=species,label=n)) +
#     geom_tile()+
#     geom_text()+
#     coord_flip()+
#     labs(title="Top 5 names by Species", x = "",y="",
#          subtitle="in registered data",
#          caption = "Tidytuesday 2019-03-26: Seattle Pet Names ")+
#     theme_minimal()+
#     theme(legend.position = "none")


asd_yearly <-  seattle_pets %>%
        count(license_year,species,animals_name,sort = TRUE) %>% 
        arrange(license_year,species) %>% 
        group_by(license_year,species) %>% 
        slice(1:5 )


# seattle_pets %>% filter(license_year==2014)

plot_yearly <- asd_yearly %>% 
    filter(license_year>=2014) %>% 
    ungroup() %>% 
    mutate(animals_name=fct_reorder(animals_name,n)) %>% 
    mutate(species=fct_reorder(species,n)) %>% 
    ggplot(aes(x=animals_name,y=species,fill=species,label=as.factor(n))) +
    geom_tile()+
    geom_text()+
    coord_flip()+
    labs(title="Top 5 names by Species by year", x = "",y="",
         subtitle="in registered data {closest_state}",
         caption = "Tidytuesday 2019-03-26: Seattle Pet Names ")+
    theme_minimal()+
    theme(legend.position = "none")+
    transition_states(license_year,
                    transition_length = 1,
                    state_length =8)

plot_yearly


```



## communicate


### general view:


### summary: 

wanted to do the bar race, but mutated into tile map animation. to see  if got shared names across species.