---
title: "week 2019-01-29 #rstats #tidytuesday mortgages"
author: "_jas"
date: 20190129
output: html_document
---

# Tidy tuesday challenge: Week 2019-01-29 mortgages

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

### Data:

this week data its related to mortgages & recesions in USA.

### objectives:

- use the anomalize package on mortgages

## import data

```{r echo=FALSE,message=FALSE,warning=FALSE}
library(here)
library(readr)
library(dplyr)
library(tidyr)
library(scales)
library(ggplot2)
library(forcats)#install.packages('forcats')
library(lubridate)
library(visdat)#install.packages('visdat')
library(stringr)
library(naniar)
library(rpart)
library(rpart.plot)
library(cowplot)
library(rstan) # install.packages("rstan")
library(prophet)# install.packages("prophet")
library(ggrepel)
library(anomalize)
```



```{r echo=FALSE,message=FALSE,warning=FALSE}
mortgage_file <- here::here("data","2019","2019-02-05","mortgage.csv")
mortgage_raw <- readr::read_csv(mortgage_file)

mortgage_process <- mortgage_raw %>% 
    mutate(decade = floor(year / 10)*10 ) %>% 
    naniar::add_any_miss()

dplyr::glimpse(mortgage_raw)
skimr::skim(mortgage_raw)
```

looking at data , and time span ... i'll focus on looking for anomalies on fixed_rate_30_yr

```{r echo=FALSE,message=FALSE,warning=FALSE}

mortgage_anomalies <- mortgage_raw %>% select(date,fixed_rate_30_yr)

mortgage_anomalies %>% 
    ggplot(aes(x=date,y=fixed_rate_30_yr)) +
    geom_line()

anomalize::time_frequency(mortgage_anomalies, period = "auto")

# mortgage_anomalies %>%  filter(year(date)==1972)

mortgage_anom_4_plot <- mortgage_anomalies %>% 
    time_decompose(fixed_rate_30_yr, 
                   method = "stl", # stl = seasonal decomposition.
                     trend =  paste0(13," weeks")
                       # "auto"
                   ) %>%
      anomalize(remainder, method = "gesd") %>%
time_recompose() 

mortgage_anom_4_plot %>% 
    plot_anomalies(time_recomposed = TRUE) %>% 
    labs(x="")

mortgage_anom_4_plot %>% 
    plot_anomaly_decomposition() 

```

# removing years that are not complete

```{r echo=FALSE,message=FALSE,warning=FALSE}

mgyearly <- mortgage_anomalies %>%  mutate(year = year(date))
range(mgyearly$year)

years_to_remove <- mgyearly %>% count(year) %>% filter(n<52) %>%  pull(year)

years_removed <- mgyearly %>% filter(! (year %in% years_to_remove))

years_removed %>%  
    # filter(year==1972) %>%  
    time_decompose(fixed_rate_30_yr, 
                   method = "stl", # stl = seasonal decomposition.
                     trend = "13 weeks"
                   ) %>%
      anomalize(remainder, method = "gesd") %>%
time_recompose()  
    
    
years_removed %>%  
    # filter(year==1972) %>%  
    time_decompose(fixed_rate_30_yr, 
                   method = "stl", # stl = seasonal decomposition.
                     trend = "13 weeks"
                   ) %>%
      anomalize(remainder, method = "gesd") %>%
time_recompose() %>% 
    
    plot_anomalies(time_recomposed = TRUE)


yr_plot <- years_removed %>%  
    # filter(year==1972) %>%  
    time_decompose(fixed_rate_30_yr, 
                   method = "stl", # stl = seasonal decomposition.
                     trend = "13 weeks"
                   ) %>%
      anomalize(remainder, method = "gesd") %>%
time_recompose() 



yr_plot %>% 
ggplot(aes(date, observed,colour=anomaly)) +
    geom_point(alpha = 0.25) +
    scale_color_manual(values = c("Yes"="#e31a1c","No"="#2c3e50"))+

    # geom_point(data=subset(yr_plot, anomaly == "No"),color = "#2c3e50", alpha = 0.25) +
    # geom_point(data=subset(yr_plot, anomaly == "Yes"),color = "#e31a1c", alpha = 0.25) +
    # facet_wrap(~ package, scale = "free_y", ncol = 3) +
    theme_light() +
    # scale_y_continuous(labels = scales::percent_format())+
    scale_x_date(date_labels = "%Y-%m",date_breaks = "5 years") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = "USA Mortgages: anomalies on Fixed rate 30 year mortgage (%) ",
         subtitle = "from 1972-2018 [ decompose: stl-13 weeks & anomalies: gesd ]",
         y="Fixed rate 30 year mortgage (%)",x="",
         caption = "Data from TidyTuesday 2019-02-05 : USA Mortgages")

```


# communicate

time counting ... so got to adjust schedule. 

only for mortgages anomalies, 

a way to improve its to put recesions periods on the same plot. 


## some usefull data in: 

- anomalize package and ggplot tuning with that data.
- https://github.com/business-science/anomalize

