---
title: "Week 20181023"
author: "_jas"
date: 20181023
output: html_document
---

# Tidy tuesday challenge: Week 2018-10-23: Horror Movies and Profit

* article: https://fivethirtyeight.com/features/scary-movies-are-the-best-investment-in-hollywood/

* data: movie_profit.csv

* link to readme: https://github.com/rfordatascience/tidytuesday/tree/master/data/2018-10-23

## keep it simple:

### Objectives: 

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate

https://rviews.rstudio.com/2017/06/08/what-is-the-tidyverse/ 

### this week expected

* try to apply some hints from last week
* think questions to answer with the data given
* try to drive analysis for what i'll probably like or think


# Workflow:

## Import

```{r echo=FALSE,message=FALSE,warning=FALSE}
library(here)
library(readr)
library(tibble)
library(dplyr)
library(tidyr)
library(skimr)#install.packages('skimr')
library(scales)
library(ggplot2)
library(forcats)
library(lubridate)
# library(geofacet)#install.packages('geofacet')
# library(sf)#install.packages('sf')

```

```{r echo=FALSE,message=FALSE,warning=FALSE}
horror_movies_raw <- readr::read_csv(here::here("data","2018-10-23","movie_profit.csv"))
```

## Tidy

```{r echo=FALSE,message=FALSE,warning=FALSE}
glimpse(horror_movies_raw)
horror_movies <- horror_movies_raw %>% 
    rename(id = X1) %>% 
    mutate(release_date_d=mdy(release_date)) %>% 
    select (-release_date)

skimr::skim(horror_movies)

horror_movies %>% filter(!complete.cases(horror_movies))
```


## Understand ( iteration 1 )

### Transform


```{r echo=FALSE,message=FALSE,warning=FALSE}

```

```{r echo=FALSE,message=FALSE,warning=FALSE}

```

```{r echo=FALSE,message=FALSE,warning=FALSE}

```



```{r echo=FALSE,message=FALSE,warning=FALSE}

```


### Visualize

making the geofaceted map on the ratio.

```{r echo=FALSE,message=FALSE,warning=FALSE}

```

### Model

--

## Communicate


### learnt:
