---
title: "Week 20181023"
author: "_jas"
date: 20181023
output: html_document
---

# Tidy tuesday challenge: Week 2018-10-23: Horror Movies and Profit

* article: https://fivethirtyeight.com/features/scary-movies-are-the-best-investment-in-hollywood/

* data: movie_profit.csv

* link to readme: https://github.com/rfordatascience/tidytuesday/tree/master/data/2018-10-23

## keep it simple:

### Objectives: 

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate

https://rviews.rstudio.com/2017/06/08/what-is-the-tidyverse/ 

### this week expected

* try to apply some hints from last week
* think questions to answer with the data given
* try to drive analysis for what i'll probably like or think


# Workflow:

## Import

```{r echo=FALSE,message=FALSE,warning=FALSE}
library(here)
library(readr)
library(tibble)
library(dplyr)
library(tidyr)
library(skimr)#install.packages('skimr')
library(scales)
library(ggplot2)
library(forcats)
library(lubridate)
library(cowplot)#install.packages('cowplot')
# library(geofacet)#install.packages('geofacet')
# library(sf)#install.packages('sf')

```

```{r echo=FALSE,message=FALSE,warning=FALSE}
horror_movies_raw <- readr::read_csv(here::here("data","2018-10-23","movie_profit.csv"))
```

## Tidy

```{r echo=FALSE,message=FALSE,warning=FALSE}
glimpse(horror_movies_raw)
horror_movies <- horror_movies_raw %>% 
    rename(id = X1) %>% 
    mutate(release_date_d=mdy(release_date)) %>% 
    select (-release_date)

skimr::skim(horror_movies)

horror_movies %>% filter(!complete.cases(horror_movies))
```

### valid date range ?

```{r echo=FALSE,message=FALSE,warning=FALSE}

range(horror_movies$release_date_d )
```

looks like there is future movies ... well they might be still in production.
also ... to be aware of sample ... which is the sample date cutpoint ?

wich is the cutpoint of the sample ?

evaluated by year can thing on 2018 , but lets see by month, as 2018 its not finished yet.


```{r echo=FALSE,message=FALSE,warning=FALSE}

horror_movies %>% 
    # filter(domestic_gross == 0) %>%  
    arrange(desc(release_date_d)) %>% 
    select(movie, release_date_d, production_budget,domestic_gross,worldwide_gross) %>% 
    mutate(year=year(release_date_d),month = month(release_date_d)) %>% 
    group_by(year,month) %>% 
    summarise(number_of_movies=n(),
              domestic_gross_total = sum(domestic_gross),
              worldwide_gross_total= sum(worldwide_gross),
              production_budget_total=sum(production_budget)) %>% 
    arrange(desc(year),desc(month)) %>% 
    head(10)
```

looks like today might be a good cut point ( today = "2018-10-23")

which movies should be excluded  ?
```{r echo=FALSE,message=FALSE,warning=FALSE}
horror_movies %>% filter(release_date_d > ymd("2018-10-23")) %>%  select(movie, release_date_d, production_budget)
```


for this analysis this 4 movies will be excluded.
date its taken as ( today = "2018-10-23")

```{r echo=FALSE,message=FALSE,warning=FALSE}
horror_movies_bkp <- horror_movies
horror_movies <- horror_movies %>% filter(release_date_d <= ymd("2018-10-23"))
```

## future work:

its an interesting point of view as ... 
can movies be capitalized before release date ?

my guess sure thing, pre event marketing & other hype stuff.

how much time before release can get this pre event capitalizations going ?
can it be measured in someway ?



## Understand ( iteration 1 )

* wich are the most profitable movies of all time ? ( gross / budget | ww_gross / budget)
* how ratio vary over decades ( gross / budget | ww_gross / budget)
* how genres vary over decades ( count movies / genre / time span)
* how genres vary over decades ( profitable movies / genre / time span)

for all this quiestions, none is affected by the NA's
for later we'll like to see what about NA's ( mpaa_rating / distributor) ; may be related to time stuff.


### Transform

to answer profit answers gotta make some columns 
to answer time span ( decade ) answers we also gotta add the decade to the movies

```{r echo=FALSE,message=FALSE,warning=FALSE}

horror_movies <- horror_movies %>% 
    mutate(usa_profit_ratio = domestic_gross / production_budget) %>% 
    mutate(ww_profit_ratio = worldwide_gross / production_budget) %>% 
    mutate(usa_profit = domestic_gross - production_budget) %>% 
    mutate(ww_profit = worldwide_gross - production_budget) %>% 
    mutate(usa_ww_gross_ratio = domestic_gross / worldwide_gross) %>% 
    mutate(usa_ww_profit_ratio = usa_profit / ww_profit) %>% 
    mutate(usa_ww_profit_ratio_ratio = usa_profit_ratio / ww_profit_ratio) %>% 
    mutate(decade = as.factor(floor(year(release_date_d)/10)*10)) %>% 
    mutate(month = as.factor(month(release_date_d)))

```


#### wich are the most profitable movies of all time ? ( gross / budget | ww_gross / budget)


```{r echo=FALSE,message=FALSE,warning=FALSE}




# movie - genre - location - type - value : gotta practice gather.
# horror_movies %>% 
#     select(movie,genre, usa_profit_ratio,ww_profit_ratio,
#            usa_profit,ww_profit,
#            usa_ww_profit_ratio,usa_ww_profit_ratio_ratio) %>% 
#     gather(location,value,genre, usa_profit_ratio,ww_profit_ratio,
#            usa_profit,ww_profit,
#            usa_ww_profit_ratio,usa_ww_profit_ratio_ratio)

plot_usa_profit_genre <- horror_movies %>% 
    ggplot(aes(genre,usa_profit/1000,fill=genre)) +
    geom_boxplot() +
    scale_y_continuous(labels = dollar_format())+
    labs(title="All time: Movies",
         subtitle="USA profit by genre",
         x="",
         y="gross - prod budget, in K USD")+
    coord_flip() +  # rotate the boxplot as there is not clear 
    theme_light() + # theme light
    theme(legend.position = "none")

plot_ww_profit_genre <- horror_movies %>% 
    ggplot(aes(genre,ww_profit/1000,fill=genre)) +
    geom_boxplot() +
    scale_y_continuous(labels = dollar_format())+
    labs(title="All time: Movies ",
         subtitle="World Wide profit by genre",
         x="",
         y="gross - prod budget, in K USD")+
    coord_flip() +  # rotate the boxplot as there is not clear 
    theme_light() + # theme light
    theme(legend.position = "none")

plot_grid(plot_usa_profit_genre, plot_ww_profit_genre)

```

if movies are evaluated by profit, top outliers will be more successfull; down outliers will be big loses.
ranges move from -200k to 1.2M aproximattly,
ww vs usa, tend to be similar, but less outliers on the loses. only drama got an outlier on the lower bound if seen ww.

### lets check the WW - usa: gross / budget ratio


```{r echo=FALSE,message=FALSE,warning=FALSE}
plot_usa_profit_ratio_genre <- horror_movies %>% 
    ggplot(aes(genre,usa_profit_ratio,fill=genre)) +
    geom_boxplot() +
    # scale_y_continuous(labels = dollar_format())+
    labs(title="All time: Movies",
         subtitle="USA profit ratio by genre",
         x="",
         y="gross / prod budget, times budget")+
    coord_flip() +  # rotate the boxplot as there is not clear 
    theme_light() + # theme light
    theme(legend.position = "none")

plot_ww_profit_ratio_genre <- horror_movies %>% 
    ggplot(aes(genre,ww_profit_ratio,fill=genre)) +
    geom_boxplot() +
    # scale_y_continuous(labels = numdollar_format())+
    labs(title="All time: Movies ",
         subtitle="World Wide profit ratio by genre",
         x="",
         y="gross / prod budget, times budget")+
    coord_flip() +  # rotate the boxplot as there is not clear 
    theme_light() + # theme light
    theme(legend.position = "none")

plot_grid(plot_usa_profit_ratio_genre, plot_ww_profit_ratio_genre)
```

if movies are evaluated by profit ratio , top outliers will be the more succesfull. 
we cross check that horror movies are the more profitable ones, follwoed by drama, and comedy.

#### Lets see times budget ratio on top 20 movies.

```{r echo=FALSE,message=FALSE,warning=FALSE}

plot_ww_profit_ratio_top_20_order_usa <- horror_movies %>% 
    select(movie,usa_ww_profit_ratio_ratio,usa_profit_ratio,ww_profit_ratio) %>% 
    arrange(desc(usa_profit_ratio)) %>% 
    head(20) %>% 
    mutate(movie=fct_reorder(movie,usa_profit_ratio)) %>%
    gather(ratio_type,value,usa_profit_ratio,ww_profit_ratio) %>% 
    ggplot(aes(movie,value,fill=ratio_type)) +
    geom_col() +
    scale_y_continuous(labels = comma_format())+
        labs(title="All time: Movies top #20 ",
         subtitle="profit ratio by usa/ww order usa",
         x="",
         y="gross / prod budget, times budget")+
    coord_flip() +  # rotate the boxplot as there is not clear 
    theme_light()


plot_ww_profit_ratio_top_20_order_ww <- horror_movies %>% 
    select(movie,usa_ww_profit_ratio_ratio,usa_profit_ratio,ww_profit_ratio) %>% 
    arrange(desc(ww_profit_ratio)) %>% 
    head(20) %>% 
    mutate(movie=fct_reorder(movie,ww_profit_ratio)) %>%
    gather(ratio_type,value,usa_profit_ratio,ww_profit_ratio) %>% 
    ggplot(aes(movie,value,fill=ratio_type)) +
    geom_col() +
    scale_y_continuous(labels = comma_format())+
        labs(title="All time: Movies top #20 ",
         subtitle="profit ratio by usa/ww order ww",
         x="",
         y="gross / prod budget, times budget")+
    coord_flip() +  # rotate the boxplot as there is not clear 
    theme_light()

# plot_grid(plot_ww_profit_ratio_top_20_order_usa, plot_ww_profit_ratio_top_20_order_ww,ncol = 1)
plot_ww_profit_ratio_top_20_order_usa
plot_ww_profit_ratio_top_20_order_ww


```

#### Tail 20 movies

when it comes to ratio, several got 0 times, so its not worth it seeing it that way 
we see the raw profit this for tails.

```{r echo=FALSE,message=FALSE,warning=FALSE}

plot_ww_profit_tail_20_order_usa <- horror_movies %>% 
    select(movie,usa_ww_profit_ratio_ratio,usa_profit,ww_profit) %>% 
    arrange(desc(usa_profit)) %>% 
    tail(20) %>% 
    arrange(usa_profit) %>% 
    mutate(movie=fct_reorder(movie,usa_profit)) %>%
    gather(ratio_type,value,usa_profit,ww_profit) %>% 
    ggplot(aes(movie,value/1000,fill=ratio_type)) +
    geom_col() +
    # scale_y_continuous(labels = comma_format())+
    scale_y_continuous(labels = dollar_format())+
        labs(title="All time: Movies tail #20 ",
         subtitle="profit ratio by usa/ww order usa",
         x="",
         y="gross / prod budget, K USD")+
    coord_flip() +  # rotate the boxplot as there is not clear 
    theme_light()


plot_ww_profit_tail_20_order_ww <- horror_movies %>% 
    select(movie,usa_ww_profit_ratio_ratio,usa_profit,ww_profit) %>% 
    arrange(desc(ww_profit)) %>% 
    tail(20) %>% 
    arrange(ww_profit) %>% 
    mutate(movie=fct_reorder(movie,ww_profit)) %>%
    gather(ratio_type,value,usa_profit,ww_profit) %>% 
    ggplot(aes(movie,value/1000,fill=ratio_type)) +
    geom_col() +
    scale_y_continuous(labels = dollar_format())+
    # scale_y_continuous(labels = comma_format())+
        labs(title="All time: Movies tail #20 ",
         subtitle="profit ratio by usa/ww order ww",
         x="",
         y="gross / prod budget, K USD")+
    coord_flip() +  # rotate the boxplot as there is not clear 
    theme_light() 

# plot_grid(plot_ww_profit_ratio_top_20_order_usa, plot_ww_profit_ratio_top_20_order_ww,ncol = 1)
plot_ww_profit_tail_20_order_usa
plot_ww_profit_tail_20_order_ww

```
as we can see if we order by the USA profit ; we see that some loses on USA might not be doomed to failure in Worldwide.


#### ADDING TIME 

#### any favorite publication month ?

```{r echo=FALSE,message=FALSE,warning=FALSE}
month_view <- horror_movies %>% 
    
    group_by(month) %>% 
    summarise(n=n(),
              usa_profit_ratio_total=sum(usa_profit_ratio),
              ww_profit_ratio_total=sum(ww_profit_ratio),
              usa_profit_total=sum(usa_profit),
              ww_profit_total=sum(ww_profit))
month_view %>%
    ggplot(aes(month,n, fill=month)) +
    geom_col()+
    
    # scale_y_continuous(labels = comma_format())+
        labs(title="All time: Movies Release months ",
         subtitle="# relseases by month",
         x="",
         y="# relseases")+
    coord_flip() +  # rotate the boxplot as there is not clear 
    theme_light()


    
```

what about profit by released on a month ?

```{r echo=FALSE,message=FALSE,warning=FALSE}

profit_month_usa <- month_view %>%
    ggplot(aes(month,usa_profit_ratio_total, fill=month)) +
    geom_col()+
    
    # scale_y_continuous(labels = comma_format())+
        labs(title="All time: Movies Release months ",
         subtitle="profit usa ",
         x="",
         y="times budget")+
    coord_flip() +  # rotate the boxplot as there is not clear 
    theme_light()

profit_month_ww<- month_view %>%
    ggplot(aes(month,ww_profit_ratio_total, fill=month)) +
    geom_col()+
    
    # scale_y_continuous(labels = comma_format())+
        labs(title="All time: Movies Release months ",
         subtitle="profit ww ",
         x="",
         y="times budget")+
    coord_flip() +  # rotate the boxplot as there is not clear 
    theme_light()

plot_grid(profit_month_usa, profit_month_ww)

```

#### last of today: over decades
```{r echo=FALSE,message=FALSE,warning=FALSE}

decades_view <- horror_movies %>% 
    
    group_by(decade) %>% 
    summarise(n=n(),
              usa_profit_ratio_total=sum(usa_profit_ratio),
              ww_profit_ratio_total=sum(ww_profit_ratio),
              usa_profit_total=sum(usa_profit),
              ww_profit_total=sum(ww_profit))
decades_view %>%
    ggplot(aes(decade,n, fill=decade)) +
    geom_col()+
    
    # scale_y_continuous(labels = comma_format())+
        labs(title="All time: Movies Release months ",
         subtitle="# relseases by decade",
         x="",
         y="# relseases")+
    coord_flip() +  # rotate the boxplot as there is not clear 
    theme_light()


decades_view %>%
    ggplot(aes(decade,usa_profit_ratio_total, fill=decade)) +
    geom_col()+
    
    # scale_y_continuous(labels = dollar_format())+
        labs(title="All time: Movies  ",
         subtitle="# USA times budget by decade",
         x="",
         y="# times budget")+
    coord_flip() +  # rotate the boxplot as there is not clear 
    theme_light()


decades_view %>%
    ggplot(aes(decade,ww_profit_ratio_total, fill=decade)) +
    geom_col()+
    
    # scale_y_continuous(labels = dollar_format())+
        labs(title="All time: Movies decades ",
         subtitle="# WW times budget by decade",
         x="",
         y="# times budget")+
    coord_flip() +  # rotate the boxplot as there is not clear 
    theme_light()

decades_view %>%
    ggplot(aes(decade,ww_profit_total/1000000, fill=decade)) +
    geom_col()+
    
    scale_y_continuous(labels = dollar_format())+
        labs(title="All time: Movies decades ",
         subtitle="# WW profit by decade",
         x="",
         y="USD Millons profit")+
    coord_flip() +  # rotate the boxplot as there is not clear 
    theme_light()



```

#### decades & genres

will  be awsome to do the percent of movies, but too much time invsted ATM.

```{r echo=FALSE,message=FALSE,warning=FALSE}
gen_decades_view <- horror_movies %>% 
    
    group_by(genre,decade) %>% 
    summarise(n=n(),
              usa_profit_ratio_total=sum(usa_profit_ratio),
              ww_profit_ratio_total=sum(ww_profit_ratio),
              usa_profit_total=sum(usa_profit),
              ww_profit_total=sum(ww_profit))
gen_decades_view %>%
    ggplot(aes(decade,ww_profit_total/1000000, fill=genre)) +
    geom_col()+
    
    scale_y_continuous(labels = dollar_format())+
        labs(title="All time: Movies decades & genre ",
         subtitle="# WW profit by decade & genre",
         x="",
         y="USD Millons profit")+
    coord_flip() +  # rotate the boxplot as there is not clear 
    theme_light()


gen_decades_view %>%
    ggplot(aes(decade,ww_profit_ratio_total, fill=genre)) +
    geom_col()+
    
    # scale_y_continuous(labels = dollar_format())+
        labs(title="All time: Movies decades & genre ",
         subtitle="# WW times budget by decade & genre ",
         x="",
         y="# times budget")+
    coord_flip() +  # rotate the boxplot as there is not clear 
    theme_light()


gen_decades_view %>%
    ggplot(aes(decade,n, fill=genre)) +
    geom_col()+
    
    # scale_y_continuous(labels = dollar_format())+
        labs(title="All time: Movies decades & genre ",
         subtitle="# movies by decade & genre",
         x="",
         y="# movies")+
    coord_flip() +  # rotate the boxplot as there is not clear 
    theme_light()

```

### Visualize

making the geofaceted map on the ratio.



### Model

--

## Communicate


### learnt:

practiced ggplot bars & some boxplot. 

it was quite interesting :D

wanted to do more but well. enough for today :p 
