---
title: "week 2018-12-25 no dataset / pick a dataset"
author: "_jas"
date: 20181225
output: html_document
---

# Tidy tuesday challenge: Week 2018-12-25

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

### Data:

No dataset / free week / use previous / your own

data choosen: 

argentina wild fires: 
All data: http://datos.ambiente.gob.ar/dataset/incendios-forestales


i picked 2:

* causes & quantity per province ('states') : "Causas y cantidad de incendios por provincia" 
* affected area per province ('states') and vegetation type : "Superficie afectada por incendios por provincia y tipo de vegetación"

in case data portal wont work, i upload to my github.

### objectives:

unify the data on a single data set. 
do some map stuff, may be a map animation.


## import data

```{r echo=FALSE,message=FALSE,warning=FALSE}
library(here)
library(readr)
library(dplyr)
library(tidyr)
library(janitor)#install.packages('janitor')
library(scales)
library(ggplot2)
library(forcats)#install.packages('forcats')
library(lubridate)
library(visdat)#install.packages('visdat')
library(stringr)
# library(DataExplorer)
# library(ggrepel)#install.packages('ggrepel')
# library(rebus)
# library(igraph)
# library(ggraph)
# library(tidygraph)#install.packages('tidygraph')
library(gganimate)

library(maptools)
library(rgdal)
library(raster)
library(rgeos)
library(ggplot2)
library(ggalt) #; install.packages("ggalt")
library(ggthemes)#; install.packages("ggthemes")
library(viridis)
library(magrittr)

```

```{r echo=FALSE,message=FALSE,warning=FALSE}
file_name_1 <- here::here("jas1_weeks","2018-12-25","incendios-cantidad-causas-provincia.csv")
causes_raw <- readr::read_delim(file_name_1,delim = ";")
glimpse(causes_raw)

file_name_2 <- here::here("jas1_weeks","2018-12-25","superficie-incendiada-provincias-tipo-de-vegetacion.csv")
area_raw <- readr::read_delim(file_name_2,delim = ";")
glimpse(area_raw)


causes_processed <- causes_raw %>% 
    rename_all(gsub, pattern = 'incendio_', replacement = '') %>% 
    rename_all(gsub, pattern = '_numero', replacement = '_causa') 

area_processed <- area_raw %>% rename_all(gsub, pattern = 'superficie_afectada_por_incendios_', replacement = '')%>% 
    rename_all(gsub, pattern = '_hectareas', replacement = '_area') 

skimr::skim(causes_processed)
skimr::skim(area_processed)

```

```{r echo=FALSE,message=FALSE,warning=FALSE}
causes_processed %>% count(anio, provincia)
area_processed %>% count(anio, provincia)

fuegos_processed <- causes_processed %>% 
    left_join(area_processed, by=c("anio","provincia"))

# skimr::skim(fuegos_processed)
# 
```

```{r echo=FALSE,message=FALSE,warning=FALSE}
    # area_processed %>% 
    # gather(key,value,-anio,-provincia) %>% 
    # filter(key!="total_area") %>% 
    # rename_all(gsub, pattern = '_area', replacement = '') %>% 
    # # filter(anio==1993) %>% 
    # mutate(provincia=fct_reorder(provincia,desc(provincia))) %>% 
    # ggplot(aes(x=provincia, y=value,fill=key)) +
    #     geom_col()+
    #     scale_y_continuous(labels = scales::number) +
    #     labs(title="Affected Area by province",
    #          x="",
    #          y="#hectareas")+
    #     coord_flip()+
    #     theme_light()
    # 
    # area_processed %>% 
    # gather(key,value,-anio,-provincia) %>% 
    # filter(key!="total_area") %>% 
    # rename_all(gsub, pattern = '_area', replacement = '') %>% 
    # # filter(anio==1993) %>% 
    # mutate(provincia=fct_reorder(provincia,desc(provincia))) %>% 
    # mutate(anio=factor(anio)) %>% 
    # 
    # ggplot(aes(x=provincia, y=value,fill=anio)) +
    #     geom_col()+
    #     scale_y_continuous(labels = scales::number) +
    #     labs(title="Affected Area by province",
    #          x="",
    #          y="#hectareas")+
    #     coord_flip()+
    #     theme_light()

```

###  Total  wildfire causes over the years

```{r echo=FALSE,message=FALSE,warning=FALSE}
    causes_processed %>% 
    dplyr::select(-anio) %>% 
    group_by(provincia ) %>% 
    summarize_all(sum) %>% 
    rename_all(gsub, pattern = '_causa', replacement = '') %>% 
    mutate(provincia=fct_reorder(provincia,total)) %>% 
    gather(key,value,-provincia) %>% 
    filter(key!="total") %>% 
    ggplot(aes(x=provincia, y=value,fill=key)) +
        geom_col()+
        scale_y_continuous(labels = scales::number) +
        labs(title="Total Wildfires by province 1993-2016",
             x="",
             y="#wildfires",
             fill="cause",
             caption = "http://datos.ambiente.gob.ar/dataset/incendios-forestales")+
        coord_flip()+
        theme_light()
```


## total area burnt by province & vegetation type

```{r echo=FALSE,message=FALSE,warning=FALSE}
    area_processed %>% 
    dplyr::select(-anio) %>% 
    group_by(provincia) %>% 
    summarize_all(sum) %>% 
    rename_all(gsub, pattern = '_area', replacement = '') %>% 
    mutate(provincia=fct_reorder(provincia,total)) %>% 
    gather(key,value,-provincia) %>% 
    filter(key!="total") %>% 
    ggplot(aes(x=provincia, y=value,fill=key)) +
        geom_col()+
        scale_y_continuous(labels = scales::number) +
        labs(title="Total area burnt by province 1993-2016",
             subtitle="remember it can be burnt again the same spot",
             x="",
             y="# hectares",
             fill="vegetation type",
             caption = "http://datos.ambiente.gob.ar/dataset/incendios-forestales")+
        coord_flip()+
        theme_light()


```


```{r}
#https://github.com/thomasp85/gganimate
total_wildfires_arg_gif <- causes_processed %>% 
    mutate(provincia=fct_reorder(provincia,total_causa)) %>%
    rename_all(gsub, pattern = '_causa', replacement = '') %>% 
    gather(key,value,-provincia,-anio) %>% 
    filter(key!="total") %>% 
ggplot(aes(x=provincia,
           y=value,
           fill=key)) +
    geom_col()+
    coord_flip()+
    theme_light()+
    scale_y_continuous(labels = scales::number) +
  labs(title = 'Argentina: Total Wildfires by province; 1993-2016',
       subtitle='Year: {frame_time}', 
       x = '', 
       y = '# wildfire',
       fill="cause",
       caption = "http://datos.ambiente.gob.ar/dataset/incendios-forestales") +
  transition_time(anio) +
  ease_aes('linear')

```


```{r}
#https://github.com/thomasp85/gganimate
total_wildfires_area_arg_gif <- area_processed %>% 
    mutate(provincia=fct_reorder(provincia,total_area)) %>%
    rename_all(gsub, pattern = '_area', replacement = '') %>% 
    gather(key,value,-provincia,-anio) %>% 
    filter(key!="total") %>% 
ggplot(aes(x=provincia,
           y=value,
           fill=key)) +
    geom_col()+
    scale_y_continuous(labels = scales::number) +
    coord_flip()+
    theme_light()+
  labs(title = 'Argentina: Total Wildfires burnt area by province; 1993-2016',
       subtitle='Year: {frame_time} ; remember it can be burnt again the same spot', 
       x = '', 
       y = '# hectares',
       fill="vegetation type",
       caption = "http://datos.ambiente.gob.ar/dataset/incendios-forestales") +
  transition_time(anio) +
  ease_aes('linear')

```


# Area burnt over time

```{r echo=FALSE,message=FALSE,warning=FALSE}
# # load argentina provinces sf
# # testing: https://stackoverflow.com/questions/37517731/decreasing-processing-time-for-shp-map-choropleth
# # as stated in the other answer, this is the same as your shapefile
# arg_adm <- raster::getData('GADM', country='ARG', level=1)
# 
# # make the polygons a bit less verbose
# gSimplify(arg_adm, 0.01, topologyPreserve=TRUE) %>% 
#   SpatialPolygonsDataFrame(dat=arg_adm@data) -> arg_adm
# 
# # turn them into a data frame
# arg_map <- fortify(arg_adm, region="NAME_1")
# 
# # use a gd projection for this region
# arg_proj <- "+proj=aeqd +lat_0=-37.869859624840764 +lon_0=-66.533203125"
# 
# # reproducibly simulate some data
# set.seed(1492)
# # 
# prov <- c("Buenos Aires", "Córdoba", "Catamarca", "Chaco", "Chubut",
#                           "Ciudad de Buenos Aires", "Corrientes", "Entre Ríos", "Formosa", 
#                           "Jujuy", "La Pampa", "La Rioja", "Mendoza", "Misiones", "Neuquén", 
#                           "Río Negro", "Salta", "San Juan", "San Luis", "Santa Cruz", 
#                           "Santa Fe", "Santiago del Estero", "Tierra del Fuego", "Tucumán")
# anios <- c("1990","2000", "2010")
# 
# puntos <- data.frame(id=prov,anio=anios,value=sample(100, length(anios)*length(prov)),stringsAsFactors = FALSE)


# 
# 
# puntos2 <-  area_processed %>% 
#     # dplyr::select(-anio) %>% 
#     mutate(anio=as.character(anio)) %>% 
#     mutate(bins=cut_interval(total_area,5,length = )) %>%  # 0-500;500-1500;1500-5000;5000-20000;+20000  .856 4514 24150
#     # group_by(anio,provincia) %>% 
#     # summarize_all(sum) %>% 
#     rename_all(gsub, pattern = '_area', replacement = '') %>% 
#     mutate(provincia=fct_reorder(provincia,total)) 
#     
#     levels(puntos2$bins)
# puntos2 %>%  
#     group_by(bins) %>% 
#     summarise(mini=min(total),maxi=max(total),n=n()) %>% 
#     mutate(bins_2=paste0(mini,"-",maxi))
# 
# 
# #856 4514 24150
# boxplot(puntos2$total)
#  skimr::skim(puntos2$total)
#  boxplot(puntos3$total)
#  skimr::skim(puntos3$total)
#  
# glimpse(puntos2)
# 
# puntos3 <- causes_processed %>% 
#     mutate(anio=as.character(anio)) %>% 
#     # group_by(anio,provincia) %>% 
#     # summarize_all(sum) %>% 
#     rename_all(gsub, pattern = '_causa', replacement = '') %>% 
#     mutate(provincia=fct_reorder(provincia,total)) %>% 
#     # mutate(bins=cut_interval(total,5)) # 20  88 275 10321 -> 0-100 ; 100-500; 500-1000;1000-5000 ;+5000 +1000
# 
# 
# 
# hist(puntos3$total)
# hist(puntos2$total)

```

```{r echo=FALSE,message=FALSE,warning=FALSE}
# puntos %>% ggplot() +
    

```


```{r echo=FALSE,message=FALSE,warning=FALSE}
# 
# haceme_mapa_anio <- function(data,arg_map,anio){
#     
#     datos_filtrados <- data %>%  filter(anio == anio)
#         
#     gg <- ggplot() 
#     
#     # necessary in the new world of ggplot2 mapping O_o
#     gg <- gg + geom_blank(data=arg_map, aes(long, lat))
#     
#     # draw the base polygon layer
#     gg <- gg + geom_map(data=arg_map, map=arg_map, 
#                         aes(fill="bins", map_id=id,frame = "anio"),
#                         color="#b2b2b2", size=0.15, fill=NA)
#     # fill in the polygons
#     animacion <- gg + geom_map(data=datos_filtrados, map=arg_map,
#                         aes(fill=bins, map_id=provincia,frame = anio),
#                         color="#b2b2b2", size=0.15)
#     
#     # animacion <- animacion + scale_fill_viridis(name="Scale Title")
#     animacion <- animacion + coord_proj(arg_proj)
#     animacion <- animacion + theme_map()
#     animacion <- animacion + theme(legend.position=c(0.8, 0.1))
#     animacion 
# }
# 
# anios_disponibles <- data_prov %>% count(anio) %>% pull(anio)
# listado_mapas <- purrr::map(anios_disponibles, ~ haceme_mapa_anio(data_prov,arg_map,anio=.) )
# 
# puntos2 %>% count(bins)
# 
#  haceme_mapa_anio(puntos2,arg_map,anio=2001)
#  haceme_mapa_anio(puntos3,arg_map,anio=2000)
# 
#  boxplot(puntos2$total)
#  skimr::skim(puntos2$total)
#  
# length(anios_disponibles)
# which(anios_disponibles==2001)
# listado_mapas[[9]]
# # data_prov %>%  filter(anio == 1993)
# # data_prov %>%  filter(anio == 2003)
# # data_prov %>%  filter(anio == 2013)
# 
# # haceme_mapa_anio(data_prov,arg_map = arg_map,anio = 2016)
# 
# animacion + gganimate::transition_manual(anio,frames = anio)
# 
# data_prov <-  area_processed %>% 
#     rename_all(gsub, pattern = '_area', replacement = '') 
# 
# range(data_prov$total)
# data_prov %>%  filter(total == 2913817)
# 
# 
# 
# data_prov_2 <- area_processed %>% 
#     rename_all(gsub, pattern = '_area', replacement = '') %>% 
#     left_join(arg_map,by=c("provincia"="id")) %>% 
#     mutate(anio=factor(anio))
```
```{r echo=FALSE,message=FALSE,warning=FALSE}

```

```{r echo=FALSE,message=FALSE,warning=FALSE}
# gif_final <- data_prov_2 %>% 
#     ggplot()+
#     geom_sf(aes(fill=total))+
#     # theme_void()+
#   theme(axis.title=element_blank(),
#         axis.text=element_blank(),
#         axis.ticks = element_blank(),
#         strip.text.y = element_text(size = 10),
#         panel.background=element_rect(fill="white"))+
#     scale_fill_gradient(low="blue", high="red",na.value = "black")+
#     gganimate::transition_manual(anio,frames = anio)+
#     labs(title="Total area burnt by province and year {current_frame}",
#          subtitle = paste0(""),
#          caption = "http://datos.ambiente.gob.ar/dataset/incendios-forestales",
#          x="",y="")
#      #%>% gganimate::save_animation(file = "tdor_yearly_2018.gif",)
# 
# # gganimate::anim_save(filename = "tdor_yearly_2018.gif")
# gif_final
```


```{r echo=FALSE,message=FALSE,warning=FALSE}

# %>% 
#     # mutate(provincia=fct_reorder(provincia,total)) %>% 
#     gather(key,value,-provincia,-anio) %>% 
#     filter(key!="total") %>% 
#     ggplot() +
#         geom_sf(aes(fill=deaths))+
#         geom_col()+
#         scale_y_continuous(labels = scales::number) +
#         labs(title="Total area burnt by province 1993-2016",
#              subtitle="remember it can be burnt again the same spot",
#              x="",
#              y="# hectareas",
#              fill="vegetation type")+
#         coord_flip()+
#         theme_light()

# data_prov_2 %>% ggplot() +
#     geom_blank(aes(long, lat)) +
#     geom_map( map=arg_map, 
#                     aes(map_id=provincia),
#                     color="#b2b2b2", size=0.15, fill=NA) +
#     geom_map(map=arg_map,
#                     aes(fill=total, map_id=provincia),
#                     color="#b2b2b2", size=0.15) +
#     scale_fill_viridis(name="Scale Title") +
#     coord_proj(arg_proj) +
#     theme_map() +
#     theme(legend.position=c(0.8, 0.1))+
#     gganimate::transition_manual(anio,frames = anio)+
#     labs(title="Total area burnt by province and year {current_frame}",
#          subtitle = paste0(""),
#          caption = "http://datos.ambiente.gob.ar/dataset/incendios-forestales",
#          x="",y="")
# #     
# 
# shape_file <- here::here("jas1_weeks","2018-12-25","provincias","provincias.shp")
# load_shape_file <- sf::read_sf(shape_file) %>% 
#      mutate(NAM=if_else(str_detect(NAM,"Tierra del Fuego"),"Tierra del Fuego",NAM)) 
#  
# 
#  area_processed %>% count(provincia) 
#   
# 
# con_geometry_2 <-load_shape_file %>% left_join(area_processed,by=c("NAM"="provincia")) %>% 
#     rename_all(gsub, pattern = '_area', replacement = '')
# 
# con_geometry <- arg_map %>% 
#     left_join(area_processed,by=c("id"="provincia")) %>% 
#     rename_all(gsub, pattern = '_area', replacement = '')
#     
# # object.size(con_geometry_2)/1024 /1024
# glimpse(con_geometry_2)
# gif_final <- con_geometry_2 %>% 
#     ggplot()+
#     geom_sf(aes(fill=total))+
#     # theme_void()+
#   theme(axis.title=element_blank(),
#         axis.text=element_blank(),
#         axis.ticks = element_blank(),
#         strip.text.y = element_text(size = 10),
#         panel.background=element_rect(fill="white"))+
#     scale_fill_gradient(low="blue", high="red",na.value = "black")+
#     gganimate::transition_manual(anio,frames = anio)+
#     labs(title="Total area burnt by province and year {current_frame}",
#          subtitle = paste0(""),
#          caption = "http://datos.ambiente.gob.ar/dataset/incendios-forestales",
#          x="",y="")
# library(gapminder)#; install.packages("gapminder")
# ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop, colour = country)) +
#   geom_point(alpha = 0.7, show.legend = FALSE) +
#   scale_colour_manual(values = country_colors) +
#   scale_size(range = c(2, 12)) +
#   scale_x_log10() +
#   facet_wrap(~continent) +
#   # Here comes the gganimate specific bits
#   labs(title = 'Year: {frame_time}', x = 'GDP per capita', y = 'life expectancy') +
#   transition_time(year) +
#   ease_aes('linear')
    
```

# comunicate.

well this time kind of a failure.

leson learnt: need more practice with custom map data
not the " world " or " usa" data; real map data related to my places.

still get complicated with maps some times.

need to practice more maps.

( refresh map courses & do practice far from the default maps ( world / usa) )