---
title: "Tidy tuesday challenge: Week 2019-04-02 Seattle bikes"
author: "julio"
date: "02-04-2019"
output: html_document
---

# Tidy tuesday challenge: Week 2019-04-03 Seattle bikes

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

### Data:

this week data its related to Seattle bikes
https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-04-02

### objectives:

- replicate @drob's faceted  day/trails plot
- may be something else


## details:


## import data
```{r echo=FALSE,message=FALSE,warning=FALSE}
library(dplyr)
library(tidyr)
library(skimr)
library(readr) 
library(ggplot2)
library(scales)
library(forcats)
library(lubridate)
library(gganimate)
library(animation)

```

```{r echo=FALSE,message=FALSE,warning=FALSE}
# bike_traffic_raw <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-04-02/bike_traffic.csv")
# getting local as its large file.
bike_traffic_raw <- readr::read_csv(here::here("data","2019","2019-04-02","bike_traffic.csv"))

# utils::head(bike_traffic_raw)
# dplyr::glimpse(bike_traffic_raw)
# skimr::skim(bike_traffic_raw)

# to check drob-s assumption on dates> he-s rigth about format.
# bike_traffic_raw %>% slice(5000:7000)
  
bike_traffic <- bike_traffic_raw %>% 
    mutate(date = mdy_hms(date)) %>% 
    select(-ped_count) # too much missings
```

# drob's plot weekday/hour

```{r}
#credits to drob: 
# https://github.com/dgrtwo/data-screencasts/blob/master/bike_traffic.Rmd

# week day start
# https://lubridate.tidyverse.org/reference/day.html

Sys.setlocale("LC_TIME", "English")
bike_traffic %>%
  group_by(crossing,
           weekday = wday(date, label = TRUE,
                          week_start = getOption("lubridate.week.start", 1)),
           hour = hour(date)) %>%
  summarize(total_bikes = sum(bike_count, na.rm = TRUE)) %>%
  group_by(crossing) %>%
  mutate(pct_bike = total_bikes / sum(total_bikes)) %>% # aesome 
  ggplot(aes(hour, pct_bike, color = crossing)) +
  geom_line(show.legend = FALSE) +
  facet_grid(crossing ~ weekday,labeller = label_wrap_gen(width = 16)) +
    theme(strip.text.y = element_text(angle = 360))+
  scale_y_continuous(labels = percent_format()) +
  labs(x = "Time of week",
       y = "% of bike crossings happening in this hour",
       title = "When in the week do people in Seattle bike?",
       subtitle = "Based on crossings from 2014-February 2019",
       caption = "tidytuesday 2019-04-02")
```

# calendar heat map 

always its really awsome to get a calendar heat map to get similar stuff.
it gives you other perspective of same data


```{r}

#bike_traffic
#date , hourly, by crossing, direction and bike count.
# i'll agregate by totalbikes by date.

total_by_hour <- bike_traffic %>% group_by(date) %>% 
    summarise(total_bikes=sum(bike_count))

total_by_date <- bike_traffic %>%
    mutate(day=lubridate::as_date(date)) %>% 
    mutate(day_of_year=lubridate::yday(date)) %>% 
    mutate(week=lubridate::week(date)) %>% 
    mutate(year=lubridate::year(date)) %>% 
    mutate(month=lubridate::month(date)) %>% 
    mutate(year_month=paste0(year,"-",month)) %>% 
    filter(year > 2013, year < 2019) %>% 
    group_by(year_month,year,month,week,day,day_of_year) %>% 
    summarise(total_bikes=sum(bike_count))

# range(total_by_date$week)
# range(total_by_date$day_of_year)

plot_month <- total_by_date %>% 
    ggplot(aes(x=month,y=year,fill = total_bikes)) +
    geom_tile() +
    scale_x_continuous(breaks = c(1:12))+
    scale_fill_viridis_c()+
    theme_light()+
    labs(title="Total monthly bike rides",
         subtitle = "from 2014 to 2018",
         x= "Month",y="",fill="bikes",
         caption = "tidytuesday 2019-04-02")

plot_week <- total_by_date %>% 
    ggplot(aes(x=week,y=year,fill = total_bikes)) +
    geom_tile() +
    scale_x_continuous(breaks = c(1,seq(from=10, to=53,by =10),53))+
    scale_fill_viridis_c()+
    theme_light()+
    labs(title="Total weekly bike rides",
         subtitle = "from 2014 to 2018",
         x= "Week",y="",fill="bikes",
         caption = "tidytuesday 2019-04-02")

plot_day_year <- total_by_date %>% 
    ggplot(aes(x=day_of_year,y=year,fill = total_bikes)) +
    geom_tile() +
    scale_x_continuous(breaks = c(1,seq(from=50, to=350,by =50),366))+
    scale_fill_viridis_c()+
    theme_light()+
    labs(title="Total daily bike rides",
         subtitle = "from 2014 to 2018",
         x= "Day of the year",y="",fill="bikes",
         caption = "tidytuesday 2019-04-02")


# to plot stuff from different plots. and merge em in a gif.
animation::saveGIF({
    
  plots <- list(plot_month,plot_week,plot_day_year)
    
    
    
  for (i in 1:length(plots)){  
      print(plots[i])
  }
}, interval = 1, movie.name="drill_down.gif")


```

# communicate: 

- this week got inspired by drob's podcast
- wanted to do a calendar heatmap , long time ago used: rChartsCalmap, discarded de idea
- mutated to do a month / week / day of year plot 
- added the drilldown effect with a gif
- may be cool stuff will be to add functionality to: rChartsCalmap
- other idea discarded by time constraints: using network animations on: 


## some resources: 

- seeking for calendar heatmaps: https://towardsdatascience.com/time-series-calendar-heatmaps-9f576578fcfe
- an amazing poster like: https://github.com/othomantegazza/code-tidytuesday/blob/master/2-14-seattle-bike-draft-grid.R

- rotate labels on facet grid: https://stackoverflow.com/questions/40484090/rotate-switched-facet-labels-in-ggplot2-facet-grid

- change localtime : to see days in english: 
https://stackoverflow.com/questions/8145886/change-time-locale-for-r

- combining different plots in a gif: 
https://stackoverflow.com/questions/32442276/combine-multiple-plots-to-a-gif

- drobs podcast: https://www.youtube.com/watch?v=sBho2GJE5lc




# not used stuff
```{r}
# https://towardsdatascience.com/time-series-calendar-heatmaps-9f576578fcfe

# total_by_date %>% 
#     ggplot(aes(monthweek, weekdayf, fill = amznStock$adjusted)) + 
#     geom_tile(colour = "white") + 
#     facet_grid(year(amznStock$date)~monthf) + scale_fill_gradient(low="red", high="green") + xlab("Week of Month") + ylab("") + ggtitle("Time-Series Calendar Heatmap: AMZN Stock Prices") + labs(fill = "Price")


# count(total_by_date)
# count(total_by_hour)

# hourly, date

library(devtools)
# install.packages('htmlwidgets')
install_github("Tutuchan/rChartsCalmap")
library(rChartsCalmap)

range(total_by_date$total_bikes,na.rm = TRUE)
hist(total_by_date$total_bikes)
# 24029 / 7 
# seq(by = 3500, from = 0, to = 25000)


# r1 <- 
calheatmap(x = 'day', y = 'total_bikes',
  data = total_by_date %>% filter(year==2014), 
  domain = 'month',
  start = min(total_by_date$day),
  legend = c(0,1000,3000,5000,10000,25000),
  # legend = seq(by = 3500, from = 0, to = 25000),
  itemName = 'bikes'#,
  # range = 7
)
calheatmap(x = 'day', y = 'total_bikes',
  data = total_by_date %>% filter(year==2015), 
  domain = 'month',
  start = min(total_by_date$day),
  legend = c(0,1000,3000,5000,10000,25000),
  # legend = seq(by = 3500, from = 0, to = 25000),
  itemName = 'bikes'#,
  # range = 7
)
calheatmap(x = 'day', y = 'total_bikes',
  data = total_by_date, 
  domain = 'month',
  start = min(total_by_date$day),
  legend = c(0,1000,3000,5000,10000,25000),
  # legend = seq(by = 3500, from = 0, to = 25000),
  itemName = 'bikes',#,
  # range = 7
  itemSelector= "#previousSelector-a",
  previousSelector= "#previousSelector-a-previous",
  nextSelector= "#previousSelector-a-next"
)
# r1
```

