---
title: "week 2018-12-18 challenge"
author: "_jas"
date: 20181218
output: html_document
---

# Tidy tuesday challenge: Week 2018-12-18

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

Captive Cetacean Dataset

## import data


```{r echo=FALSE,message=FALSE,warning=FALSE}
library(here)
library(readr)
library(purrr)
library(dplyr)
library(tidyr)
library(janitor)#install.packages('janitor')
library(scales)
library(ggplot2)
library(forcats)#install.packages('forcats')
library(lubridate)
library(visdat)#install.packages('visdat')
library(stringr)
library(DataExplorer)
library(ggrepel)#install.packages('ggrepel')
library(rebus)
library(igraph)
library(ggraph)
library(tidygraph)#install.packages('tidygraph')
library(gganimate)


# library(rpart)
# library(rpart.plot)
# library(cluster)
# library(leaflet)
```


```{r echo=FALSE,message=FALSE,warning=FALSE}
file_name <- here::here("data","2018-12-18","allCetaceanData.csv")
cetacean_raw <- readr::read_csv(file_name)
glimpse(cetacean_raw)
```


# lets check missing data

```{r echo=FALSE,message=FALSE,warning=FALSE}
DataExplorer::plot_missing(cetacean_raw)
```
X1, its a rowid 
transfer date like 99% of cases dont have transfer date, 
but have other interesting columns as transfers & transfers

a question might be " its common to transfer cetaceans ?"

## is it common to transfer cetaceans ?

```{r echo=FALSE,message=FALSE,warning=FALSE}
transfers <- cetacean_raw %>% select(id,transfer,transfers,transferDate)


# US centric & not detailed, at least no NA's
# transfers %>% count(transfer,sort=TRUE)

# half of data ~= NA : 1392
# transfers %>% count(transfers,sort=TRUE)


# transfers %>% count(transfers,sort=TRUE) %>% 
#     mutate(str_longest=str_length(transfers)) %>% 
#     arrange(desc(str_longest)) %>% 
#     pull(transfers)

count_transfers <- transfers %>% 
    mutate(amount_transfers=str_count(transfers," to ")) %>% 
    arrange(desc(amount_transfers))

count_transfers %>% 
    count(amount_transfers) %>% 
    mutate(amount_transfers=if_else(is.na(amount_transfers),as.integer(0),amount_transfers)) %>% 
    mutate(amount_transfers=factor(amount_transfers)) %>% 
    mutate(amount_transfers=fct_reorder(amount_transfers,n)) %>% 
    ggplot(aes(x=amount_transfers,y=n,fill=amount_transfers)) + 
    geom_col()+
    coord_flip()+
    labs(title="Amount of transfers by Cetacean",
         x="",
         y="# cetaceans with N transfers")+
    theme_light()+
    theme(legend.position = "none")
```


## Cetacean destination graph

as its a path, its a directed graph
but to get that i need all the " from / to "
splited by " to " in the transfer date
by cetacean


### prepare graph data:

```{r echo=FALSE,message=FALSE,warning=FALSE}
# split transfers destinations

transfers_rows <- cetacean_raw %>% 
    rename(row_id=X1) %>% 
    select(row_id,transfers)

transfers_from_to <- transfers_rows %>% 
    separate_rows(transfers, sep = " to ", convert = FALSE) %>% # separate destinations of registry
    separate(transfers,c("destiny","date_aprox"),sep=rebus::OPEN_PAREN) %>%
    mutate(destiny=str_trim(destiny)) %>% 
    mutate(date_aprox=str_replace(date_aprox,rebus::CLOSE_PAREN,"")) %>% 
    mutate(from=if_else(row_id==lag(row_id),lag(destiny),"")) %>% 
    select(row_id,from,destiny) %>% 
    filter(!is.na(destiny),from!="") %>% 
    rename(to=destiny)

# transfers_from_to %>% filter(from==" ")
transfers_from_to_graph_df_weighted <- transfers_from_to %>%  count(from,to) %>%  rename(weight=n)


# transfers_from_to %>% filter(row_id==4)
# transfers_rows %>% filter(row_id==4)
# transfers_rows %>% filter(row_id==3) %>% pull(transfers) %>% str_count(" to ") 
# transfers_rows %>% filter(row_id==5) %>% pull(transfers) %>% str_count(" to ") 
# transfers_from_to %>% filter(row_id==5)
```


### make the graph

```{r echo=FALSE,message=FALSE,warning=FALSE}

transfers_graph <- igraph::graph_from_data_frame(transfers_from_to_graph_df_weighted) %>% as_tbl_graph() 

igraph::V(transfers_graph)$degree <- igraph::degree(transfers_graph)
igraph::V(transfers_graph)$between <- igraph::betweenness(transfers_graph)

# igraph::average.path.length(transfers_graph)
# igraph::transitivity(transfers_graph)
# igraph::diameter(transfers_graph)
filtered_graph <- transfers_graph %>% 
    mutate(community = group_walktrap()) %>% 
    filter(community %in% 1:3)

# is_directed(transfers_graph)
# filtered_graph <- transfers_graph %>% 
    

# V(transfers_graph)
filtered_graph %>% 
    ggraph(layout = 'mds') +
    geom_node_point(aes(size = degree,col = as.factor(community))) +
    ggraph::geom_edge_link(aes(edge_alpha = weight))+
    theme_void()+ 
    theme(legend.position = 'none') +
    labs(title="Top 3 Cetacean Transfer Communities",
         subtitle = "Node size: degree; Link alpha: weight; Communities: walktrap; Layout: mds")

filtered_graph %>%
    ggraph(layout = 'lgl') +
    ggraph::geom_node_point(aes(size = degree,col = as.factor(community))) +
    ggraph::geom_edge_link(aes(edge_alpha = weight))+
    theme_void()+
        theme(legend.position = 'none') +
    labs(title="Top 3 Cetacean Transfer Communities: walktrap",
         subtitle = "Node size: degree; Link alpha: weight")
    # ggraph::geom_node_label(filter=V(transfers_graph)$degree>40,aes(label=name)) 
```


# animation, comunities, layouts

```{r echo=FALSE,message=FALSE,warning=FALSE}



# https://perrystephenson.me/2018/09/29/the-r-twitter-network/

layout_list <- list(
  list(layout = 'star'),
  list(layout = 'circle'),
  list(layout = 'gem'),
  list(layout = 'graphopt'),
  list(layout = 'grid'),
  list(layout = 'mds'),
  list(layout = 'randomly'),
  list(layout = 'fr'),
  list(layout = 'kk'),
  list(layout = 'nicely'),
  list(layout = 'lgl'),
  list(layout = 'drl'))




layouts <- filtered_graph %>% 
  invoke_map('create_layout', layout_list, graph = .) %>% 
  set_names(unlist(layout_list)) %>% 
  bind_rows(.id = 'layout')

dummy_layout <- create_layout(filtered_graph, 'nicely')

attr(layouts, 'graph') <- attr(dummy_layout, 'graph')
attr(layouts, 'circular') <- FALSE


animated_graph <- ggraph(layouts) +
  geom_node_point(aes(size = degree,col = as.factor(community))) +
    # geom_edge_link(aes(edge_alpha = weight))+ # not working properly
  theme_void() +
  theme(legend.position = 'none') +
  labs(title = 'Cetacean Transfer Graph; Node Size: degree ; Color: top 3 communities',
       subtitle = 'Using {closest_state} layout') +
  transition_states(layout, 1, 2) +
  ease_aes('linear') +
  view_follow()

# gganimate::anim_save(filename = "cetacean_2018.gif")

# animate(animated_graph, fps = 10, nframes = 100)


```


# Some graph statistics

```{r echo=FALSE,message=FALSE,warning=FALSE}

# [V(transfers_graph)$degree > 25]
# [V(transfers_graph)$degree > 25]
plot_labels_df <- data.frame(name=V(transfers_graph)$name,
           degree=as.integer(V(transfers_graph)$degree)) %>%  as_tibble() %>% 
    group_by(degree) %>% 
    summarize(label=paste0(name,collapse="; "))


# threshold_degree <- 25

# plot_labels_df %>% filter(degree>threshold_degree) 



# transfers_graph

degree_distribution_plot_data <- transfers_graph %>% 
    degree_distribution() %>% 
            as_tibble() %>% 
            rename(degree_distribution=value) %>% 
            tibble::rownames_to_column("degree") %>% 
            mutate(degree=as.integer(degree)) %>% 
            arrange(degree) %>% 
    left_join(plot_labels_df,by=("degree"="degree"))

glimpse(degree_distribution_plot_data)
glimpse(plot_labels_df)


degree_distribution_plot_data %>% 
    ggplot(aes(x=degree,y=degree_distribution,label=label)) + 
    geom_col() +
    labs(title="Transfers Graph: Degree distribution",
         x="Degree",
         y="P degree")+
    theme_light()+
    ggrepel::geom_text_repel(data=subset(degree_distribution_plot_data,
                                        degree > 25 & degree < 40 ),
                            nudge_y      = 0.05,
                            direction    = "x",
                            angle        = 90,
                            vjust        = 0,
                            segment.size = 0.2,
                             segment.color="blue",
                             color="blue") +
    ggrepel::geom_text_repel(data=subset(degree_distribution_plot_data,
                                        degree > 40 ),
                            nudge_y      = 0.05,
                            direction    = "x",
                            angle        = 90,
                            vjust        = 0,
                            segment.size = 0.2,
                             segment.color="red",
                             color="red")

```


```{r echo=FALSE,message=FALSE,warning=FALSE}


#     
# grrr i can still be doing this for hours and add new stuff all the time ...
# i rather close on those.
# btw <- transfers_graph %>% betweenness()
# 
# 
# btw_histo_plot_data <- data.frame(names=names(btw),value=btw) %>% 
#             as_tibble() %>% 
#             rename(betweenness=value)
# 
# 
# btw_histo_plot_data %>% 
#             ggplot(aes(betweenness)) +
#             geom_histogram() +
#             labs(title="Betweenness Histogram ",
#                  x= "Betweenness",
#                  y= "# nodes")+
#             theme_light()+
#         ggrepel::geom_text_repel(data=subset(btw_histo_plot_data,
#                                         betweenness > 4800 & betweenness < 7500 ),
#                             nudge_y      = 0.05,
#                             direction    = "x",
#                             angle        = 90,
#                             vjust        = 0,
#                             segment.size = 0.2,
#                              segment.color="blue",
#                              color="blue")

    
```


# lets check missing data

```{r echo=FALSE,message=FALSE,warning=FALSE}
# DataExplorer::plot_missing(cetacean_raw)
```


# are 1 registry per cetacean?

search with more than 1 of same id:

```{r echo=FALSE,message=FALSE,warning=FALSE}
# more_than_1_registry <- cetacean_raw %>% count(id,sort=TRUE) %>%  filter(n>1) 
# 
# more_than_1_registry 
# 
# total with more than 1 registry: `r more_than_1_registry %>% count()`
# total registries: `r nrow(cetacean_raw)`
# proportion: `r more_than_1_registry / nrow(cetacean_raw)`
```




# communicate

- this week was another really late tuesday :p 
- was fun that got the idea how to work with graphs here . 
- will be awsome to write some special case explorer.
- database was not clean, and i did not expend really much time on that
- for duplications i rather prefer open refine & text clustering; as i dont know & was a short try & also that task consumes a lot of time, it was not done.
- focused on transforming the transfers to from/to for each rowid , and then, summarized,to get weights on edges.




