
---
title: "Tidy tuesday challenge: Week 2019-07-30 Video Games Dataset"
author: "julio"
date: "2019-08-04"
output: html_document
---

# Tidy tuesday challenge: Week  2019-07-30 Video Games Dataset

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

### Data:
 
2019-07-30 Video Games Dataset
 
### objectives:

check the data and do something

## details:


## import data
```{r echo=FALSE,message=FALSE,warning=FALSE}
library(magrittr) # para el %T>%
library(tidyverse)
# library(sf)
library(dplyr)
library(stringr)#;
# library(rebus)#; install.packages('rebus')
# library(tidytext)
library(prophet)


# install.packages("Rcpp")
# remotes::install_github("tylermorganwall/rayshader")
# library(rayshader)
library(lubridate)
library(ggforce)
library(ggrepel)

# library

```


```{r echo=FALSE,message=FALSE,warning=FALSE}
video_games <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-30/video_games.csv")
```

# explore data


```{r echo=FALSE,message=FALSE,warning=FALSE}
video_games %>% head()

```

```{r echo=FALSE,message=FALSE,warning=FALSE}
video_games %>% glimpse()

```

```{r echo=FALSE,message=FALSE,warning=FALSE}
video_games %>% skimr::skim()


```


# longest living popular games in steam ! 

this have to be crossed to game real release date.
example quake @ 1996, not 2007 , but steam started 2003~2004
also steam is owned by Valve , creators of half life, early games was more related to them.


```{r echo=FALSE,message=FALSE,warning=FALSE}
video_games %>% filter(!is.na(owners) & !is.na(release_date)) %>% 
    mutate(release_date=lubridate::mdy(release_date)) %>% 
    filter(str_detect(str_to_lower(game),"quake"))

```



```{r echo=FALSE,message=FALSE,warning=FALSE}
games_owners <- video_games %>% filter(!is.na(owners) & !is.na(release_date)) %>% 
    mutate(release_date=lubridate::mdy(release_date)) %>% 
    count(owners) %>% 
        mutate(owners=fct_reorder(owners,n)) %>% 
        ggplot(aes(x=owners,y=log(n),fill=owners))+
        geom_col()+
        coord_flip()+
    labs(title="Amount of games by quantity of owners, Log scale", 
         x="#owners",y="Log(#games)", 
         caption = "#TidyTuesday")
    # filter(str_detect(str_to_lower(game),"quake"))

ggsave(games_owners,filename = "games_owners.png")

```


```{r echo=FALSE,message=FALSE,warning=FALSE}
explore_boxplot <- video_games %>% filter(!is.na(price) & !is.na(release_date)) %>% 
    mutate(release_date=lubridate::mdy(release_date)) %>% 
    mutate(year_release=lubridate::year(release_date)) %>% 
        ggplot(aes(x=as.character(year_release),y=price,fill=as.character(year_release)))+
        geom_boxplot()+
        coord_flip()+
    labs(title="Games price distribution by year", 
         x="",y="USD",fill="year", 
         caption = "#TidyTuesday")
explore_boxplot
# plotly::ggplotly(explore_boxplot)

```



```{r echo=FALSE,message=FALSE,warning=FALSE}

owner_levels_vec <- c("100,000,000 .. 200,000,000",
"50,000,000 .. 100,000,000",
"20,000,000 .. 50,000,000",
"10,000,000 .. 20,000,000",
"5,000,000 .. 10,000,000",
"2,000,000 .. 5,000,000",
"1,000,000 .. 2,000,000",
"500,000 .. 1,000,000",
"200,000 .. 500,000",
"100,000 .. 200,000",
"50,000 .. 100,000",
"20,000 .. 50,000",
"0 .. 20,000")
owner_levels <- data.frame(owners=owner_levels_vec,owners_order=seq(1,length(owner_levels_vec),1),stringsAsFactors = FALSE)
relevel_owners <- video_games %>% 
    left_join(owner_levels,by = c("owners")) %>% 
    mutate(label_plot=if_else(is.na(price),paste0("[NA]:",game),game)) %>% 
    mutate(price=if_else(is.na(price),0.001,price)) %>% 
    mutate(release_date=lubridate::mdy(release_date)) %>% 
    mutate(year_release=lubridate::year(release_date))


relevel_owners_plot <- relevel_owners %>%
    
    mutate(owners=fct_reorder(owners,owners_order)) %>% 
    ggplot(aes(x=log(median_playtime),
                   y=log(price),
                   color=owners,
               label=label_plot))+
        geom_point()+
    labs(title="Games price vs median playtime (log / log)",
         subtitle="better to explore interactively with plotly\nNA price under-4 log(price)",
    x="log(median playtime) (minutes)",y="log(price)(usd)" ,colour="# of owners",
    caption = "#TidyTuesday")

plotly::ggplotly(relevel_owners_plot)

ggsave(relevel_owners_plot,filename = "relevel_owners_plot.png")

```


#tweet: 

2019-07-30 #TidyTuesday #rstats Steam Video Games Dataset ! 
just explored a little price vs time spent in game.
Better to check it interactively with plotly.
can see NA price games, had a wide range of playtime & popularity.



# communicate

just checked spent money vs spent time.







