---
title: "week 2019-01-29 #rstats #tidytuesday diary"
author: "_jas"
date: 20190129
output: html_document
---

# Tidy tuesday challenge: Week 2019-01-29 diary

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

### Data:

this week data its related to diary in USA.

### objectives:

try prophet again

- do a TS forecast on some subset of the data.
- do some cheese colors on ggplot

## import data

```{r echo=FALSE,message=FALSE,warning=FALSE}
library(here)
library(readr)
library(dplyr)
library(tidyr)
library(scales)
library(ggplot2)
library(forcats)#install.packages('forcats')
library(lubridate)
library(visdat)#install.packages('visdat')
library(stringr)
library(naniar)
library(rpart)
library(rpart.plot)
library(cowplot)
library(rstan) # install.packages("rstan")
library(prophet)# install.packages("prophet")
library(ggrepel)
```



```{r echo=FALSE,message=FALSE,warning=FALSE}
milkcow_facts_file <- here::here("data","2019","2019-01-29","milkcow_facts.csv")
milkcow_facts_raw <- readr::read_csv(milkcow_facts_file)

milkcow_facts_processed <- milkcow_facts_raw %>% 
    mutate(decade = floor(year / 10)*10 ) %>% 
    naniar::add_any_miss()

dplyr::glimpse(milkcow_facts_processed)
skimr::skim(milkcow_facts_processed)
```



```{r echo=FALSE,message=FALSE,warning=FALSE}

clean_cheese_file <- here::here("data","2019","2019-01-29","clean_cheese.csv")
clean_cheese_raw <- readr::read_csv(clean_cheese_file)

clean_cheese_processed <- clean_cheese_raw %>% janitor::clean_names()

dplyr::glimpse(clean_cheese_processed)
skimr::skim(clean_cheese_processed)
# m <- rpart(outcome ~ loan_amount + credit_score, data = loans, method = "class")
```


```{r echo=FALSE,message=FALSE,warning=FALSE}

fluid_milk_sales_file <- here::here("data","2019","2019-01-29","fluid_milk_sales.csv")
fluid_milk_sales_raw <- readr::read_csv(fluid_milk_sales_file)

fluid_milk_sales_processed <- fluid_milk_sales_raw %>% janitor::clean_names()

dplyr::glimpse(fluid_milk_sales_processed)
skimr::skim(fluid_milk_sales_processed)


data_for_ts <- fluid_milk_sales_processed %>% 
    mutate(ds=ymd(paste(year,'12','31'))) %>% # for prophet
    rename(y=pounds) %>% 
    select(milk_type,ds,y) %>% 
    filter(milk_type != 'Total Production')

# fluid_milk_sales_processed %>% pull(year) %>%  range()

# 
# data_for_ts %>% count(milk_type)%>% pull(milk_type)
# [1]                             
# [4]                      
# [7]                   
milk_colors <-  c("Whole"="#fff7d7",
                  "Skim" = "#f0e7e0",
                  "Reduced Fat (2%)"="#dbd6ab",
                  "Low Fat (1%)" = "#d8c49c",
                  "Flavored (Whole)"="#cca472",
                  # "Flavored  (Not Whole)"="#dfb593",
                  "Flavored  (Not Whole)"="#ecddcd",
                  
                  "Eggnog"="#d4a17f",
                  "Buttermilk"="#cc615b")
data_for_ts %>% 
    ggplot(aes(y=y/1000000,x=ds,fill=milk_type))+
    geom_col() +
    scale_fill_manual(values = milk_colors)+
    # scale_y_continuous(labels = scales::number_format)+
    labs(title="Evolution of milk type sales in USA (1975-2017)",
         subtitle="in Millions pounds",
         caption = "#rstats # tidytuesday week 2019-01-29",
         fill="Milk Type",
         x="", y="")
    
    

# 
# make_forecast <- function(data_forecast,filter_var="rural"){
#     prophet_total <- data_forecast %>% 
#         filter(pop_category=="Total") %>% 
#         filter(urbanicity==param_urbanicity) %>%
#         ungroup() %>% 
#         select(ds,y)
#     }



# m <- rpart(outcome ~ loan_amount + credit_score, data = loans, method = "class")
```

# communicate

time counting ... so got to adjust schedule. 

so gone only for the colors =) 

## some usefull data in: 

https://ggplot2.tidyverse.org/reference/scale_manual.html

also take care on scale_fill_manual, or scale_colour_manual . 


## some  colours inspirations on: 

https://www.color-hex.com/color-palette/70954

https://www.color-hex.com/color-palettes/?keyword=milk

https://www.color-hex.com/color-palette/14317

https://color.adobe.com/nuvole-color-theme-11975068/
