
---
title: "Tidy tuesday challenge: Week 2019-07-09 Womens World Cup"
author: "julio"
date: "2019-06-25"
output: html_document
---

# Tidy tuesday challenge: Week 2019-07-09 Womens World Cup

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

### Data:
 
 Womens World Cup
 
### objectives:

check the data and do something.

## details:


## import data
```{r echo=FALSE,message=FALSE,warning=FALSE}
library(magrittr) # para el %T>%
library(tidyverse)
# library(sf)
library(dplyr)
library(stringr)#;
library(rebus)#; install.packages('rebus')
library(tidytext)

# install.packages("Rcpp")
# remotes::install_github("tylermorganwall/rayshader")
library(rayshader)
library(lubridate)


```


```{r echo=FALSE,message=FALSE,warning=FALSE}
wwc_outcomes <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-09/wwc_outcomes.csv")
squads <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-09/squads.csv")
codes <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-09/codes.csv")
```

# explore data


```{r echo=FALSE,message=FALSE,warning=FALSE}
wwc_outcomes %>% head()
squads %>% head()
codes %>% head()

```

```{r echo=FALSE,message=FALSE,warning=FALSE}
glimpse(wwc_outcomes)
glimpse(squads)
glimpse(codes)
```

```{r echo=FALSE,message=FALSE,warning=FALSE}
wwc_outcomes %>% skimr::skim()
squads %>% skimr::skim()
codes %>% skimr::skim()

```


# how long ago wwc exists ?
```{r echo=FALSE,message=FALSE,warning=FALSE}
wwc_outcomes %>% 
    count(year) %>%
    mutate(year_date=lubridate::ymd(paste0(year,"-01-15"))) %>% 
    ggplot(aes(x=year_date,y=n,color=n))+
    geom_line()+
    geom_point()+
    scale_x_date(date_labels = "%Y",date_breaks = "4 years",
                 limits = c(lubridate::ymd("1989-01-15"),lubridate::ymd("2020-01-15")))+
    labs(title="# of Countries per WWC edition",
         x="",y="",color="# countries",
         caption="#TidyTuesday")

```

# how long ago wwc exists ?
```{r echo=FALSE,message=FALSE,warning=FALSE}

ww_plot <-  wwc_outcomes %>% 
    count(year) %>%
    mutate(year_date=lubridate::ymd(paste0(year,"-01-15"))) %>% 
    ggplot(aes(x=year_date,y=n,color=n,label=n))+
    geom_line()+
    geom_point()+
    ggrepel::geom_label_repel()+
    scale_x_date(date_labels = "%Y",date_breaks = "4 years",
                 limits = c(lubridate::ymd("1989-01-15"),lubridate::ymd("2020-01-15")))+
    labs(title="# of Countries per WWC edition",
         subtitle="Womens World Cup",

         x="",y="",color="# countries",
         caption="#TidyTuesday")+
    theme(legend.position = "none")

ggsave(filename = "participants.png",plot = ww_plot)

```

# top age distribution by country in 2019

```{r echo=FALSE,message=FALSE,warning=FALSE}

# order_sqd <- squads %>% 
    # group_by(country) %>% 
    # summarise(median_age_country=median(age),
    #           max_age_country=max(age),
    #           min_age_country=min(age),
    #           iqr_age_country=IQR(age),
    #           var_age_country=var(age))

is_outlier <- function(x) {
  return(x < quantile(x, 0.25) - 1.5 * IQR(x) | x > quantile(x, 0.75) + 1.5 * IQR(x))
}

sqd_out_2 <- squads %>% 
    nest(-country) %>% 
    mutate(team_data=purrr::map(data,.f=function(data){
        ret <- data %>%  mutate(is_outlier=is_outlier(age))
        ret
    })) %>% 
    unnest(team_data)
   
# 
# sqd_out <- squads %>% 
#     mutate(is_outlier=is_outlier(age)) 

team_var <- sqd_out_2 %>% 
    ggplot(aes(x=fct_reorder(country,age),y=age,
               fill=country))+
    geom_boxplot()+
    ggrepel::geom_label_repel(
        data          = subset(sqd_out_2, is_outlier),
        aes(label=paste0(player,"(",age,")")),
        # hjust=-0.3,
        # nudge_y = 0.3,
        segment.size = 0.2
        # nudge_x      = 0.1,
        # nudge_x      = -0.35,

    )+
    coord_flip()+
    theme(legend.position = "none")+
    labs(title = "Team Age Variance by Country",
         subtitle="Womens World Cup 2019 rosters",
         caption="#TidyTuesday",
         x="",y="age")
        
ggsave(filename = "team_age_variance.png",plot = team_var)

```


#tweet: 

2019-07-09 #TidyTuesday #rstats Womens World Cup! 
Amount of teams by cup & age variance per team.


# communicate

well almost there, some errors on the years elapsed label :/.

credit of the bar to: 

https://ogustavo.com/post/dotplot-ggplot/





