---
title: "Week 20181016"
author: "_jas"
date: 20181016
output: html_document
---

# Tidy tuesday challenge: Week 2018-10-16: 538 - Economic Guide to Picking a College Major

## keep it simple:

### Objectives: 

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate

https://rviews.rstudio.com/2017/06/08/what-is-the-tidyverse/ 

### this week expected

well its kind of ... friday ... again :p ... 
i generally dont see tidytuesday responses untill i do mine. 

this week got an interesting exception, @drob done a screencast on the challenge.

https://www.youtube.com/watch?v=nx5yhXAQLxw

after done this ... gone to drob github: 
https://github.com/dgrtwo/data-screencasts/blob/master/college-majors.Rmd



it was awsome to see how a person who deal with data everyday works with new data w/o knowing whats about it.

a lot of little workflow thingies arise during the workflow so,
as im short of time, i'll write the ones that didnt knew, so i got some new stuff on my toolbelt.

in the following i mix some stuff i learnt previously + some screencast stuff that i loved.


# Workflow:

## Import

```{r echo=FALSE,message=FALSE,warning=FALSE}
library(here)
library(readr)
library(tibble)
library(dplyr)
library(tidyr)
library(skimr)#install.packages('skimr')
library(scales)#install.packages('scales')
library(ggplot2)
library(forcats)#install.packages('forcats')
library(plotly)#install.packages('plotly')


```

```{r echo=FALSE,message=FALSE,warning=FALSE}
recent_grads <- readr::read_csv(here::here("data","2018-10-16","recent-grads.csv"))
```

## Tidy

```{r echo=FALSE,message=FALSE,warning=FALSE}
glimpse(recent_grads)
recent_grads <- recent_grads %>% janitor::clean_names()
skimr::skim(recent_grads)
```

there are some NA
```{r echo=FALSE,message=FALSE,warning=FALSE}
recent_grads %>% filter(!complete.cases(recent_grads)) %>% glimpse()

```


## Understand ( iteration 1 )

### Transform

removing NA's: just for the sake of practice, but in real case should evaluate WHY it is NA.

```{r echo=FALSE,message=FALSE,warning=FALSE}
recent_grads_no_NA <- recent_grads %>% filter(complete.cases(recent_grads))

recent_grads_no_NA
```




```{r echo=FALSE,message=FALSE,warning=FALSE}

```


### Visualize

Variability by major category:

```{r echo=FALSE,message=FALSE,warning=FALSE}
recent_grads_no_NA %>% 
    mutate(major_category=fct_reorder(major_category,median)) %>% 
    ggplot(aes(major_category,median,fill=major_category)) +
    geom_boxplot() +
    scale_y_continuous(labels = dollar_format())+
    labs(title="Variability by major category",
         subtitle="",
         x="",
         y="Median")+
    coord_flip() +  # rotate the boxplot as there is not clear 
    theme_light() + # theme light
    theme(legend.position = "none") # no legend just colors for categories.

```

total of graduates

```{r echo=FALSE,message=FALSE,warning=FALSE}
recent_grads_no_NA %>% 
    count(major_category,wt=total,sort=TRUE) %>% 
    mutate(major_category=fct_reorder(major_category,n)) %>%
    ggplot(aes(major_category,n,fill=major_category)) +
    geom_col() +
    scale_y_continuous(labels = comma_format())+
    labs(title="Major categories more common",
         subtitle="",
         x="",
         y="Total # Graduates")+
    coord_flip() +  # rotate the boxplot as there is not clear 
    theme_light() + # theme light
    theme(legend.position = "none") # no legend just colors for categories.

```

majors, categories , amount of graduates

```{r echo=FALSE,message=FALSE,warning=FALSE}
recent_grads_no_NA %>% 
    mutate(major=fct_reorder(major,total)) %>%
    arrange(desc(total)) %>% 
    head(20) %>% 
    ggplot(aes(major,total,fill=major_category)) +
    geom_col() +
    scale_y_continuous(labels = comma_format())+
    labs(title="Major categories more common",
         subtitle="",
         x="",
         y="Total # Graduates")+
    coord_flip() +  # rotate the boxplot as there is not clear 
    theme_light() + # theme light
    theme(legend.position = "none") # no legend just colors for categories.

```


sample size vs median

because sample size if small it should no be taken in account.
this visualization helps getting the idea where to make a cutoff.in case needed.

```{r echo=FALSE,message=FALSE,warning=FALSE}
recent_grads_no_NA %>%
    ggplot(aes(sample_size,median)) +
    geom_point()+
    # geom_text(aes(label=major),check_overlap = TRUE,vjust=1,hjust=1)+ # nice but much
    scale_x_log10()
    
```


majors, categories , amount of graduates by gender

```{r echo=FALSE,message=FALSE,warning=FALSE}
recent_grads_no_NA %>% 
    arrange(desc(total)) %>% 
    head(20) %>%
    mutate(major=fct_reorder(major,total)) %>%
    gather(gender,number,men,women) %>% 
    ggplot(aes(major,number,fill=gender)) +
    geom_col() +
    scale_y_continuous(labels = comma_format())+
    labs(title="Major categories more common by gender",
         subtitle="",
         x="",
         y="Total # Graduates")+
    coord_flip() +  # rotate the boxplot as there is not clear 
    theme_light() 

```

last one, regresion to see if its correlated gender and low income.

```{r echo=FALSE,message=FALSE,warning=FALSE}
for_interactive_graph <- recent_grads_no_NA %>% 
    mutate(major_category=fct_lump(major_category,4)) %>%
    ggplot(aes(share_women,median,color=major_category,size=sample_size,label=major)) +
    geom_point() +
    geom_smooth(aes(group=1),method="lm") +
    expand_limits(y=0)+
    scale_y_continuous(labels = dollar_format())+
    labs(title="relation share woman / median income by major category",
         subtitle="",
         x="share woman",
         y="median income")+
    theme_light() # theme light

ggplotly(for_interactive_graph)

```

### Model

--

## Communicate

This time was more of a " how to replicate some cool stuff" than things that i thought myself; nevertheless more tools on the belt =) 

### learnt:

it was cool to see the workflow, and how lots of concepts come togheter into an exploratory analysis.

cool stuff: 

- the factors package ( forcats, and fct_reorder )
- the easy ggplot labels ( labs: title, subtitle, x,y  ), always forgot about it
- ggplot remove side reference ( theme(legend.position = "none") )
- the changes on ggplot labels ( dollar_format() ; comma_format() )
- the way of constructing it to an interesting stuff & ordering it for a final pruduct, a blog post.

### future work: 

try to evolve to something like this and spend less amount of time on tech stuff.
more analizing