---
title: "week 2018-11-20 challenge"
author: "_jas"
date: 20181023
output: html_document
---

# Tidy tuesday challenge: Week 2018-11-20

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

- thanksgiving
- tdor (  Transgender Day of Remembrance )

will go for tdor, as looks like more world related =).

## import data

data comes from package: 
details on: https://github.com/CaRdiffR/tdor



```{r echo=FALSE,message=FALSE,warning=FALSE}
library(here)
library(readr)
library(purrr)
library(dplyr)
library(tidyr)
library(janitor)#install.packages('janitor')
library(scales)
library(ggplot2)
library(forcats)#install.packages('forcats')
library(lubridate)
library(visdat)#install.packages('visdat')
library(stringr)
library(DataExplorer)#install.packages('DataExplorer')

library(remotes)#install.packages('remotes')
install_github("CaRdiffR/tdor")
library(tdor)

library(tmap)#install.packages('tmap')
library(magick)#install.packages('remotes')
library(gganimate)#install.packages('gganimate')
library(sf)
# install.packages('ncdf4', type = "source")
data("World")

# library(skimr)#install.packages('skimr')
# library(scales)
# library(ggplot2)
# library(forcats)
# library(lubridate)
# library(cowplot)#install.packages('cowplot')
# library(geofacet)#install.packages('geofacet')
# library(sf)#install.packages('sf')

tdor_raw <- janitor::clean_names(tdor::tdor)
```

# Dataset technical exploration

```{r echo=FALSE,message=FALSE,warning=FALSE}
# glimpse(tdor_raw)
```

as we can see we got a lot of NAs, to 

visually exploring missing data:

```{r echo=FALSE,message=FALSE,warning=FALSE}
DataExplorer::plot_missing(tdor_raw)

```

lets keep no missings or green missings.
```{r echo=FALSE,message=FALSE,warning=FALSE}
tdor_small_missing <- tdor_raw %>% select(-date_1,-source,-tgeu_ref,-photo_source,-photo)
```

# Checking ages:

- might be a range or aproximation: as consecuence we need to lookout before making any graph.

```{r echo=FALSE,message=FALSE,warning=FALSE}
amount_not_taken <- tdor_small_missing %>% filter(str_length(age)>2 ) %>% nrow()
total_amount <- tdor_small_missing %>% nrow()

tdor_age_filter <- tdor_small_missing %>% filter(str_length(age)<=2 ) %>% 
    mutate(five_yrs_min = factor(ceiling(as.integer(age)/5)*5)) %>% 
    mutate(five_yrs_max = factor(floor(as.integer(age)/5)*5)) %>% 
    mutate(five_yrs = paste0(five_yrs_max,"-",five_yrs_min))

tdor_age_filter  %>% 
    count(five_yrs) %>%  
    ggplot(aes(x=five_yrs,y=n,fill=five_yrs)) + 
    geom_col()+
    scale_colour_grey(aesthetics = "fill") +
    coord_flip()+
    labs(title="Transgender deaths by age, 5 year bin",
         subtitle = paste0("age ranges not taken in account (",amount_not_taken," / ",total_amount," registries)"),
         caption = "TDOR data 2018",
         x="",y="# deaths")+
    theme_light()+
    theme(legend.position = "none")
```

# Lets check by country:

```{r echo=FALSE,message=FALSE,warning=FALSE}

tdor_age_filter <- tdor_small_missing %>% filter(str_length(age)<=2 ) %>% 
    mutate(five_yrs_min = factor(ceiling(as.integer(age)/5)*5)) %>% 
    mutate(five_yrs_max = factor(floor(as.integer(age)/5)*5)) %>% 
    mutate(five_yrs = paste0(five_yrs_max,"-",five_yrs_min))


country_names <- maps::iso3166 %>%  
    mutate(mapname_l=str_to_lower(mapname),
           ISOname_l=str_to_lower(ISOname)) %>% 
    count(mapname_l,ISOname_l,a3) %>% select(-n) %>% 
    mutate(std_iso=mapname_l) %>% 
    mutate(std_map=ISOname_l)
# 
# country_names %>% filter(str_detect(country_std,"China"))
# country_names %>% filter(str_detect(std_iso,"Trinidad and Tobago") | 
#                              str_detect(std_map,"Trinidad"))
# country_names %>% filter(str_detect(ISOname_l,"united kingdom"))
# 
# country_names %>% filter(str_detect(ISOname_l,"united kingdom")) %>% pull(ISOname_l)
# 
# maps::iso3166 %>% filter(str_detect(ISOname,"Great Britain"))

tdor_age_filter_map <- tdor_age_filter %>% 
    mutate(country_l=str_to_lower(country)) %>% 
    mutate(country_l=if_else(country_l=="united kingdom",
                             "united kingdom of great britain and northern ireland",
                             country_l)) %>% 
    left_join(country_names,by=c("country_l"="ISOname_l"),suffix = c("_tdor","_iso")) %>% 
    left_join(country_names,by=c("country_l"="mapname_l"),suffix = c("_tdor_iso","_map")) %>% 
    mutate(iso_a3=if_else(is.na(a3_map),a3_tdor_iso,a3_map)) 

tdor_map_count_by_year <- tdor_age_filter_map %>% count(iso_a3,year) %>% rename(deaths=n)

# glimpse(tdor_age_filter_map)

```

# gganimate yearly 

```{r echo=FALSE,message=FALSE,warning=FALSE}
# library(maps)
# library(ggmap)
world_tdor <- World %>% left_join(tdor_map_count_by_year,by=c("iso_a3"="iso_a3")) %>%
    select(iso_a3,name,continent,year,deaths,geometry)

# for each year add NA deaths
tdor_all_years <- tdor_map_count_by_year %>% count(year) %>% pull(year)
tdor_all_countries <- World %>% filter(iso_a3!="ATA")%>%  pull(iso_a3) # all countries except antartica

# tdor_miss_country <- World %>%  anti_join(tdor_map_count_by_year,by=c("iso_a3"="a3_code")) %>% pull(iso_a3)
# all countries and years with miss
tdor_impute_miss <- purrr::map(tdor_all_countries,function(x){
    paste0(tdor_all_years,"|",x) }) %>% 
    unlist() %>% data.frame(stringsAsFactors = FALSE) %>%
    rename(yr_cont=".") %>% as_tibble() %>% 
    separate(col = yr_cont,into=c("year","iso_a3"),sep = "\\|") %>% 
    mutate(year=as.integer(year),iso_a3=as.character(iso_a3)) %>% 
    mutate(deaths=NA)

#
world_tdor_imputed_2 <- World %>% # to world 
    select(iso_a3,name,continent,geometry) %>%  # extract columns of interest
    filter(iso_a3!="ATA") %>%  # remove antartica
    left_join(tdor_impute_miss,by=c("iso_a3"="iso_a3")) %>% # join on impute NA
    left_join(tdor_map_count_by_year,by=c("iso_a3","year")) %>% 
    mutate(deaths=deaths.y) %>%
    select(-deaths.x,-deaths.y) %>% 
    mutate(year=factor(year))

glimpse(world_tdor_imputed_2)
 
world_tdor_imputed_3857 <- sf::st_transform(world_tdor_imputed_2, "+init=epsg:3857")
world_tdor_imputed_3857_gif <- world_tdor_imputed_3857%>% 
    ggplot()+
    geom_sf(aes(fill=deaths))+
    # theme_void()+
  theme(axis.title=element_blank(),
        axis.text=element_blank(),
        axis.ticks = element_blank(),
        strip.text.y = element_text(size = 10),
        panel.background=element_rect(fill="white"))+
    scale_fill_gradient(low="blue", high="red",na.value = "black")+
    gganimate::transition_manual(year,frames = year)+
        labs(title="Transgender deaths by year a country {current_frame}",
         subtitle = paste0("age ranges not taken in account (",amount_not_taken," / ",total_amount," registries)"),
         caption = "TDOR data 2018",
         x="",y="") #%>% gganimate::save_animation(file = "tdor_yearly_2018.gif",)

# gganimate::anim_save(filename = "tdor_yearly_2018.gif")
world_tdor_imputed_3857_gif
```

# tmap yearly 

```{r echo=FALSE,message=FALSE,warning=FALSE}
tdor_yearly_map <- tmap::tm_shape(world_tdor_imputed_2) +
    tm_polygons("deaths")+
    tm_facets(by = "year", nrow=1,ncol=1) +
    tm_legend(#legend.position = c("left", "top"),
      main.title = "Transgender deaths by year and country",
      # title="per 1000 population at risk",
      main.title.size=0.75,
      title.size=0.25,
      legend.text.size=0.25,
      legend.hist.size=0.25,
      # main.title.position = "right",
      legend.outside=TRUE,
      legend.show=FALSE,
      # title.snap.to.legend=FALSE,
      title.position = c("right","bottom"),
      legend.position = c("right","top"))

tmap_animation(tdor_yearly_map, filename="tdor_yearly_map.gif", width=800, delay=100)

```

# communicate:

## data improvments:

- data labels: suggest snake case: use janitor::clean_names
- age: year 1 ?: might be wrong ?
- age: add age bins ? ( 5 / 10 years): so specific age miss can be taken in account
- cause of death: some typos , can be improved via openrefine clustering for similarities.
- cause of death: might be handled as a list ?
- cause of death: some other language issues (ex: 'desmembramiento')
- country: may be add 3 leter code to easier data crossing.
