---
title: "week 2018-12-11 challenge"
author: "_jas"
date: 20181211
output: html_document
---

# Tidy tuesday challenge: Week 2018-12-11

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

nyc restaurants.


## import data


```{r echo=FALSE,message=FALSE,warning=FALSE}
library(here)
library(readr)
library(purrr)
library(dplyr)
library(tidyr)
library(janitor)#install.packages('janitor')
library(scales)
library(ggplot2)
library(forcats)#install.packages('forcats')
library(lubridate)
library(visdat)#install.packages('visdat')
library(stringr)
library(DataExplorer)
library(ggrepel)#install.packages('ggrepel')
# library(rpart)
# library(rpart.plot)
# library(cluster)
# library(leaflet)
```

raw data:

```{r echo=FALSE,message=FALSE,warning=FALSE}
# ~ 100 MB 
# full data at: 
nyc_resto_raw <- readr::read_csv(here::here("data","2018-12-11","nyc_restaurants.csv"))
glimpse(nyc_resto_raw)


nyc_resto_full <- read_csv("https://data.cityofnewyork.us/api/views/43nn-pn8j/rows.csv")
nyc_resto_full <- nyc_resto_full %>% janitor::clean_names()
glimpse(nyc_resto_full)
# Cleaning names with janitor, sampling 300,000 records, and dropping some variables
# nyc_resto_full <- df %>% 
#         janitor::clean_names()
```

# see missing compositions

```{r echo=FALSE,message=FALSE,warning=FALSE}
nyc_resto_processed <- nyc_resto_full %>% 
    mutate(boro=ifelse(boro=="Missing",NA,boro)) %>%  # have to use ifelse , not if_else , nor str_replace
    mutate(inspection_date_parsed=mdy(inspection_date)) %>% 
    mutate(inspection_year=year(inspection_date_parsed),
           inspection_month=month(inspection_date_parsed)) %>% 
    filter(inspection_year>2000)
    
DataExplorer::plot_missing(nyc_resto_processed)

```


```{r echo=FALSE,message=FALSE,warning=FALSE}
nyc_resto_processed %>% 
    count(grade,sort=TRUE)
```


```{r echo=FALSE,message=FALSE,warning=FALSE}
nyc_resto_processed %>% 
    count(boro,sort=TRUE)
```

# amount of inspections by year

* warning 1: for each day inspection, you can have multiple violation codes.


```{r echo=FALSE,message=FALSE,warning=FALSE}
nyc_resto_processed %>% filter(camis==50082219, inspection_date_parsed==date("2018-11-05"))
```

* warning 2: for each day inspaction you can have multiple violation code of same category

```{r echo=FALSE,message=FALSE,warning=FALSE}
nyc_resto_processed %>% 		
    filter(camis==50047379, 
           inspection_date_parsed==date("2018-01-27"),
           violation_code=="15F") 
```

```{r echo=FALSE,message=FALSE,warning=FALSE}

# nyc_resto_processed %>% 
#     count(camis,inspection_date,violation_code,sort=TRUE)
# nyc_resto_processed %>% filter(camis==50082219, inspection_date==date("2018-11-05"))

```

# till now: 

1st we need to understand data structure
as have inspections per day, 
but in each day you can have multiple inspections , 
that was related because to violations. 
if a day got multiple violations , you got multiple registries.

* also we saw that there are many dates out of range; so lets get all the dates before 2000 just to see

```{r echo=FALSE,message=FALSE,warning=FALSE}
nyc_resto_processed %>% filter(inspection_date_parsed<"2000-01-01") %>% count(inspection_year)
```

as seen year is 1900 , so that inspections will be ignored.

ill remove that from the dataset.


lets group by: 
how many violations got each reastaurant per day.



```{r echo=FALSE,message=FALSE,warning=FALSE}
inspections_by_date_by_resto <- nyc_resto_processed %>% 
    mutate(inspection_year=factor(inspection_year)) %>% 
    mutate(score_imputed=if_else(is.na(score),as.integer(0),score)) %>% 
    group_by(inspection_month, inspection_year,  inspection_date_parsed,
           camis,dba,boro,cuisine_description) %>% 
    summarise(violations_per_day=n(),
              violation_codes=paste0(violation_code,collapse = ";"),
              violation_descriptions=paste0(violation_description,collapse = ";"),
              critical_flags=paste0(critical_flag,collapse = ";"),
              total_inspection_score=sum(score_imputed),
              inspection_types=paste0(inspection_type,collapse = ";"),
              total_critical=sum(critical_flag=="Critical"),
              total_not_critical=sum(critical_flag=="Not Critical"),
              total_not_applicable=sum(critical_flag=="Not Applicable"))

inspections_by_date_by_resto %>% 
    ggplot(aes(x=inspection_year,
               y=violations_per_day,
               fill=inspection_year))+ 
    geom_boxplot()+
    theme_light()+
    theme(legend.position = "none")+
    labs(title="Violations per day per restaurant, yearly",
         x="Year",
         y="#Violations per day per restaurant")+
    coord_flip()
```

as seen amount of inspections grown a lot in time; might be more sanity campaings, or that data wasnt registered on the database.


# Wich restaurants have more inspections over time

```{r echo=FALSE,message=FALSE,warning=FALSE}
# Restaurants / inspections
amount_inspections_by_resto <- inspections_by_date_by_resto %>% 
    arrange(inspection_date_parsed) %>% 
    ungroup() %>% 
    group_by(camis,dba,boro,cuisine_description) %>% 
    summarise(inspections_path=paste0(inspection_date_parsed,collapse="->"),
              amount_inspections=n()) %>% 
    nest(-camis,-dba,-boro,-cuisine_description)

nested_amount_inspections_by_resto <- inspections_by_date_by_resto %>% 
    arrange(inspection_date_parsed) %>% 
    ungroup() %>% 
    nest(-camis,-dba,-boro,-cuisine_description)
# 
# nested_amount_inspections_by_resto %>% 
#     glimpse()

nested2 <- nested_amount_inspections_by_resto %>%
    mutate(amount_inspections=unlist(map(data,nrow))) %>% 
    mutate(have_critical=unlist(map(data,function(x){any(x$total_critical>1)}))) %>% 
    mutate(total_critical=unlist(map(data,function(x){sum(x$total_critical)}))) %>% 
    mutate(total_not_critical=unlist(map(data,function(x){sum(x$total_not_critical)}))) %>% 
    mutate(total_not_applicable=unlist(map(data,function(x){sum(x$total_not_applicable)}))) %>% 
    mutate(inspections_path=unlist(map(data,function(x){paste0(x$inspection_date_parsed,collapse="->")})))

unnested2 <- nested2 %>% unnest()

plot_data <- nested2 %>% 
    arrange(desc(amount_inspections)) %>% 
    head(20) %>% 
    mutate(dba=fct_reorder(dba,amount_inspections)) %>% 
    unnest() %>% 
    filter(inspection_year=="2018")

sk <- skimr::skim(plot_data$violations_per_day)
violation_level_threshold_min <- sk %>% filter(stat%in%c("p50")) %>% pull(value) %>% mean()
violation_level_threshold <- sk %>% filter(stat%in%c("p75","p100")) %>% pull(value) %>% mean()

plot_data <- plot_data %>% 
    mutate(violation_amount_level=ifelse(violations_per_day > violation_level_threshold, "red", 
                              ifelse(violations_per_day<=violation_level_threshold_min,"green","blue"))) %>% 
    mutate(total_secerity=total_critical+total_not_critical+total_not_applicable) %>% 
    mutate(ratio_critical=total_critical/total_secerity)

# plot_data %>% filter(ratio_critical >= 0.5)


# plot_data %>% filter(violation_amount_level=="green")
plot_nested <- plot_data %>% 
    ggplot(aes(x=inspection_date_parsed,
               y=dba,
               label=paste0("#violations: ", violations_per_day),
               color=violation_amount_level,
               size=violations_per_day))+
    # geom_point(alpha=0.2,
    #            size=plot_data$violations_per_day+1,
    #            color=ifelse(plot_data$ratio_critical>0.7,"orange","grey"))+
    geom_point(alpha=0.5)+
    # geom_text_repel(data=subset(plot_data,ratio_critical>0.7) ,
    #                 hjust=50,
    #                 # x_nudge=50,
    #                 direction = "x",
    #                 size=3,
    #                 segment.size = 0.4,
    #                 segment.color="blue",
    #                 color="blue"
    #                 )+
    geom_text_repel(data=subset(plot_data,violations_per_day > violation_level_threshold) ,

                    hjust=25,
                    # x_nudge=50,
                    direction = "x",
                    size=3,
                    segment.size = 0.4,
                    segment.color="blue",
                    color="blue"
                    )+
    scale_fill_gradient(low="blue", high="red",na.value = "black")+
    labs(title="Inspections by restaurant in 2018 ,Top #20",
         subtitle=paste0("highlight threshold: ",violation_level_threshold_min,"<= violations < ",violation_level_threshold),
         x="",
         y="")+
    scale_x_date(date_breaks = "1 month", date_labels = "%Y-%m-%d") +
    theme_light()+
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 45, hjust = 1))


plot_nested
# https://cran.r-project.org/web/packages/ggrepel/vignettes/ggrepel.html
```


```{r echo=FALSE,message=FALSE,warning=FALSE}

```


```{r echo=FALSE,message=FALSE,warning=FALSE}
# nested_amount_inspections_by_resto$data[[1]]
# nested_amount_inspections_by_resto$amount_inspections[[1]]


# str(nested_amount_inspections_by_resto$data)
# 
# 
# glimpse(amount_inspections_by_resto)
# 
# glimpse(amount_inspections_by_resto %>% separate(inspections_path,sep="->"))
# 
# amount_inspections_by_resto %>% 
#     arrange(desc(amount_inspections)) %>% 
#     # head(50) %>% 
#     # ungroup() %>% 
#     # mutate(dba=fct_reorder(dba,amount_inspections)) %>% 
#     ggplot(aes(y=amount_inspections)) +
#     geom_boxplot()+
#     labs(title="# inspections by restaurant",
#          y="# inspections by restaurant",
#          x="")+
#     theme_light()+
#     theme(legend.position = "none")+
#     coord_flip()
# 
# amount_inspections_by_resto %>% select(amount_inspections)
#     skimr::skim(amount_inspections_by_resto$amount_inspections)
#     
#     amount_inspections_by_resto %>% 
#     ggplot(aes(x=dba,y=am))

    
    
    
amount_inspections_by_resto %>% 
    arrange(desc(amount_inspections)) %>% 
    head(50) %>% 
    ungroup() %>% 
    mutate(dba=fct_reorder(dba,amount_inspections)) %>% 
    ggplot(aes(x=dba,y=amount_inspections,fill=dba)) +
    geom_col()+
    labs(title="# inspections by restaurant",
         y="# inspections by restaurant",
         x="")+
    theme_light()+
    theme(legend.position = "none")+
    coord_flip()

```


```{r echo=FALSE,message=FALSE,warning=FALSE}

```




```{r echo=FALSE,message=FALSE,warning=FALSE}
# nyc_resto_processed %>% filter(camis==50082219, inspection_date==date("2018-11-05"))
```


```{r echo=FALSE,message=FALSE,warning=FALSE}
# why there are so much NA in violation_description ?

# guess = explore data
# well im just being vague ... not having an objective atm ... 
#that will lead to N time , need to be more effective
# think = make a model
# its fun but ... not having a clear obejctive ... about all the analysis
# ill better think a little 1st.
# missing_violation_code <- nyc_resto_processed %>% 
#     mutate(missing_violation_code=is.na(violation_code))
# 
# rpart
# 
# 
#     filter(camis==50047379, 
#            inspection_date_parsed==date("2018-01-27"),
#            violation_code=="15F") 


```



```{r echo=FALSE,message=FALSE,warning=FALSE}
# 
# %>% 
#     select(camis,inspection_date,violation_code,violation_description,action) %>% 
#     count(camis,inspection_date,violation_code,sort=TRUE)
# 
# nyc_resto_processed %>% 
#     count(violation_code,violation_description) %>% 
#     filter(is.na(violation_description))
# 
# 
# nyc_resto_processed %>% filter(is.na(violation_code))
# 
# nyc_resto_processed %>% 
#     filter(is.na(violation_code)) %>% 
#     select(camis,inspection_date,inspection_date_parsed) %>% 
#     filter()
# 
#     min(nyc_resto_processed$inspection_date)
# nyc_resto_raw$inspection_date
```

# communicate

too much random time assigned to this week . 
cannot focus on one stuff in a session.
for next time, just remember: 
- explore the data structure of the columns
- think some questions to ask , may be thinking who can be interested can guide.
- focus on answer some questions

learnt:

ggrepel 
more ggplot
and realizing that having inspection "paths" might not be a good point of view for just a static visualization; may be experiment with something interactive in shiny.

