---
title: "Tidy tuesday challenge: Week 2019-05-07 UNESCO Student Teacher Ratio"
author: "julio"
date: "12 de mayo de 2019"
output: html_document
---

# Tidy tuesday challenge: Week 2019-05-07  UNESCO Student Teacher Ratio

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

### Data:

UNESCO Student Teacher Ratio

### objectives:

- explore and decide

## details:


## import data
```{r echo=FALSE,message=FALSE,warning=FALSE}
library(dplyr)
library(tidyr)
library(skimr)
library(readr) 
library(ggplot2)
library(scales)
library(forcats)
library(lubridate)
library(gganimate)
library(animation)

library(stringr)#;
library(rebus)#; install.packages('rebus')
library(tidytext)

library(rworldxtra) # para datos iso3, segun gabo
library(sf)

# devtools::install_github('rensa/ggflags')
library(ggflags) # https://github.com/rensa/ggflags

# library(widyr) # for pairwise core
# library(ggraph) # for graph drawing
# library(igraph)  # for graph managing

```


```{r echo=FALSE,message=FALSE,warning=FALSE}
student_ratio <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-05-07/student_teacher_ratio.csv")
```

# explore data

```{r echo=FALSE,message=FALSE,warning=FALSE}
glimpse(student_ratio)
# glimpse(mp_light)

# bird_collisions %>% count(family,genus,species,sort = TRUE)

```

```{r echo=FALSE,message=FALSE,warning=FALSE}
skimr::skim(student_ratio)
# skimr::skim(mp_light)

```


```{r echo=FALSE,message=FALSE,warning=FALSE}
student_ratio %>% 
    count(flags)

student_ratio %>% 
    count(flag_codes)

student_ratio %>% 
    count(indicator)

student_ratio %>% 
    count(year)
```


```{r echo=FALSE,message=FALSE,warning=FALSE}
student_ratio %>% 
    filter(year < 2018) %>% 
    group_by(indicator,year) %>% 
    summarise(world_total_ratio=sum(student_ratio,na.rm = TRUE)) %>% 
    ungroup() %>% 
    mutate(indicator=fct_reorder(indicator,world_total_ratio)) %>% 
    ggplot(aes(x=year,y=world_total_ratio,fill=indicator))+
    geom_area()+
    labs(x="",y="Sumemd Ratio",fill="Education Level",
         title="World total Student/Teacher Ratio",
         subtitle="From 2012 to 2017")
```
```{r echo=FALSE,message=FALSE,warning=FALSE}
# Based on: https://twitter.com/packpridejones/status/1127032282517266433


data("countriesHigh", package = "rworldxtra")
# listado paises para SF ( los mapas )
countries <- countriesHigh %>% 
    st_as_sf() %>%
    as_tibble() %>% 
    filter(REGION != "Antarctica") %>% 
    select(NAME, ISO_A3,ISO_A2,REGION) %>% 
    mutate(ISO_A2=str_to_lower(ISO_A2))

student_ratio_with_codes <- student_ratio %>% 
    left_join(countries,by = c("country_code"="ISO_A3"))
student_ratio_with_codes %>% filter(str_detect(REGION,"South")) %>% count(NAME,ISO_A2)

student_ratio_with_codes %>% count(indicator)


```


```{r echo=FALSE,message=FALSE,warning=FALSE}

filtered_data <- student_ratio_with_codes %>% 
    filter(year < 2018) %>% 
    # filter(str_detect(REGION,"South")) %>% 
    filter(ISO_A2 %in% c('ar','br','cl','co','ec','gd','gy','py','pe','uy','ve')) %>%     filter(indicator=="Primary Education") #%>% 
    # mutate(year=as.character(year)) 

# but with flags :D 
mean_y <-  mean(filtered_data$student_ratio,na.rm = TRUE)

original_anim <- filtered_data%>% 
    ggplot(aes(year,student_ratio,color=country_code))+
    # geom_jitter(alpha=0.5)+
    geom_line()+
    geom_point(alpha=0.5)+
    geom_flag(data=filtered_data,
        aes(country = ISO_A2),
        size = rel(5),alpha=0.5)+
    labs(x="",y="Student/Teacaher Ratio",
         color="Country",
         title="Primary Education Student Teacher Ratio on South America",
         subtitle="2012/2017; missing values on some countries.",
         caption = "#tidytuesday")

#+
    # geom_hline(yintercept = mean_y,color="orange")+
    # coord_flip()+
    # theme(panel.background = element_blank())+
    # transition_states(year,
    #                   transition_length = 1,
    #                   state_length = 0.1,
    #                   wrap = FALSE)+
    # enter_fade()+
    # exit_fade()+
    # labs(title="Year: {closest_state}",
    #      x="",y="student ratio",
    #      caption="#tidytuesday")
    

# animate(original_anim, fps = 6)
# 
# anim_save("student_ratio.gif",
#           original_anim)


ggsave(plot = original_anim,filename = "20190507_challenge.png")


```


# communicate

well not best week not best times ... just keep rowing.



