---
title: "Tidy tuesday challenge: Week 2019-02-26 french train delays"
author: "julio"
date: "2019-03-03"
output: html_document
---

# Tidy tuesday challenge: Week 2019-02-26 french train delays

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

### Data:

this week data its related to PhDs Awarded by Field

### objectives:

- issues on git syhcroing, switching to an interesting visualization and reproduce.
- just went to a graph visualization. showing how are linked from/to , and if its tgv as color.

## details:

- selected small trains dataset

## import data

```{r echo=FALSE,message=FALSE,warning=FALSE}
library(dplyr)
# library(Rcpp) #; install.packages("Rcpp")
library(skimr) #; install.packages("skimr")
# library(circlepackeR) #; devtools::install_github("jeromefroe/circlepackeR")
# library(data.tree)  #; install.packages("data.tree")

library(igraph)
library(visNetwork)

```



```{r echo=FALSE,message=FALSE,warning=FALSE}
small_trains <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-02-26/small_trains.csv") 


```


```{r echo=FALSE,message=FALSE,warning=FALSE}
glimpse(small_trains) 


small_trains %>% 
    count(service)

```

```{r echo=FALSE,message=FALSE,warning=FALSE}
skimr::skim(phd_raw) 


```

## tidy

always wante to try circle pack, so there we go: prepare data: deprecated -- shifted to the shiny app of the guy as i wan to have a functional deployable that i can give other people.
not just my pc.

```{r echo=FALSE,message=FALSE,warning=FALSE}
# library(circlepackeR) #; devtools::install_github("jeromefroe/circlepackeR")
# 
# phd_processed <- phd_raw %>% filter(year==2008) %>% select(-year) %>%  
#         mutate(path_string = paste("phds", broad_field, major_field, field, sep = "/"))
# # sample: http://shiny.rstudio.com/gallery/submitbutton-demo.html
# # UI
# 
# # library(data.tree)
# # phd_processed_circler <- data.tree::as.Node(phd_processed)
# 
# 
# phd_processed %>% count(field) %>%  filter(n>1)
# 
# phd_processed %>% filter(field=="Environmental toxicologyc")
# 
# circlepackeR::circlepackeR(phd_raw)
# 
# 
# 
# phd_processed2 <- phd_raw %>% filter(year==2008) %>% select(-year) %>% 
#     mutate(from=) %>% mutate( to="")
# 
# 
# # Libraries
# library(ggraph)
# library(igraph)
# library(tidyverse)
# library(viridis)
#  
# # We need a data frame giving a hierarchical structure. Let's consider the flare dataset:
# edges=flare$edges
# vertices = flare$vertices
# mygraph <- graph_from_data_frame( edges, vertices=vertices )
#  
# # Control the size of each circle: (use the size column of the vertices data frame)
# # png("~/Dropbox/R_GG/R_GRAPH/#314_custom_circle_packing1.png", height = 480, width=480)
# ggraph(mygraph, layout = 'circlepack', weight="size") + 
#   geom_node_circle() +
#   theme_void()


```

## visualize

as i want interactive for the shiny app im using package: circlepackeR: canceled.
using the guy shiny app as ive made some stuff with that library before should not be complex.

```{r echo=FALSE,message=FALSE,warning=FALSE}
# library(circlepackeR)
# library(data.tree)
# library(treemap)#;install.packages("treemap"); install.packages("httpuv"); install.packages("mime")
# 
# data(GNI2014)
# head(GNI2014)
# 
# GNI2014$pathString <- paste("world", 
#                             GNI2014$continent, 
#                             GNI2014$country, 
#                             sep = "/")
# population <- as.Node(GNI2014)
# 
# circlepackeR(population, size = "population", color_min = "hsl(56,80%,80%)", 
#              color_max = "hsl(341,30%,40%)")


```


## communicate

### general view:

#### issues with git 

remeber any step on command line, you previous gotta be on the repository folder. 
like: 'cd /user/me/git/tidytuesday'

got issues with git syncrho. 
generally use windows desktop , now on linux notebook
i got the repo on dropbox folder. so config of windows its shared on dropbox and autosynchroed.
that causes issues as both computers got different context. 

- on windows got the  autocrlf, to true. 

so when i downloaded from github, it transformed the lf, to crlf > making the linux version see 'differences' just when i downloaded the repo.

to solve this i've set it to false:

git config --global core.autocrlf false

- issues on this computer sychro, on windows i use notepad++ , it wont work right on linux, so as alternative i use atom. 

nevertheless, it is not so happy when using terminal, at least on my experience at setting it as default editor. 
after commiting and filling the message on the editor, and closing it ( file & atom ) , the terminal keeps saying waiting to close ... 

i used this command to set atom as default:

git config --global core.editor "atom --wait"

i used this one to fall back to default editor on my notebook 
( my default is pluma, you gotta search which is the name of your default editor. )

git config --global core.editor "pluma"


as i already was in the problem, of getting commited or undone things ... 
i just fall back to previous commit that ive done. 

to know which was i looked at: 

git log

and see the log searching for the commit that i was the last to submit.
then when i see: 
Merge: 75253ed 354de7b

the id to reset its the 1st part '75253ed'

WARNING: THIS WILL DROP WHATEVER YOU HAVE DONE AFTER THAT POINT !

if you decide to move further, to do the reset all you need to go is: 

git reset --hard 75253ed3

Then after that i resinchroed to the current week data, uploads like: 

git pull https://github.com/rfordatascience/tidytuesday.git master

it was quite time consuming to debug this, so ill just go for one interesing visualization and try to reproduce.


