
---
title: "Tidy tuesday challenge: Week 2019-03-12 board games data"
author: "julio"
date: "16 de febrero de 2019"
output: html_document
---

# Tidy tuesday challenge: Week 2019-03-12 board games data

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

### Data:

this week data its related to 2019-03-12 board games data

### objectives:

- ggforce voronoi stuff, and try to make it usefull :p 

## import data

```{r echo=FALSE,message=FALSE,warning=FALSE}
library(here)
library(readr)
library(dplyr)
library(tidyr)
library(scales)
library(ggplot2)
library(forcats)#install.packages('forcats')
library(lubridate)
# library(visdat)#install.packages('visdat')
library(stringr)
# library(naniar)
# library(rpart)
# library(rpart.plot)
library(cowplot)
# library(rstan) # install.packages("rstan")
# library(prophet)# install.packages("prophet")
# library(ggrepel)
library(ggforce)# install.packages("ggforce")
# devtools::install_github("9thcirclegames/bgg-analysis")
library(bggAnalysis) # install.packages("bggAnalysis") 
```


# first test the existent default example

```{r echo=FALSE,message=FALSE,warning=FALSE}

# voronoi
ggplot(iris, aes(Sepal.Length, Sepal.Width)) + 
  geom_voronoi_tile(aes(fill = Species, group = -1L), max.radius = 1,
                    colour = 'black', expand = unit(-0.5, 'mm'), 
                    radius = unit(0.5, 'mm'), show.legend = FALSE)


# parallel sets

titanic <- reshape2::melt(Titanic)
# This is how we usually envision data for parallel sets
# head(titanic)
titanic <- gather_set_data(titanic, 1:4)
# head(titanic)
ggplot(titanic, aes(x, id = id, split = y, value = value)) +
  geom_parallel_sets(aes(fill = Sex), alpha = 0.3, axis.width = 0.1) +
  geom_parallel_sets_axes(axis.width = 0.1) +
  geom_parallel_sets_labels(colour = 'white')

```

# think about current week data and put it there

```{r echo=FALSE,message=FALSE,warning=FALSE}

data(BoardGames)
head(BoardGames)

bg_processed <- BoardGames %>% janitor::clean_names()
glimpse(bg_processed)
```


```{r echo=FALSE,message=FALSE,warning=FALSE}
# ok ... lets go for the shor tversion: 

board_games <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-03-12/board_games.csv")
glimpse(board_games)
```


```{r echo=FALSE,message=FALSE,warning=FALSE}

boards_processed <- board_games %>% 
    mutate(decade= floor(year_published / 10)*10) 


board_categories <- boards_processed %>% 
    select(category,decade,year_published,game_id) %>% 
    separate_rows(category,sep = ",") 


board_mechanics <- boards_processed %>% 
    separate_rows(mechanic,sep = ",") %>% 
    count(decade,mechanic,sort = TRUE)

boards_processed %>% count(decade,sort = TRUE)
```



```{r echo=FALSE,message=FALSE,warning=FALSE}
boards_processed %>% 
    count(category,sort = TRUE) %>% 
    filter(str_detect(str_to_lower(category),"war"))

```

# visualization

## avg rating by decade

```{r echo=FALSE,message=FALSE,warning=FALSE}

font_selected <- "Roboto"
col_brew_selected <-  "Accent" # RColorBrewer::display.brewer.all()
boards_processed %>%
    ggplot(aes(x=year_published,y=average_rating,colour=as.factor(decade))) +
    geom_point()+
    scale_y_continuous(limits = c(1,10), breaks = seq(1,10,by = 1)) +
    scale_x_continuous(breaks = c(seq(1950,2010, by = 10))) + 
        labs(title= "Average game rating by decade",
         x="Year published",
         y="Avg. rating",
         color="Decade") +
    theme_dark()+
    # scale_color_brewer(palette = col_brew_selected, direction = -1) +
    scale_color_brewer(palette = col_brew_selected, direction = 1) +
    theme(
    text = element_text(family = font_selected, color = "gray90"),
    legend.position = "bottom",
    plot.background = element_rect(fill = "gray30"),
    legend.background = element_rect(fill = "gray30"), 
    plot.title = element_text(size = 20,face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.text = element_text(family = font_selected, color = "gray90"),
    axis.title = element_text(family = font_selected, color = "gray90", face = "bold"),
    panel.grid.minor.y = element_blank(), 
    panel.grid.major.y = element_line(size =0.1),
    panel.grid.major.x = element_line(size =0.1),
    panel.grid.minor.x = element_blank()
    )+
      guides(
    color = guide_legend(
      title = "Decade",
      title.position = "top",
      title.hjust = 0.5,
      nrow = 1,
      label.position = "bottom",
      override.aes =list(alpha = 1,size = 4))
)+
    ggsave(filename=here::here("jas1_weeks","2019","2019-03-12","harro_exp.png"))


    

```

## game categoriew by decade


```{r echo=FALSE,message=FALSE,warning=FALSE}


board_categories %>%
    count(decade,category) %>% 
    filter(n>10) %>% 
    filter(!is.na(category)) %>% 
    mutate(category=fct_reorder(category,n)) %>% 
    ggplot(aes(x=category ,y=n,fill=as.factor(decade))) +
    geom_col()+
    # theme(
    #     # legend.position = "none",
    #     # text = element_text(size = rel(0.2)),
    #     axis.text.y = element_text( hjust = 1,size = rel(0.5))
    # )+
    coord_flip()+
    labs(title = "Board game categories by decade",
         subtitle="filtered NA & N < 10",
         caption = "#tidytuesday 2019-03-12 ",
         x="",
         y="# category/decade",
         fill="Decade")+
    theme_dark()+
    # scale_color_brewer(palette = col_brew_selected, direction = -1) +
    scale_color_brewer(palette = col_brew_selected, direction = 1) +
    theme(
    text = element_text(family = font_selected, color = "gray90"),
    # legend.position = "bottom",
    plot.background = element_rect(fill = "gray30"),
    legend.background = element_rect(fill = "gray30"), 
    plot.title = element_text(size = 20,face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.text = element_text(family = font_selected, color = "gray90"),
    axis.text.y = element_text(family = font_selected, color = "gray90",size = rel(0.5)),
    axis.title = element_text(family = font_selected, color = "gray90", face = "bold"),
    panel.grid.minor.y = element_blank(), 
    panel.grid.major.y = element_line(size =0.1),
    panel.grid.major.x = element_line(size =0.1),
    panel.grid.minor.x = element_blank()
    ) +
    ggsave(filename=here::here("jas1_weeks","2019","2019-03-12","category_decade_exp.png"))

```


## paralell set (~sankey)


```{r echo=FALSE,message=FALSE,warning=FALSE}

boards_for_parallel_set <- board_games %>% 
    mutate(decade= floor(year_published / 10)*10) %>% 
    select(category,decade,game_id,mechanic) %>% 
    separate_rows(category,sep = ",") %>% 
    separate_rows(mechanic,sep = ",") %>% 
    count(decade,category,mechanic,sort = TRUE) %>% 
    nest(-decade) %>% 
    mutate(top_10=purrr:::map(.x=data,function(x){
        # print(head(x))
        res <- data %>% head(10) # %>% select(category,n)
        unlist(res)
    })) %>% unnest(top_10)

boards_for_parallel_set_plot <- gather_set_data(boards_for_parallel_set, 1:3)

boards_for_parallel_set_plot_2 <- boards_for_parallel_set_plot %>% 
    nest(-decade) %>% 
    mutate(top_10=purrr:::map(.x=data,function(x){
        print(head(x))
        
        data %>% unlist() %>% head(10) %>% select(category,n)
    })) %>% unnest(top_10)
    
  boards_for_parallel_set_plot_2 %>% 
  ggplot(aes(x, id = id, split = y, value = n)) +
  geom_parallel_sets(aes(fill = category), alpha = 0.3, axis.width = 0.1) +
  geom_parallel_sets_axes(axis.width = 0.1) +
  geom_parallel_sets_labels(colour = 'white') +
  theme(legend.position = "none")
```


# voronoi

```{r echo=FALSE,message=FALSE,warning=FALSE}

board_games_voro <-  board_games %>% 
    mutate(decade= as.factor(floor(year_published / 10)*10)) #%>% 
    # head(board_games_voro)
```
 
```{r echo=FALSE,message=FALSE,warning=FALSE}   

ggplot(board_games_voro, aes(average_rating, users_rated)) +
    geom_point(aes(fill = decade),shape=21) 

board_games_voro %>% 
    filter(users_rated>300 & users_rated<1000) #%>% 
    # count()

ggplot( aes(decade, users_rated)) +
    geom_boxplot()

ggplot(board_games_voro, aes(decade, users_rated)) +
    # geom_boxplot()
    geom_line()

board_games_voro %>% 
    # filter(users_rated>300 & users_rated<1000) %>% 
    head(50) %>% 
ggplot(aes(average_rating, users_rated)) +
    # geom_point(aes(fill = decade),shape=21)
  geom_voronoi_tile(aes(fill = decade, group = -1L), max.radius = 1,
                    colour = 'black', expand = unit(-0.5, 'mm'), 
                    radius = unit(0.5, 'mm'), show.legend = FALSE)
```

