
---
title: "Tidy tuesday challenge: Week 2019-05-28 wines!"
author: "julio"
date: "21 - 5 - 2019"
output: html_document
---

# Tidy tuesday challenge: Week 2019-05-28 wines!

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

### Data:

Global Plastic Waste our world in data

### objectives:

- explore & do a quick one , short time this week

## details:


## import data
```{r echo=FALSE,message=FALSE,warning=FALSE}
library(dplyr)
library(tidyr)
library(skimr)
library(readr) 
library(ggplot2)
library(scales)
library(forcats)
library(lubridate)
library(gganimate)
library(animation)

library(stringr)#;
library(rebus)#; install.packages('rebus')
library(tidytext)
# 
# library(rworldxtra) # para datos iso3, segun gabo
# library(sf)

# devtools::install_github('rensa/ggflags')
# library(ggflags) # https://github.com/rensa/ggflags

# library(widyr) # for pairwise core
# library(ggraph) # for graph drawing
# library(igraph)  # for graph managing

# install.packages("waffle")
# devtools::install_github("hrbrmstr/waffle")
library(waffle)
library(countrycode)
```


```{r echo=FALSE,message=FALSE,warning=FALSE}
wine_ratings <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-05-28/winemag-data-130k-v2.csv")

wine_ratings_clean <- wine_ratings %>% rename(id=X1)
```

# explore data

```{r echo=FALSE,message=FALSE,warning=FALSE}
glimpse(wine_ratings)

```


```{r echo=FALSE,message=FALSE,warning=FALSE}
wine_ratings_clean %>% skimr::skim()

# nobel_winner_all_pubs$
```

# wineries by country & price

```{r echo=FALSE,message=FALSE,warning=FALSE}
wineries_country_price <-  wine_ratings_clean %>% 
    select(country,winery,price) %>% 
    filter(complete.cases(.))

wine_data_plot_1 <- wineries_country_price %>% arrange(country) %>% 
    group_by(country) %>% 
    summarise(q01=quantile(price,probs = 0.01),
              q25=quantile(price,probs = 0.25),
              q5=quantile(price,probs = 0.5),
              q75=quantile(price,probs = 0.75),
              q99=quantile(price,probs = 0.99)) %>% 
    gather(quantile,price,-country) 
    

wine_data_plot_1 %>% 
    ungroup() %>% 
    mutate(country=fct_reorder(country,price)) %>% 
    ggplot(aes(x=quantile,y=price,fill=country))+
    geom_col()+
    coord_flip()

# wine_data_plot_1 %>% filter(str_detect(country,"gu") )
wine_data_plot_1 %>% 
    filter(country %in% 
               c('Argentina','Brazil','Chile','Colombia','Ecuador',
                 'French Guyana','Suriname','Guyana',
                 'Paraguay','Peru','Uruguay','Venezuela')) %>% 
    ungroup() %>% 
    mutate(country=fct_reorder(country,price)) %>% 
    ggplot(aes(x=country,y=price,fill=quantile))+
    geom_col()+
    coord_flip()
```
# types of wines & countries

```{r}
country_variety <-  wine_ratings_clean %>% 
    select(country,variety) %>% 
    filter(complete.cases(.)) 

with_100_reviews <- country_variety %>% 
    count(country,variety) %>% 
    filter(n>100)
    
top_20_varieties <- wine_ratings_clean %>% 
    select(country,variety) %>% 
    filter(complete.cases(.)) %>% 
    count(variety,sort=TRUE) %>% 
    head(20)

top_20_countries <- wine_ratings_clean %>% 
    select(country,variety) %>% 
    filter(complete.cases(.)) %>% 
    count(country,sort=TRUE) %>% 
    head(20)

top_20_critics <- wine_ratings_clean %>% 
    select(taster_name,variety) %>% 
    filter(complete.cases(.)) %>% 
    count(taster_name,sort=TRUE) %>% 
    head(20)


top_20_varieties_country <- wine_ratings_clean %>% 
    select(country,variety) %>% 
    filter(complete.cases(.)) %>% 
    count(country,variety,sort=TRUE) %>% 
    head(20)


```

## Chord diagram



```{r}
# install.packages("circlize")
library(circlize)
# RColorBrewer::display.brewer.all()
pallete_colors <- RColorBrewer::brewer.pal(n=8,"Set3")
grid.col = c(Argentina = pallete_colors[1],
             Portugal = pallete_colors[2],
             Germany = pallete_colors[3],
             US = pallete_colors[4],
             France = pallete_colors[5],
             Italy = pallete_colors[6])

#https://www.color-hex.com/color-palette/74955
#https://www.color-hex.com/color-palette/59373

# grid.col = c(Argentina = "#842896",
#              Portugal = "#c099b8",
#              Germany = "#927498",
#              US = "#653e52",
#              France = "#6c003a",
#              Italy = "#9c2b48")

#eebed8 	(238,190,216)
	#c099b8 	(192,153,184)
	#927498 	(146,116,152)
	#7f3a69 	(127,58,105)
	#6c003a 	(108,0,58)


top_20_varieties_country_2 <- top_20_varieties_country %>% 
    mutate(variety=
               if_else(variety=="Bordeaux-style Red Blend",
        "Bordeaux-style\nRed Blend",
        if_else(variety=="Cabernet Sauvignon",
                "Cabernet\nSauvignon",
                variety)))

chordDiagram(top_20_varieties_country_2, 
annotationTrack = "grid", 
preAllocateTracks = list(track.height = 0.1),
grid.col = grid.col)


panel_fun_1 <- function(x, y) {
  xlim = get.cell.meta.data("xlim")
  ylim = get.cell.meta.data("ylim")

  sector.name = get.cell.meta.data("sector.index")

  circos.text(mean(xlim), ylim[1] + .1, sector.name,
              facing = "clockwise", niceFacing = TRUE,
              adj = c(0, 0.5))

  circos.axis(h = "top", labels.cex = 0.5,
              major.tick.percentage = 0.2,
              sector.index = sector.name, 
              track.index = 2)
}
#https://jokergoo.github.io/circlize_book/book/the-chorddiagram-function.html

# https://rstudio-pubs-static.s3.amazonaws.com/145337_0ecf43312d7b42aaa6b4687649915879.html
panel_fun_2 <-  function(x, y) {
  xlim = get.cell.meta.data("xlim")
  xplot = get.cell.meta.data("xplot")
  ylim = get.cell.meta.data("ylim")
  sector.name = get.cell.meta.data("sector.index")
  if(abs(xplot[2] - xplot[1]) < 10) {
    circos.text(mean(xlim), ylim[1], sector.name, facing = "clockwise",
                niceFacing = TRUE, adj = c(0, 0.5))
  } else {
    circos.text(mean(xlim), ylim[1], sector.name, facing = "inside",
                niceFacing = TRUE, adj = c(0.5, 0))
  }
}

circos.trackPlotRegion(track.index = 1,
                       panel.fun = panel_fun_2,
bg.border = NA)
# RColorBrewer::display.brewer.all()
title(main = "Top Countries & wine Varieties",sub = "#TidyTuesday")

```

#tweet: 

https://twitter.com/jspairani/status/1135403749416067073


# communicate

well chord diagram finally but i really dont like the syntax :/

love the customization.

here some usefull links: 

- https://jokergoo.github.io/circlize_book/book/the-chorddiagram-function.html

- https://rstudio-pubs-static.s3.amazonaws.com/145337_0ecf43312d7b42aaa6b4687649915879.html

also some palettes: but didnt use em in the final version.

- https://www.color-hex.com/color-palette/74955
- https://www.color-hex.com/color-palette/59373
