
---
title: "Tidy tuesday challenge: Week 2019-06-18 xmas birds"
author: "julio"
date: "2019-06-11"
output: html_document
---

# Tidy tuesday challenge: Week 2019-06-18 xmas birds !

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

### Data:

xmas birds


### objectives:

check the data and do something.

## details:


## import data
```{r echo=FALSE,message=FALSE,warning=FALSE}
library(magrittr) # para el %T>%
library(tidyverse)

library(stringr)#;
library(rebus)#; install.packages('rebus')
library(tidytext)


```


```{r echo=FALSE,message=FALSE,warning=FALSE}
bird_counts <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-18/bird_counts.csv")
```

# explore data

```{r echo=FALSE,message=FALSE,warning=FALSE}
glimpse(bird_counts)

```


```{r echo=FALSE,message=FALSE,warning=FALSE}
bird_counts %>% skimr::skim()

```

can see several missing on how many counted by hour, and total hours.
will omit the missing val columns

```{r echo=FALSE,message=FALSE,warning=FALSE}

bird_counts %>% 
    select(species_latin,year,how_many_counted) %>% 
    group_by(species_latin,year) %>% 
    summarise(total_how_many_counted=sum(how_many_counted), total_obs=n()) %>% 
    filter(total_obs >1)

```

can see that there are no multiple inputs by species / year.
so every year have the total amount of species.
so lets take a look to how much species by year


```{r echo=FALSE,message=FALSE,warning=FALSE}

bird_counts %>% 
    select(species_latin,year,how_many_counted) %>% 
    group_by(year) %>% 
    summarise(total_how_many_counted=sum(how_many_counted), total_species_obs=n())  %>% 
    filter(total_species_obs != 199)

```
also can see that every species have an input each year. at least with 0 value.
as we dont want the zeroes we want the total species, lets ommit the how many counted == 0

```{r echo=FALSE,message=FALSE,warning=FALSE}

total_species_by_year <- bird_counts %>% 
    select(species_latin,year,how_many_counted) %>% 
    filter(how_many_counted > 0) %>% 
    group_by(year) %>% 
    summarise(total_counted=sum(how_many_counted), total_species_obs=n()) %>% 
    mutate(counted_vs_species=total_counted/total_species_obs) %>% 
    gather(time_series,value,-year) %>% 
    mutate(xmas_date=lubridate::as_date(paste0(year,"-12-25")))

```


# 3 series , 
```{r echo=FALSE,message=FALSE,warning=FALSE}


# for_labels <- lubridate::year(total_species_by_year$xmas_date)

series_plot <- total_species_by_year %>% 
    
    ggplot(aes(x = xmas_date, y = value,color=time_series))+
    geom_line()+
    scale_x_date(date_breaks = "5 years",date_labels = "%Y")+
    facet_grid(time_series~.,scales = "free_y")+
    theme_light()+
    theme(legend.position = "none",
          axis.text.x = element_text(angle=90))+
    labs(x="",y="",
         title="Progression over time of Christmas Bird Counts from Hamilton-Ontario",
         subtitle=paste0(min(total_species_by_year$year),"-",max(total_species_by_year$year)),
         caption = "#TidyTuesday") 


series_plot %T>% 
    print() %T>%
    ggsave(plot=.,"xmas_birds_ontario.png",width = 10,height = 7)
    

```


# idea , do prophet stuff to see how it evolves over time.

well not enough time today to do this.
will be next time :p 

```{r echo=FALSE,message=FALSE,warning=FALSE}
# library(prophet)
# 


```



#tweet: 

2019-06-18 #TidyTuesday #rstats Xmas Birds Hamilton-Ontario! 
something simpel this time, saw the time evolution of the bird counts, then thested the %T>% operator, posted recently by: 
@andrewheiss here: https://twitter.com/andrewheiss/status/1141853448591200256
next time: would be nice to do some forecast for 2018/2019 and then check with the actual values =).

https://twitter.com/jspairani/status/1142957944310386688

# communicate

- not much time , so jus tchecked the data do some ggplot & teste dthe %T>% operator. its amazing :D


