---
title: "Week 20181113"
author: "_jas"
date: 2018-11-13
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tidy tuesday challenge: Week 2018-11-13

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

- malaria atlas : just a wish: discarded for this week challenge, nevertheless might get in later.
- will use the malaria simpler csv's ; just to get into the domain

# Workflow:

## Import

```{r load_libs }
library(here)
library(readr)
library(janitor)
library(skimr)
library(ggplot2)
library(dplyr)
library(DataExplorer)
library(tmap)
library(tidyr)
# library(rlang)#; install.packages("rlang")
# library(Rcpp)#; install.packages("Rcpp")
# devtools::install_github('thomasp85/gganimate')
library(magick)#;install.packages("magick")
# library(gganimate)#; install.packages("gganimate")
# install.packages('devtools')
library(ggridges)
# library(dplyr)
# library(malariaAtlas)#; install.packages("malariaAtlas")
```



```{r load_data }
# listData(datatype = "pr points")
# listData(datatype = "vector points")
# listData(datatype = "raster")
# listData(datatype = "shape")
# arg_pr_data <- getPR(country = "Argentina", species = "both")
# glimpse(arg_pr_data)

list.files(here("data","2018-11-13"))

malaria_data_deaths <- readr::read_csv(here("data","2018-11-13","malaria_deaths.csv"))%>% janitor::clean_names(case = "snake")
malaria_data_deaths_age <- readr::read_csv(here("data","2018-11-13","malaria_deaths_age.csv"))%>% janitor::clean_names(case = "snake")
malaria_data_inc <- readr::read_csv(here("data","2018-11-13","malaria_inc.csv")) %>% janitor::clean_names(case = "snake")

# skimr::skim(malaria_data_deaths)
# DataExplorer::create_report(malaria_data_deaths)
```

Lets explore the data:

```{r see_missing }
DataExplorer::plot_missing(malaria_data_deaths)
DataExplorer::plot_missing(malaria_data_deaths_age)
DataExplorer::plot_missing(malaria_data_inc)

```


in every missing report we see that there is several missing values related to code. 

checking what code means:

```{r check_missing }
malaria_data_inc %>% 
    filter(is.na(code)) %>% 
    count(entity) 
```

as we can see code is relatd to entities that are not countries, but aglomeration of them.


as seen that we have aglomerated countries we should try to see wich countries are OR.
try to plot ina tmap wha i already got :P

```{r tmap_experience }
data("World")
clean_malaria <- malaria_data_inc %>% filter(!is.na(code)) #%>% filter(year==2015)
world_2 <- World %>% left_join(clean_malaria,by=c("iso_a3"="code")) %>% 
    rename(malaria_risk=incidence_of_malaria_per_1_000_population_at_risk_per_1_000_population_at_risk)

# world_2 %>% glimpse()
all_years <- clean_malaria %>% count(year) %>% pull(year)
miss_country <- World %>%  anti_join(clean_malaria,by=c("iso_a3"="code")) %>% pull(iso_a3)

impute_miss <- lapply(miss_country,function(x){
    paste0(all_years,"|",x)
}) %>% unlist() %>% data.frame(stringsAsFactors = FALSE) %>% 
    rename(yr_cont=".") %>% as_tibble() %>% 
    separate(col = yr_cont,into=c("year","code"),sep = "\\|") %>% 
    mutate(year=as.integer(year),code=as.character(code)) %>% 
    mutate(malaria_risk=NA)

world_3 <- world_2 %>% left_join(impute_miss, by=c("iso_a3"="code")) %>%
    mutate(year=if_else(!is.na(year.x),year.x,year.y)) %>% 
    # mutate(malaria_risk=if_else(!is.na(malaria_risk.x),malaria_risk.x,malaria_risk.y)) %>% 
     mutate(malaria_risk=malaria_risk.x) %>% 
    select(-malaria_risk.x,-malaria_risk.y,-year.x,-year.y)

map_1 <- tmap::tm_shape(world_3) +
    tm_polygons("malaria_risk",n=25)+
    tm_facets(by = "year", nrow=1,ncol=1) +
    tm_legend(#legend.position = c("left", "top"),
      main.title = "Incidence of malaria ",
      # title="per 1000 population at risk",
      main.title.size=0.75,
      title.size=0.25,
      legend.text.size=0.25,
      legend.hist.size=0.25,
      # main.title.position = "right",
      legend.outside=TRUE,
      legend.show=FALSE,
      # title.snap.to.legend=FALSE,
      title.position = c("right","bottom"),
      legend.position = c("right","top"))

map_2 <- tmap::tm_shape(world_3) +
    tm_bubbles(size="malaria_risk",col = "malaria_risk")+
    tm_facets(by = "year", nrow=1,ncol=1) +
    tm_legend(#legend.position = c("left", "top"),
      main.title = "Incidence of malaria ",
      # title="per 1000 population at risk",
      main.title.size=0.75,
      # title.size=0.5,
      title.size=0.25,
      legend.text.size=0.25,
      legend.hist.size=0.25,
      # legend.title.size=0.25,
      # legend.title.size=0.25,
      # main.title.position = "right",
      # legend.outside=TRUE,
      # legend.show=FALSE,
      # title.snap.to.legend=FALSE,
      # title.position = c("right","bottom"),
      legend.position = c("right","bottom"))


# m3 <- tmap_arrange(map_1, map_2)

tmap_animation(map_1, filename="malaria_risk.gif", width=800, delay=100)
tmap_animation(map_2, filename="malaria_risk_bubbles.gif", width=800, delay=100)


```

```{r boxplot }
world_3 %>%
    filter(!is.na(malaria_risk)) %>%
    mutate(year=as.factor(year)) %>% 
    ggplot(aes(y = malaria_risk,x = year,fill=year))+
    geom_boxplot() +
    labs(x="", y="Malaria Risk" )+
    theme_light()+
    theme(legend.position = "none")+
    coord_flip()
```

as can see on 2000's got an extreme outlier,
also over the years malaria risk gone down.
```{r boxplot_outlier }
world_3 %>%
    as_tibble() %>% 
    filter(year==2000,malaria_risk> 1500) %>% 
    select(iso_a3,name,-geometry)
    
```

can see that the country on the outlier is Turkey.

# Communicate

#### technical stuff

this week saw tmap, 

- added data to the world map

- done choroplet
- done bubblemap+color
- some issues with legends & titles  ( sizes & positions )
- some issues when faceting and got NA's

- done gif's
- tried to go trhough gganimate, but it was not needed. 
- issues on gif: got to install packages and after that ImageMagic software, and restart rstudio


#### data stuff

- saw some NA's: they where related to aggregations of data but didnt have the exploded data.
- to solve this: imputed 0 to NA's , and added the years. for each of missing countries.
- one extreme outlier on 2000's , that was turkey.


