---
title: "week 2018-11-27"
author: "julio"
date: "2018-11-27"
output: html_document
---

# Tidy tuesday challenge: Week 2018-11-27

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

- baltimore bridges
- some animation to practice > cool stuff
- some map to practice > location
- some model to practice > predict 'ok'ness


## import data



```{r echo=FALSE,message=FALSE,warning=FALSE}
library(here)
library(readr)
library(purrr)
library(dplyr)
library(tidyr)
library(janitor)#install.packages('janitor')
library(scales)
library(ggplot2)
library(forcats)#install.packages('forcats')
library(lubridate)
library(visdat)#install.packages('visdat')
library(stringr)
library(DataExplorer)
library(rpart)
# library(cluster)
library(leaflet)
```

raw data:

```{r echo=FALSE,message=FALSE,warning=FALSE}
bridges_raw <- readr::read_csv(here::here("data","2018-11-27","baltimore_bridges.csv"))
glimpse(bridges_raw)
```

looks like vehicles and avg_dialy_traffic are the same. lets check

```{r echo=FALSE,message=FALSE,warning=FALSE}
bridges_raw %>% 
    mutate(vehicles_no_text=readr::parse_number(vehicles)) %>%
    mutate(eq_number=avg_daily_traffic==vehicles_no_text) %>% 
    count(eq_number)

```

yap. its replicated.

now lets take a look at owner / responsability

```{r echo=FALSE,message=FALSE,warning=FALSE}
bridges_raw %>% 
    mutate(eq_number=owner==responsibility) %>% 
    count(eq_number)

bridges_raw %>% 
    mutate(eq_number=owner==responsibility) %>% 
    count(owner,responsibility,eq_number) %>% 
    filter(eq_number==FALSE)

bridges_raw %>% 
    mutate(eq_number=owner==responsibility) %>% 
    count(owner,responsibility,eq_number) %>% 
    filter(is.na(eq_number))

```
don't happen the same, are a few with different number and some NA's

lets check NA generally:

```{r echo=FALSE,message=FALSE,warning=FALSE}
DataExplorer::plot_missing(bridges_raw)
```

after this see that responsability and owner are not that bad in NA's lets check which cases are:
```{r echo=FALSE,message=FALSE,warning=FALSE}
bridges_raw %>% filter(is.na(responsibility) | is.na(owner))

```

its weird as its quite new data. but just 2 registries.

about other missing, the total cost, it wont be taken in account in the analysis as 69% its NA 
```{r echo=FALSE,message=FALSE,warning=FALSE}

bridges_no_cost <- bridges_raw %>% select(-total_improve_cost_thousands,-vehicles)
```

about , not the same owner, responsability

hypo: this might get some quality issues , as interests might not be on the same page 


```{r echo=FALSE,message=FALSE,warning=FALSE}
bridges_no_cost %>% 
    filter(owner!=responsibility) %>% 
    select(owner,responsibility,bridge_condition,yr_built,avg_daily_traffic)

```
3 of them are not "good"
2 bad , 1 fair.


bridge_condition


```{r echo=FALSE,message=FALSE,warning=FALSE}
bridges_no_cost %>%  
    count(bridge_condition) %>% 
    mutate(bridge_condition=forcats::fct_reorder(bridge_condition,n)) %>% 
    ggplot(aes(bridge_condition,n, fill=bridge_condition)) +
    geom_col() +
    theme_light() +
    theme(legend.position = "none")+
    labs(title="Bridges by condition",
         x="",
         y="# bridges")+
    expand_limits(y = c(0,1500))+
    coord_flip()+
    geom_text(aes(label=n), hjust= -0.75) 
    

```

# some map

```{r echo=FALSE,message=FALSE,warning=FALSE}

bridges_no_cost_map <- bridges_no_cost %>% 
    mutate(bridge_color= case_when( 
        bridge_condition == "Fair" ~ '#000099',
        bridge_condition == "Good" ~ '#009900',
        bridge_condition == "Poor" ~ '#990000'
    )) 

    palette_fct <- colorFactor(
        palette = c("#990000","#009900","#000099"), 
        levels = c("Poor", "Good", "Fair"))

    # pretty colors but get lost on map     
    # bridges_no_cost_map <- bridges_no_cost %>% 
    # mutate(bridge_color= case_when( 
    #     bridge_condition == "Fair" ~ '#619CFF',
    #     bridge_condition == "Good" ~ '#00BA38',
    #     bridge_condition == "Poor" ~ '#F8766D'
    # )) 
    # 
    # palette_fct <- colorFactor(
    #     palette = c("#F8766D","#00BA38","#619CFF"), 
    #     levels = c("Poor", "Good", "Fair"))
    

    leaflet() %>%
    addProviderTiles(providers$OpenStreetMap) %>% 
    setView(lng= -76.6121893, lat=39.2903848, zoom=13) %>%
    addCircleMarkers(data=bridges_no_cost_map,
               radius = 2,
               color = ~bridge_color,
               popup = ~paste0(
                   "avg daily traffic: ",avg_daily_traffic,"<br/>",
                   "year built: ",yr_built,"<br/>",
                   "owner: ",owner,"<br/>",
                   "responsibility: ",responsibility,"<br/>",
                   "year built: ",yr_built,"<br/>",
                   "location: [",lat,";",long,"]")) %>% 
            addLegend(position = "bottomright",
              pal = palette_fct, 
              values = c("Poor", "Good", "Fair"))


```


```{r echo=FALSE,message=FALSE,warning=FALSE}

status_owner_perc <- bridges_no_cost_map %>% 
    count(bridge_condition,owner) %>% 
    rename(condition_owner_n=n) %>% 
    left_join(bridges_no_cost_map %>% count(owner),by=c("owner"="owner")) %>% 
    rename(total_owner_n=n) %>% 
    mutate(ratio_by_owner=condition_owner_n/total_owner_n) %>% 
    # arrange(desc(ratio_by_owner)) %>% 
    mutate(owner=forcats::fct_reorder(owner,condition_owner_n)) %>% 
    filter(!is.na(owner)) %>% 
    ggplot(aes(owner,ratio_by_owner,fill=bridge_condition)) +
    geom_col()+
    theme_light()+
    scale_fill_manual(values = c("Poor" = "#F8766D", "Good" = "#00BA38", "Fair" = "#619CFF"))+
    labs(title = "Bridge status by owner",
         x="",
         y="% by owner",
         fill="Bridge condition")+
    scale_y_continuous(labels = scales::percent_format())+
    coord_flip()
# c("Poor" = "#F8766D", "Good" = "#00BA38", "Fair" = "#619CFF")
# c("A" = "green", "B" = "blue", "C" = "red", "D" = "grey")
    
# show_col(hue_pal()(3))
#R #F8766D
#G #00BA38
#B #619CFF

status_owner_n <- bridges_no_cost_map %>% 
    count(bridge_condition,owner) %>% 
    rename(condition_owner_n=n) %>% 
    left_join(bridges_no_cost_map %>% count(owner),by=c("owner"="owner")) %>% 
    rename(total_owner_n=n) %>% 
    mutate(ratio_by_owner=condition_owner_n/total_owner_n) %>% 
    mutate(owner=forcats::fct_reorder(owner,condition_owner_n)) %>% 
    filter(!is.na(owner)) %>% 
    ggplot(aes(owner,condition_owner_n,fill=bridge_condition)) +
    geom_col()+
    # geom_text(aes(label=total_owner_n), hjust= -0.75) +
    theme_light()+
    scale_fill_manual(values = c("Poor" = "#F8766D", "Good" = "#00BA38", "Fair" = "#619CFF"))+
    labs(title = "Bridge status by owner",
         x="",
         y="# by owner",
         fill="Bridge condition")+
    scale_y_continuous(labels = scales::number_format())+
    coord_flip()
    
cowplot::plot_grid(status_owner_perc,status_owner_n)


```


Try some models for explaining the state of the bridge

```{r echo=FALSE,message=FALSE,warning=FALSE}

# model_tree <- bridges_no_cost %>%   rpart::rpart(bridge_condition ~ ., data = .)

# plot(model_tree)

 # model_year_rules <- rpart.rules(model_year, cover = TRUE) %>% View() 

# glm(bridge_condition ~ ., data = donors, family = "binomial")

# need to do a multi class

# but before that lets see if can cluster in N
# kmeans(bridges_no_cost %>% filter(complete.cases(bridges_no_cost)),centers = 3)

# bridges_no_cost %>% filter(!complete.cases(bridges_no_cost))

# bridges_no_cost_dummy <- dist(bridges_no_cost %>% filter(complete.cases(bridges_no_cost)),method='binary')
# head(bridges_no_cost_dummy)
# cluster_dummy <- hclust(bridges_no_cost_dummy,method = "complete")
# cluster_h <- cutree(cluster_dummy,k = 3)

# bridges_no_cost <- bridges_no_cost %>% mutate(clsuter_h=cluster_h)

```

# communicate

- focused on bridge condition
- overall 
- by owner as # and %
- also some map practice on leaflet that was really handy :D, should check course deeply ( https://www.datacamp.com/courses/interactive-maps-with-leaflet-in-r )

- well i'll leave models for other time, im happy that got the feel of "should try to explain bridge condition on other variables" but well too much time invested :P . a nice place to start will be drob's screencast: https://www.youtube.com/watch?v=kzM-4jMh9Qs




