---
title: "Tidy tuesday challenge: Week 2019-09-10 Amusement Park Injuries
author: "julio"
date: "2019-09-08"
output: html_document
---

# Tidy tuesday challenge: Week 2019-09-10 Amusement Park Injuries

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

just explore and plot something =) 

### Data:
 
 Week Week 2019-09-10 Amusement Park Injuries
 
### objectives:

explore & do something

## details:


## import data
```{r echo=FALSE,message=FALSE,warning=FALSE}
library(magrittr) # para el %T>%
library(tidyverse)#; install.packages("tidyverse")
# library(tidyr)#;install.packages("tidyr");install.packages("readxl");install.packages("xml2")
# library(sf)
library(dplyr)
library(stringr)#;
# library(rebus)#; install.packages('rebus')
# library(tidytext)
# library(prophet)
# install.packages('scico')
library(scico)

# install.packages("Rcpp")
# remotes::install_github("tylermorganwall/rayshader")
# library(rayshader)
library(lubridate)
library(ggforce)

library(corrplot)

library(ggrepel)
library(tidytext)
library(tidyquant)
# library(arules)#install.packages('arules')
# library(arulesViz)#install.packages('arulesViz')
# library(igraph)

```


```{r echo=FALSE,message=FALSE,warning=FALSE}
tx_injuries <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-09-10/tx_injuries.csv")

safer_parks <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-09-10/saferparks.csv")

# now_data <- list(tx_injuries,safer_parks)

```

# explore data

```{r echo=FALSE,message=FALSE,warning=FALSE}
# now_data %>% purrr::map(now_data,head)
tx_injuries %>% head()
safer_parks %>% head()

```

```{r echo=FALSE,message=FALSE,warning=FALSE}

tx_injuries %>% glimpse()
safer_parks %>% glimpse()


```

```{r echo=FALSE,message=FALSE,warning=FALSE}

tx_injuries %>% skimr::skim()
safer_parks %>% skimr::skim()

```

```{r echo=FALSE,message=FALSE,warning=FALSE}
DataExplorer::plot_missing(tx_injuries)
DataExplorer::plot_missing(safer_parks)
```

getting no missing columns

```{r echo=FALSE,message=FALSE,warning=FALSE}
safer_parks_no_na <- safer_parks %>% select(category,tradename_or_generic,device_type,device_category,industry_sector,bus_type,source,fix_port,acc_state,acc_date,acc_id)

tx_injuries_no_na <- tx_injuries %>% select(ride_name,st,city,name_of_operation,injury_report_rec)
```

ok . now we see that reports, are mainly from TX 51 reports from TX, 1 from FL & 1 from AZ


```{r echo=FALSE,message=FALSE,warning=FALSE}
tx_injuries_no_na %>% 
    count(st,injury_report_rec) %>% 
    count(st)
```




what about safer parks data

```{r echo=FALSE,message=FALSE,warning=FALSE}
p <- safer_parks_no_na %>% 
    count(acc_state,sort = TRUE) %>% 
    mutate(acc_state=fct_reorder(acc_state,n)) %>% 
    filter(n>0) %>% 
    ggplot(aes(x=acc_state,y=n,fill=acc_state))+
    geom_col()+
    coord_flip()+
    theme(legend.position = "none")

plotly::ggplotly(p)

```

```{r echo=FALSE,message=FALSE,warning=FALSE}
pf_data <-  safer_parks_no_na %>% 
    mutate(date_acc=lubridate::mdy(acc_date)) %>% 
    mutate(year=as.factor(lubridate::year(date_acc)))


pf_data_min <- min(pf_data$date_acc)
pf_data_max <- max(pf_data$date_acc)

pf <- pf_data %>% 
    count(acc_state,year,sort = TRUE) %>% 
    mutate(acc_state=fct_reorder(acc_state,n)) %>% 
    filter(n>0) %>% 
    ggplot(aes(x=year,y=acc_state,fill=n))+
    # geom_col()+
    geom_tile()+
    scale_fill_scico(palette = "roma",direction = -1)+
    scale_x_discrete()+
    theme_light()+
    labs(x="",y="",fill="#accidents",
         title="Amusement Park Accidents In USA",
         subtitle = paste0("from ",pf_data_min, " to ",pf_data_max),
         caption="#TidyTuesday")
    # ggridges::geom_density_ridges()
    # coord_flip()
    # theme(legend.position = "none")
ggsave(filename = "accidents_by_year.png",plot = pf,width = 5,height = 7)
# plotly::ggplotly(p)

```


## items as absolute values



#tweet: 

2019-09-10 #TidyTuesday #rstats Amusement Park Injuries In USA !
Some simple plot with scico package for colors. 
Also with plumber api to test different palettes as parameter :D!
code: https://github.com/jas1/tidytuesday/tree/master/jas1_weeks/2019/2019-09-10


# communicate

plumber api with plots

here tweet: https://twitter.com/jspairani/status/1173799069761847296

# resources: 

scico palette: https://github.com/thomasp85/scico
code plumber: https://github.com/rstudio/plumber/blob/master/R/images.R
docu plumber: https://www.rplumber.io/docs/rendering-and-output.html#response-object
some stack overflow: https://stackoverflow.com/questions/50033857/serve-arbitrary-image-files-through-plumber

spanish version: 
- https://github.com/jas1/datos-de-miercoles/blob/master/jas1_semanas/2019/2019-08-21/api_co2/plumber.R
