

---
title: "Tidy tuesday challenge: Week 2019-08-13 Roman Emperors"
author: "julio"
date: "2019-08-04"
output: html_document
---

# Tidy tuesday challenge:  Week 2019-08-13 Roman Emperors

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

### Data:
 
 Week 2019-08-06 Bob Ross
 
### objectives:



## details:


## import data
```{r echo=FALSE,message=FALSE,warning=FALSE}
library(magrittr) # para el %T>%
library(tidyverse)
# library(sf)
library(dplyr)
library(stringr)#;
# library(rebus)#; install.packages('rebus')
# library(tidytext)
library(prophet)


# install.packages("Rcpp")
# remotes::install_github("tylermorganwall/rayshader")
# library(rayshader)
library(lubridate)
library(ggforce)
library(ggrepel)

# library(arules)#install.packages('arules')
# library(arulesViz)#install.packages('arulesViz')
# library(igraph)

```


```{r echo=FALSE,message=FALSE,warning=FALSE}
emperors <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-08-13/emperors.csv")
```

# explore data


```{r echo=FALSE,message=FALSE,warning=FALSE}
emperors %>% head()

```

```{r echo=FALSE,message=FALSE,warning=FALSE}
emperors %>% glimpse()

```

```{r echo=FALSE,message=FALSE,warning=FALSE}
emperors %>% skimr::skim()


```

we see some emperors got missing Day of birth.
```{r echo=FALSE,message=FALSE,warning=FALSE}
emperors %>% filter(is.na(birth))
```

ill search & impute: 

imputing march 3 for not know date 

Florian: August 19, 232 AD, Terni, Italy >> 19/08/232 ( wikipedia )
Numerian: Born: 254 AD >> 3/3/254 ( wikipedia + impute )
Carinus: Born: 257 AD, Dalmatia, Croatia >> 3/3/257 ( wikipedia + impute )
Severus II (Valerius Severus) :  around the middle of the third century AD >> 3/3/250 ( wikipedia + impute )
Vetranio: sometime in the late 3rd century >> 3/3/290 (wikipedia + impute)

imputing values and added imputed col

```{r echo=FALSE,message=FALSE,warning=FALSE}

emperors_processed <- emperors %>% 
    mutate(is_birth_imputed=is.na(birth)) %>% 
    mutate(birth_imputed=case_when(
        index==44 ~ { #Florian: August 19, 232 AD, Terni, Italy >> 19/08/232 ( wikipedia )
        lubridate::dmy("19/08/0232")
            },
        index==47 ~ { # Numerian: Born: 254 AD >> 3/3/254 ( wikipedia + impute )
            lubridate::dmy("03/03/0254")
        },
        index==48 ~ { #Carinus: Born: 257 AD, Dalmatia, Croatia >> 03/03/0257 ( wikipedia + impute )
            lubridate::dmy("03/03/0257")
        },
        index==53 ~ { #Severus II (Valerius Severus) :  around the middle of the third century AD >> 03/03/250 ( wikipedia + impute )
            lubridate::dmy("03/03/0250")
        },
        index==61 ~ { #Vetranio: sometime in the late 3rd century >> 3/3/290 (wikipedia + impute)
            lubridate::dmy("03/03/0290")
        },
        TRUE ~ birth
    )) %>% 
    mutate(impute_desc=case_when(
        index==44 ~ { "19/08/232\n( wikipedia )" },
        index==47 ~ { "Born: 254 AD\nimputed: 03/03/0254\n( wikipedia + impute )" },
        index==48 ~ { "Born: 257 AD\nimputed: 03/03/0257\n( wikipedia + impute )"},
        index==53 ~ { "around the middle of the third century AD\nimputed: 03/03/0250\n( wikipedia + impute )"},
        index==61 ~ { "sometime in the late 3rd century\nimputed: 03/03/0290\n(wikipedia + impute)"},
        TRUE ~ "--"
    ))

emperors_processed %>% select(name,birth,birth_imputed,is_birth_imputed,impute_desc) %>% 
    filter(is.na(birth))
```




```{r echo=FALSE,message=FALSE,warning=FALSE}
gathered_emperors <- emperors_processed %>% 
    select(birth_imputed,death,is_birth_imputed,impute_desc,reign_start,reign_end,
           name_full,name,dynasty,era) %>% 
    mutate(dynasty=factor(name)) %>% 
    mutate(dynasty=factor(dynasty)) %>% 
    gather(date_type,date,-name_full,-name,-dynasty,-era,-is_birth_imputed,-impute_desc) %>% 
    mutate(date_span=if_else(date_type %in% c("birth_imputed","death"), "life","reign" ))

gathered_emperors %>% glimpse()
```

```{r echo=FALSE,message=FALSE,warning=FALSE}

dynasty_geom_df <- emperors_processed %>% group_by(dynasty) %>% 
    summarise(min_date=min(reign_start),max_date=max(reign_end),
              first=head(name,1),last=tail(name,1),date_span="",name="")

era_geom_df <- emperors_processed %>% group_by(era) %>% 
    summarise(min_date=min(reign_start),max_date=max(reign_end),
              first=head(name,1),last=tail(name,1),date_span="",name="")


# geom_rect(xmin=dynasty_geom_df[1,]$min_date,
#           xmax=dynasty_geom_df[1,]$max_date,
#           ymin=dynasty_geom_df[1,]$first, ymax=dynasty_geom_df[1,]$last)+
# geom_rect(xmin=dynasty_geom_df$min_date, xmax=dynasty_geom_df$max_date,
#           ymin=dynasty_geom_df$first, ymax=dynasty_geom_df$last,
#           fill=dynasty_geom_df$dynasty)+
# scale_y_date(date_breaks = "1 year")+
plot_emperor_life <- gathered_emperors %>% 
    mutate(name=fct_reorder(name,date)) %>% 
    ggplot(aes(x=name,y=date,fill=date_span,colour=date_span)) + 
    geom_line()+
    geom_label_repel(data = gathered_emperors %>%
                        filter(is_birth_imputed) %>%
                        filter(date_type=="birth_imputed"),
                    aes(label=paste0(name,": ",impute_desc)),
                    nudge_y= -500000,
                    color="#333333",
                    fill="#ffffff",
                    size=2.2)+

    coord_flip()+
    
    labs(x="",y="",color="time span",
         title = "Roman Emperors time span",
         caption="#TidyTuesday")+
    theme_light()
    
    
    # geom_col(data=gathered_emperors %>% filter(date_type %in% birth,date))

ggsave(plot_emperor_life,filename = "plot_emperor_life.png",height = 9,width = 10)

plot_emperor_life
```




```{r echo=FALSE,message=FALSE,warning=FALSE}
plotly::ggplotly(plot_emperor_life)
```

## items as absolute values



#tweet: 

2019-08-06 #TidyTuesday #rstats Bob Ross ! 
Association rules, wanted to see how the elements were related.
some element freq/counts, then arules and a plus, @othomn, @chisatini voronoi on images; just wanted to give a try, and its amazing :D! thanks ! 
https://github.com/jas1/tidytuesday/tree/master/jas1_weeks/2019/2019-08-06



# communicate

Arules, graphs & voronois

## resources: 

Voronoi stuff: 

- seen on: https://twitter.com/othomn/status/1159110356867911680

- https://github.com/othomantegazza/code-tidytuesday/blob/master/2-32-painting-voronoi.R
- https://chichacha.netlify.com/2018/11/12/utilizing-k-means-to-extract-colours-from-your-favourite-images/

Arules/Apriori / Association rules: based on previous work and other references: 

- https://twitter.com/jspairani/status/1083717747173928960
- https://michael.hahsler.net/research/arules_RUG_2015/demo/#association-rules
- https://www.datacamp.com/community/tutorials/market-basket-analysis-r
- https://stackoverflow.com/questions/44677670/exporting-a-network-produced-with-visnetwork-for-r


