

---
title: "Tidy tuesday challenge: Week 2019-08-27 Simpsons guests
author: "julio"
date: "2019-08-20"
output: html_document
---

# Tidy tuesday challenge:  Week 2019-08-27 Simpsons guests

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

just explore and plot something =) 

### Data:
 
 Week 2019-08-27 Simpsons guests
 
### objectives:

explore & do something

## details:


## import data
```{r echo=FALSE,message=FALSE,warning=FALSE}
library(magrittr) # para el %T>%
library(tidyverse)
# library(sf)
library(dplyr)
library(stringr)#;
# library(rebus)#; install.packages('rebus')
# library(tidytext)
# library(prophet)


# install.packages("Rcpp")
# remotes::install_github("tylermorganwall/rayshader")
# library(rayshader)
library(lubridate)
library(ggforce)
library(ggrepel)

# library(arules)#install.packages('arules')
# library(arulesViz)#install.packages('arulesViz')
# library(igraph)

```


```{r echo=FALSE,message=FALSE,warning=FALSE}
simpsons <- readr::read_delim("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-08-27/simpsons-guests.csv", delim = "|", quote = "")
```

# explore data


```{r echo=FALSE,message=FALSE,warning=FALSE}
simpsons %>% head()

```

```{r echo=FALSE,message=FALSE,warning=FALSE}
simpsons %>% glimpse()

```

```{r echo=FALSE,message=FALSE,warning=FALSE}
simpsons %>% skimr::skim()


```


guest stars as themselves vs guest stars as other


```{r echo=FALSE,message=FALSE,warning=FALSE}
simpsons_process <- simpsons %>% 
    separate_rows(role,sep = ";") %>%  # becaus of multiple roles
    mutate(role=stringr::str_trim(role)) %>% # becaus of leadign or ending withe spaces
    mutate(as_self=if_else(role%in%c("Himself","Herself","Themselves"),TRUE,FALSE)) %>% 
    mutate(season=if_else(season=="Movie",as.double(18.5),as.double(season)))
    # mutate(as_self=if_else(role==guest_star,TRUE,FALSE)) 
```

```{r echo=FALSE,message=FALSE,warning=FALSE}
total_season <- simpsons_process %>%
    count(season) %>%  rename(total=n)

are_themselves_plot <- simpsons_process %>%
    group_by(season,as_self) %>% 
    left_join(total_season) %>% 
    summarise(quant=n(),total=mean(total) ,ratio=quant/total) %>% 
    # ungroup() %>% 
    # mutate(season=as.factor(season)) %>% 
    # mutate(season=if_else(season=="0","Movie",season)) %>% 
    ggplot(aes(x=as.factor(season),y=ratio,fill=as_self)) +
    geom_col()+
    scale_y_continuous(labels=scales::percent)+
    theme_light()+
    coord_flip()+
    labs(x="",y="# count",fill="as self",
         title="Simpsons Guests: Are themselves across the seasons ?",
         subtitle = "Movie recoded as 18.5 ( released on july 2007 between season 18 and 19)",
         caption="#TidyTuesday")

ggsave(plot=are_themselves_plot,filename = "are_themselves_plot.png")

# 
# simpsons_process %>% 
#     count(role,sort = TRUE)
```

```{r echo=FALSE,message=FALSE,warning=FALSE}

more_than_one_role_data <- simpsons_process %>%
    count(season,guest_star,role,sort = TRUE) %>% 
    count(season,guest_star,sort = TRUE) 

metric_str <- "n"
top_n_num <- 3
resu <- more_than_one_role_data %>% 
nest(-season) %>% 
    mutate(top_n_year=purrr::map(data,
                                    n_top=top_n_num,
                                    .f=function(d,n_top){
        ranking <- seq(from=1,to = n_top,by = 1)
        ret <- d %>% 

            arrange(desc(n)) %>% 
            head(n_top) %>% 
            mutate(rank=ranking)
        ret
    })) %>% unnest(top_n_year)

 plot_out <- resu %>% 
    mutate(guest_star=fct_reorder(guest_star,rank,.desc = TRUE)) %>% 
    mutate(rank=as.factor(rank)) %>% 
    ggplot(aes(x=as.factor(season),y=guest_star,color=rank,group=rank,label=n))+
    geom_point()+
    geom_line(show.legend = FALSE)+
    theme_light()+
    theme(axis.text.x = element_text(angle = 90, hjust = 1),
          axis.text.y = element_text(size = 6))+
    labs(x="",y="", 
         title="Top 3 guests with more characters per season",
         subtitle = "Movie recoded as 18.5 ( released on july 2007 between season 18 and 19)",
         caption="#TidyTuesday")

ggsave(plot=plot_out,filename = "top3_guest_chars.png")


```

## items as absolute values



#tweet: 

2019-08-27 #TidyTuesday #rstats Simpsons guests !
they play themselves ?
who are the top 3 playing more than one role over the seasons?



# communicate

just a simple one

## resources: 

old submissions :p 

## on demand palette

```{r echo=FALSE,message=FALSE,warning=FALSE}

total_season <- simpsons_process %>%
    count(season) %>%  rename(total=n)

are_themselves_plot_palette <- simpsons_process %>%
    group_by(season,as_self) %>% 
    left_join(total_season) %>% 
    summarise(quant=n(),total=mean(total) ,ratio=quant/total) %>% 
    # ungroup() %>% 
    # mutate(season=as.factor(season)) %>% 
    # mutate(season=if_else(season=="0","Movie",season)) %>% 
    ggplot(aes(x=as.factor(season),y=ratio,fill=as_self)) +
    geom_col()+
    scale_y_continuous(labels=scales::percent)+
    coord_flip()+
    theme_light()+
    scale_fill_manual(values = c("#dedede","#ffbb22"))+#"#8899ff"
    theme(text = element_text(colour = "#eeaa11"),#aaaaaa
          panel.grid=element_line(colour="#99aaff"),
          plot.background = element_rect(fill ="#eeeeee"),##8899ff
          panel.background = element_rect(fill ="#8899ff" ) )+
    geom_hline(yintercept = 0.25,linetype=2,colour="#99aaff" )+
    geom_hline(yintercept = 0.5,linetype=2,colour="#99aaff" )+
    geom_hline(yintercept = 0.75,linetype=2,colour="#99aaff" )+
    labs(x="",y="# count",fill="as self",
         title="Simpsons Guests: Are themselves across the seasons ?",
         subtitle = "Movie recoded as 18.5 ( released on july 2007 between season 18 and 19)",
         caption="#TidyTuesday")
are_themselves_plot_palette
ggsave(plot=are_themselves_plot_palette,filename = "are_themselves_plot_palette.png")

# 
# simpsons_process %>% 
#     count(role,sort = TRUE)
```
