---
title: "Tidy tuesday challenge: Week 2019-10-01 Pizza!
author: "julio"
date: "2019-10-07"
output: html_document
---

# Tidy tuesday challenge: Week 2019-10-01 Pizza !

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

just explore and plot something =) 

### Data:
 
 Week Week 2019-10-01 Pizza!
 
### objectives:

explore & do something

## details:


## import data
```{r echo=FALSE,message=FALSE,warning=FALSE}
# library(magrittr)
library(tidyverse)
# library(ggrepel)
library(skimr)

library(here)

library(choroplethr)#;install.packages("choroplethr")
library(choroplethrMaps)#;install.packages("choroplethrMaps") # ??choroplethrMaps

# library(dplyr)
# library(tidyr)
# library(readxl)
# library(janitor)
# library(stringr)
# library(lubridate)
# library(geojsonsf)

# library(geojsonio); #install.packages("geojsonio")

# library(grid)
# library(spatialEco)#; install.packages("spatialEco")
# library(sp)
# library(sf)# install.packages('sf')
# install.packages("Rcpp")
# remotes::install_github("tylermorganwall/rayshader")
# library(rayshader)

# library(transformr)# install.packages('transformr')
# library(gganimate)# install.packag
```


```{r echo=FALSE,message=FALSE,warning=FALSE}
pizza_jared <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-10-01/pizza_jared.csv")
pizza_barstool <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-10-01/pizza_barstool.csv")
pizza_datafiniti <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-10-01/pizza_datafiniti.csv")

```

too much data i'll just focus on parks

# explore data

```{r echo=FALSE,message=FALSE,warning=FALSE}
# now_data %>% purrr::map(now_data,head)
pizza_datafiniti %>% head()


```

```{r echo=FALSE,message=FALSE,warning=FALSE}

pizza_datafiniti %>% glimpse()



```

```{r echo=FALSE,message=FALSE,warning=FALSE}

pizza_datafiniti %>% skimr::skim()


```

```{r echo=FALSE,message=FALSE,warning=FALSE}
DataExplorer::plot_missing(pizza_datafiniti)

```

going for datafinity


```{r echo=FALSE,message=FALSE,warning=FALSE}

amount_of_pizza_places_by_state <- pizza_datafiniti %>% 
    mutate(price_range=price_range_max-price_range_min) %>% 
    group_by(province) %>% 
    summarise(price_range_min_median=mean(price_range_min),
              price_range_max_median=max(price_range_max),
              price_range_var_median=median(price_range),
              amount=n())

```


```{r echo=FALSE,message=FALSE,warning=FALSE}
# data(continental_us_states)
# ggmap::

us_states <- data.frame(state.name,state.abb,stringsAsFactors = FALSE) %>% 
    as_tibble() %>% 
    mutate(state_name=str_to_lower(state.name))


simple_choro <- amount_of_pizza_places_by_state %>%
    left_join(us_states,by=c("province"="state.abb")) %>% 
    rename(region=state_name,
           value=amount) %>%
    state_choropleth(num_colors=7,
                 zoom = continental_us_states)+
    scale_fill_brewer(palette="YlOrBr")+
    labs(title="Pizza places in Datafinity",
         fill="# of places",
         caption="#TidyTuesday")

ggsave("simple_choro.png",simple_choro,width = 7,height = 5)

```

#tweet: 

2019-10-01 #TidyTuesday #rstats Pizza places !
amount of pizza places in datafinity DB by state.

code: https://github.com/jas1/tidytuesday/tree/master/jas1_weeks/2019/2019-10-01


# communicate

here tweet: https://twitter.com/jspairani/status/1175980540576620544

