---
title: "Tidy tuesday challenge: Week 2019-03-19 The Stanford Open Policing Project"
author: "julio"
date: "24-03-2019"
output: html_document
---

# Tidy tuesday challenge: Week 2019-03-19 The Stanford Open Policing Project

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

### Data:

this week data its related to The Stanford Open Policing Project

### objectives:

not clear, lets explore data and some stuff...

## details:


## import data

```{r echo=FALSE,message=FALSE,warning=FALSE}
library(dplyr)
library(tidyr)
library(skimr)
library(readr) 
library(ggplot2)
library(forcats)
library(usmap)#;install.packages('usmap') # https://cran.r-project.org/web/packages/usmap/vignettes/mapping.html


# library(tidyverse)
library(ggimage)#; install.packages('ggimage') 
library(sf)#; install.packages('ggimage') 

```



```{r echo=FALSE,message=FALSE,warning=FALSE}
combined_data <- readr::read_csv("https://raw.githubusercontent.com/5harad/openpolicing/master/results/data_for_figures/combined_data.csv")

head(combined_data)
glimpse(combined_data)
skimr::skim(combined_data)

# head(earnings_female)
# glimpse(earnings_female)
# skimr::skim(earnings_female)
 
# head(employed_gender)
# glimpse(employed_gender)
# skimr::skim(employed_gender)
```

after taking a look at the dataset, ill choose the non missing columns to work.

```{r echo=FALSE,message=FALSE,warning=FALSE}
combined_data <- readr::read_csv("https://raw.githubusercontent.com/5harad/openpolicing/master/results/data_for_figures/combined_data.csv")

combined_processed <- combined_data %>% select(driver_race,location,state,stops_per_year)

combined_processed_by_state <- combined_processed %>% 
    group_by(driver_race,state) %>% 
    summarise(state_stops_per_year = sum(stops_per_year),
              avg_state_stops_year = mean(stops_per_year),
              min_state_stops_year = min(stops_per_year),
              max_state_stops_year = max(stops_per_year),
              median_state_stops_year = median(stops_per_year),
              iqr_state_stops_year = max(stops_per_year))
# library(devtools)
# library(treemapify) # install_github("wilkox/treemapify")
# library(ggplotify) # install.packages('ggplotify')

# if(!require(treemapify)){install_github("wilkox/treemapify")}
# treemap_coords <- treemapify(combined_processed_by_state,
#                              area="state_stops_per_year", 
#                              fill="HDI", 
#                              label="driver_race", 
#                              group="state")
# head(treemap_coords)

# library
library(treemap)#; install.packages("treemap")
library(d3treeR)#;install.packages("d3treeR")
# Build Dataset
# group=c(rep("group-1",4),rep("group-2",2),rep("group-3",3))
# subgroup=paste("subgroup" , c(1,2,3,4,1,2,1,2,3), sep="-")
# value=c(13,5,22,12,11,7,3,1,23)
# data=data.frame(group,subgroup,value)

# treemap
treemap(combined_processed_by_state,
            index=c("state","driver_race"),
            vSize="state_stops_per_year",
            type="index")

+
    labs(title="Stops per year by State and Driver race")

# make it interactive ("rootname" becomes the title of the plot):
# inter=
    # d3tree2( tree_map ,  rootname = "General" )

```


```{r echo=FALSE,message=FALSE,warning=FALSE}
# 
# plot_usmap(data = combined_processed_by_state, 
#            values = "state_stops_per_year", 
#            lines = "red") + 
#   scale_fill_continuous( low = "white", high = "red", name = "Stops per year", label = scales::comma ) +
#   theme(legend.position = "right")
# 
# 
# combined_processed_by_county <- combined_processed %>% 
#     mutate(state_location=paste0(state,'_',location)) %>% 
#     group_by(state_location) %>% 
#     summarise(state_stops_per_year = sum(stops_per_year),
#               avg_state_stops_year = mean(stops_per_year),
#               min_state_stops_year = min(stops_per_year),
#               max_state_stops_year = max(stops_per_year),
#               median_state_stops_year = median(stops_per_year),
#               iqr_state_stops_year = max(stops_per_year))
# 
# # plot_usmap(data = combined_processed_by_county, 
# #            values = "state_stops_per_year", 
# #            lines = "black",regions = "counties") + 
# #   scale_fill_continuous( low = "white", high = "red", 
# #                          name = "Stops per year", 
# #                          label = scales::comma ) +
# #   theme(legend.position = "right")
# 
# 
# 
# 
# combined_processed_by_county %>% 
#     mutate(location = fct_reorder(location,state_stops_per_year)) %>% 
#     arrange(desc(state_stops_per_year)) %>% 
#     head(20) %>% 
#     ggplot(aes(x=location,
#                y=state_stops_per_year,
#                fill=state_stops_per_year)) +
#     geom_col() +
#     coord_flip()
# 
# combined_processed_by_county %>% 
#     ungroup() %>% 
#     mutate(state=fct_reorder(state,avg_state_stops_year)) %>% 
#     mutate(location = fct_reorder(location,avg_state_stops_year)) %>% 
#     arrange(desc(avg_state_stops_year)) %>% 
#     head(20) %>% 
#   ggplot(aes(x=location,
#                y=avg_state_stops_year,
#                fill=avg_state_stops_year)) +
#     geom_col() +
#     coord_flip()

```

### selected columns

looking at the columns can spot , race, location, state, and stopes per year. 
well lets, do a cloroplet with the new ggplot stuff.


```{r echo=FALSE,message=FALSE,warning=FALSE}
# combined_data <- readr::read_csv("https://raw.githubusercontent.com/5harad/openpolicing/master/results/data_for_figures/combined_data.csv")
# 
# combined_processed <- combined_data %>% select(driver_race,location,state,stops_per_year)
# 
# 
# combined_processed %>% 
# 
# 
# # head(earnings_female)
# # glimpse(earnings_female)
# # skimr::skim(earnings_female)
#  
# # head(employed_gender)
# # glimpse(employed_gender)
# # skimr::skim(employed_gender)
# 
# # combined_processed_2 <- combined_processed %>% 
# #     gather_set_data(1:4)
# # head(titanic)
# # 
# # 
# # ggplot(combined_processed, aes(x, id = id, split = y, value = value)) +
# #   geom_parallel_sets(aes(fill = Sex), alpha = 0.3, axis.width = 0.1) 
# 


```



## communicate


### general view:


### summary: 

### full story:


### to close ...:


### useful resources:

- https://twitter.com/frizzyroselle/status/1104042846645043200
- https://ogustavo.com/post/dotplot-ggplot/
