---
title: "Tidy tuesday challenge: Week 2019-03-05 Women in the Workforce"
author: "julio"
date: "10-03-2019"
output: html_document
---

# Tidy tuesday challenge: Week 2019-03-05 Women in the Workforce

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

### Data:

this week data its related to Women in the Workforce

### objectives:

- just to do something similar to https://ogustavo.com/post/dotplot-ggplot/
as i never done one before :P. 


## details:


## import data

```{r echo=FALSE,message=FALSE,warning=FALSE}
library(dplyr)
library(tidyr)
library(skimr)
library(readr) 
library(ggplot2)
library(forcats)


```



```{r echo=FALSE,message=FALSE,warning=FALSE}
jobs_gender <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-03-05/jobs_gender.csv")
# earnings_female <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-03-05/earnings_female.csv") 
# employed_gender <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-03-05/employed_gender.csv") 

head(jobs_gender)
glimpse(jobs_gender)
skimr::skim(jobs_gender)

# head(earnings_female)
# glimpse(earnings_female)
# skimr::skim(earnings_female)
# 
# head(employed_gender)
# glimpse(employed_gender)
# skimr::skim(employed_gender)

```

after peeking the 3 datasets, i'll go for jobs_gender.

### jobs_gender

looking at how can i subset the data to get this working: 
pointing to minor categories. 

then a year
then male / female
then the difference, 

## tidy

lets make the data so i can make the plot

- category + value 1 and value 2

category = minor categories
value 1= workers_female
value 2= workers_male
year = to have the time dimension.

as they dont have missings. 



```{r echo=FALSE,message=FALSE,warning=FALSE}
jobs_gender_processed <- jobs_gender %>% 
    group_by(year,minor_category) %>%  # group
    summarise(workers_female_total = sum(workers_female),
              workers_male_total = sum(workers_male),
              workers_total_total = sum(total_workers)) %>%  # sumarize
    mutate(diff=workers_male_total-workers_female_total) %>% # diff
    mutate(percent_female=workers_female_total/workers_total_total) %>% # get percents
    mutate(percent_male=workers_male_total/workers_total_total) %>%
    mutate(perc_diff=percent_male-percent_female) 
    
# may be filter year and sort by difference

jobs_gender_processed_gather <- jobs_gender_processed %>%  
    filter(year==2016) %>% ungroup() %>% 
    select(minor_category,percent_male,percent_female,perc_diff) %>% 
    rename(male_workers=percent_male,female_workers=percent_female) %>% 
    gather(
    key = gender,
    value = workers_percent,
    male_workers, 
    female_workers
  ) %>% 
    select(minor_category,gender,workers_percent,perc_diff)
    
jobs_gender_processed_gather_2 <- jobs_gender_processed_gather %>% 
    mutate(minor_category=forcats::fct_reorder(minor_category,perc_diff))    

lev <- levels(jobs_gender_processed_gather_2$minor_category)
first_entry <- lev[length(lev)]



```

## visualize



```{r echo=FALSE,message=FALSE,warning=FALSE}

col_orange <- "#FF856B"
col_orange_text <- "#F0B9AD"

col_green_text <- "#16BFB1"
col_green <- "#3196C1"
col_background_diff <- "#F5EBCB"
col_gray <- "#bbbbbb"
col_dark_gray <- "#333333"
col_light_gray <- "#f4f4f4"
nudge_y_titles <- 0.6
font_size_titles <- 3
# percent_nudge_x <- 0.03
percent_nudge_x <- 0
percent_nudge_y <- -0.5
title_plot <- "Difference between genders by\namount of workers in\nminor working category"

p <- jobs_gender_processed_gather_2 %>% 
ggplot(aes( y=minor_category, x=workers_percent, 
    color = ifelse(perc_diff == 0, "zero", gender))) + 
  # Plot lines between points, by Country
  geom_line(mapping = aes(group = minor_category),color = col_gray,size = 1.5) +
  geom_point( size=3, pch = 19 ) + # dot type. Important to be number 19, otherwise we cannot plot the dots with the colors for different groups
  # Add % for each point
  geom_text( size = 3, colour = col_dark_gray,
             nudge_x = percent_nudge_x,
             nudge_y = percent_nudge_y,
             # colour = ifelse(perc_diff > 0,
             #                                            col_green_text,
             #                                            ifelse(perc_diff < 0,
             #                                                   col_orange_text,
             #                                                   "#333333")),
    mapping = 
      aes(
        label = 
          # If country is (the first one), plot numbers with %
          ifelse(minor_category == first_entry,
                 # Value_if_True:
                 paste0(as.character(round(workers_percent*100,0)),"%"),
                 # Value_if_False
                 paste0(as.character(round(workers_percent*100,0)))
                 )
        #,color = ifelse(perc_diff == 0, "zero", gender)
      )
  ) + 
  # Add "Legend" above (the first one)
  geom_text(
    # Font size
    size = font_size_titles, fontface = "bold", nudge_y = nudge_y_titles,
    mapping = 
      aes(label = 
          # If it is first_entry, plot legend
          ifelse(minor_category == first_entry,
                 # Value_if_True:
                 ifelse(gender == "male_workers",
                        # Value_if_True:
                        "Male",
                        # Value_if_False:
                        "Female"
                        ),
                 # Value_if_False
                 ""
          ),
        color = ifelse(perc_diff == 0, "zero", gender)
      )
  ) +
  # Change dot colors
  scale_color_manual(values = c(col_orange, col_green, col_gray)) +
  # Change scale x axis
  scale_x_continuous(
    # Set limits to 0 and 1.2 (we won't set it to 1 because we neeed some space
    # after 1 to place the values of the differences)
    limits = c(0,1.2),
    # Show tick marks at every 25%
    breaks = seq(0,1,.25),
    # Change scale to percent
    labels = scales::percent
  ) +
  # Expand y axis scale so that the legend can fit
  scale_y_discrete(expand = expand_scale(add=c(0.65,1))) +
  # Add white rectangle to set the area where the values of the differences will
  # be
  geom_rect( mapping = aes(xmin = 1.01, xmax = Inf , ymin = -Inf, ymax = Inf),
             fill = "white",
             color = "white") +
  # Add rectangle with correct banground color for the differences
  geom_rect( mapping = aes(xmin = 1.05, xmax = 1.15 , ymin = -Inf, ymax = Inf),
    fill = col_background_diff,
    color = col_background_diff ) +
  # Add Differences values
  geom_text(
    fontface = "bold",size = 4, colour = "black",
    mapping = 
      aes(
        x = 1.1,
        y = minor_category,
        label = 
          # To avoid duplicate values, plot empty text for the first group and
          # plot the difference only for the Ages_18_34 group.
          ifelse(gender == "male_workers",
                 # Value_if_True
                 "",
                 #Value_if_False
                 # If the difference is equal to zero, do not put any signal. 
                 # Otherwise, if Positive, put the + sign on the front.
                 ifelse(perc_diff == 0,
                        # Value_if_True:
                        paste0(as.character(round(perc_diff*100,0))),
                        # Value_if_False
                        ifelse(perc_diff > 0, 
                               # Value_if_True
                               paste0("+",as.character(round(perc_diff*100,0))),
                               # Value_if_False
                               paste0(as.character(round(perc_diff*100,0)))
                        )
                 )
              )
      )
  ) +
  # Insert Title of Differences
  geom_text(
    # Bold face
    fontface = "bold",
    # Font size
    size = font_size_titles,
    # Cor
    colour = col_dark_gray,
    # Set text a little above the dots
    nudge_y = nudge_y_titles,
    # Position
    mapping = 
      aes(
        x = 1.1,
        y = minor_category,
        label = 
          # If its 1st entry, plot values
          ifelse(minor_category == first_entry,
                 # Value_if_True
                 "Diff",
                 #Value_if_False
                 ""
          )
      )
  ) +
  # Plot Title and Axis Labels
  labs(
    title = title_plot,
    subtitle = "USA data on 2016",
    caption = "#tiydtuesday 2019-03-05",
    x = "", y = "" ) +
  # Change background, General font size, and other things
  theme(
    # Change font color and text for all text outside geom_text
    text = element_text(color = "#4e4d47", size = 10),
    
    # Country names in bold face
    # axis.text.y = element_text(face = "bold"),
    
    # Add space between x axis text and plot 
    axis.text.x = element_text(vjust = -0.75),
    
    # Do not show tick marks
    axis.ticks = element_blank(),
    
    # Delete original legend (keep only the one we created)
    legend.position = "none",
    
    # White background
    panel.background = element_blank(),
    
    # Country (y Axis) Lines
    panel.grid.major.y = element_line(colour = col_light_gray, size = 1),
    
    # Change Title Font
    plot.title = element_text(face = "bold", size = 16),
    
    # Change Subtitle Font and add some margin
    plot.subtitle = element_text(face = "italic", size = 12,
                                 margin = margin(b = 0.5, unit = "cm"))
  ) +
    ggsave(file=here::here("jas1_weeks","2019","2019-03-05","after-ggsave.png"),
           type="cairo",device="png")

# here::here("jas1_weeks","2019","2019-03-05","after.png")
# png(here::here("jas1_weeks","2019","2019-03-05","after.png"), height = 600,width = 700) 
# p 
```


## communicate


### general view:


### summary: 

### full story:


### to close ...:


### useful resources:

- https://twitter.com/frizzyroselle/status/1104042846645043200
- https://ogustavo.com/post/dotplot-ggplot/
