
---
title: "Tidy tuesday challenge: Week 2019-06-25 UFO's world"
author: "julio"
date: "2019-06-25"
output: html_document
---

# Tidy tuesday challenge: Week 2019-06-25 UFO's world !

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

### Data:

UFO's world

### objectives:

check the data and do something.

## details:


## import data
```{r echo=FALSE,message=FALSE,warning=FALSE}
library(magrittr) # para el %T>%
library(tidyverse)
# library(sf)
library(dplyr)
library(stringr)#;
library(rebus)#; install.packages('rebus')
library(tidytext)

# install.packages("Rcpp")
# remotes::install_github("tylermorganwall/rayshader")
library(rayshader)


```


```{r echo=FALSE,message=FALSE,warning=FALSE}
ufo_sightings <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-25/ufo_sightings.csv")

ufo_sightings_local <- readr::read_csv(here::here("data","2019","2019-06-25","ufo_sightings.csv")) %>% 
    mutate(date_time_parsed=lubridate::mdy_hm(date_time)) %>%  # remember to see data dictionary if available :p 
    mutate(year=lubridate::year(date_time_parsed)) %>% 
    mutate(decade=floor(year/10)*10) %>% 
    mutate(end_year_decade=lubridate::dmy(paste0("31-12-",decade)))
```

# explore data

```{r echo=FALSE,message=FALSE,warning=FALSE}
glimpse(ufo_sightings_local)

```

```{r echo=FALSE,message=FALSE,warning=FALSE}
ufo_sightings_local %>% skimr::skim()

```

well we got long, lat, can do something mappy . nevertheless got countries  missing, making it difficult make a map with those. 

seeing which city area is related to contury to see if we can see some relation of extraction that can be done.

also there is one with no latitude :/


```{r echo=FALSE,message=FALSE,warning=FALSE}

ufo_sightings_local %>%
    filter(is.na(country) ) %>%
    select(city_area)



```
somehting can be done with some, but its quite difficult to impute all countries. 
so will use non missing values, missing values wont be used.

other stuff that can be done. its see wich kind of ufo sighting have more movment over time. to see if there is some change on that.

## decade with "field country"

```{r echo=FALSE,message=FALSE,warning=FALSE}

year_country <- ufo_sightings_local %>%  
    filter (!is.na(country)) 



# year_country %>% filter(is.na(year))

count_by_year_country <- year_country %>% 
    group_by(end_year_decade,country) %>% 
    summarise(count=n())

ufo_sightings_by_country_data <- count_by_year_country %>%
    
    ggplot(aes(x=end_year_decade,y=count,color=country)) +
    geom_line()+
    scale_x_date(date_breaks = "5 years",date_labels = "%Y")+
    facet_grid(country~.,scales = "free_y")+
    labs(title="UFO Sightings by decade based country on data; 1910 - 2014", 
         x="",y="# UFO Sightings",
         subtitle = "Where country is not NA",
         caption="#tidytuesday")+
    
    theme_light()+
    theme(legend.position = "none",
          axis.text.x = element_text(angle=90))
# 
# ggsave(ufo_sightings_by_country_data,filename = "ufo_sightings_by_country_data.png",width = 10,height = 5)

```

ok , more usefull information will come from geoposition and countrie sor continents


```{r echo=FALSE,message=FALSE,warning=FALSE}
library(rgdal)
library(geojsonsf)#;install.packages("geojsonsf")
library(sf)
library(sp)
library(spatialEco)


# world_ggplot <- ggplot2::map_data("world")
# https://gis.stackexchange.com/questions/230608/creating-an-sf-object-from-the-maps-package
world_sf <- sf::st_as_sf(maps::map("world", plot = FALSE, fill = TRUE))


ufo_complete_points <- ufo_sightings_local %>% # complete lat long
    filter(!is.na(latitude) & !is.na(longitude))
    

ufo_sight_points <- st_as_sf(ufo_complete_points, coords = c("longitude", "latitude"))

st_crs(ufo_sight_points) <- st_crs(world_sf)

#https://stackoverflow.com/questions/3647744/intersecting-points-and-polygons-in-r
# library(spatialEco)#; install.packages("spatialEco")

ufo_by_country  <- point.in.poly(ufo_sight_points, world_sf)


ufo_country_sf <- st_as_sf(ufo_by_country) %>% 
    rename(country=ID)


```


# ggplot map
```{r echo=FALSE,message=FALSE,warning=FALSE}
ufo_summary <- ufo_country_sf %>% 
    group_by(country.1) %>% 
    summarise(total_ufos=n()) %>% 
    st_set_geometry(NULL)


ufo_country_map <- world_sf %>% 
    left_join(ufo_summary, 
               by = c("ID"="country.1"))

```



## decade with "geolocation"
```{r}
year_country_geoloc <- ufo_country_sf %>% 
    st_set_geometry(NULL)

count_by_year_country_geoloc <- year_country_geoloc %>% 
    group_by(end_year_decade,country.1) %>% 
    summarise(count=n())


filter_countries_at_least_5_years <- count_by_year_country_geoloc %>% 
    ungroup() %>% 
    count(country.1) %>% 
    filter(n > 5) %>% pull(country.1)

ufo_sightings_by_country_data_geoloc <- count_by_year_country_geoloc %>%
    filter(country.1%in%filter_countries_at_least_5_years) %>% 
    ggplot(aes(x=end_year_decade,y=count,color=country.1)) +
    geom_line()+
    geom_point()+
    scale_x_date(date_breaks = "5 years",date_labels = "%Y")+
    facet_grid(country.1~.,scales = "free_y")+
    labs(title="UFO Sightings by decade by country ( geolocation ) 1910 - 2014", 
         x="",y="# UFO Sightings",
         subtitle = "Where latitud & longitude are not NA, AND at least 5 reports in diferent years",
         caption="#tidytuesday")+
    theme_light()+
    theme(legend.position = "none",
          axis.text.x = element_text(angle=90))
# 
# ggsave(ufo_sightings_by_country_data_geoloc,filename = "ufo_sightings_by_country_data_imputed_crossing.png",width = 10,height = 20)
```


## map 2d

```{r}


# range(year_country$year)
ufo_country_map_plot <- ufo_country_map %>% 
    # filter(total_meteors<5000) %>% 
    ggplot()+
    geom_sf(aes(fill=log(total_ufos)+1),color=NA)+#+
    # scale_fill_brewer(palette = "YlGnBu",na.value="grey")+
    # scale_fill_brewer(palette = "PuBu",na.value="grey")+
    # scale_fill_viridis_d(na.value="grey")+
    scale_fill_viridis_c(na.value="grey")+
    theme( panel.border = element_blank(),
      panel.background = element_blank())+
    labs(title="Total amount of Ufo's by country",
         subtitle="# UFO's (log of total sights) ; 1910 - 2014",
         fill="# UFO's",
         caption="#tidytuesday")
ufo_country_map_plot
# ggsave(ufo_country_map_plot,filename = "ufo_country_map_plot.png",width = 10,height = 5)


# n_meteors_country
    
# pp = ggplot(data, aes(x=x, y=y)) +
#   geom_hex(bins = 20, size = 0.5, color = "black") +
#   scale_fill_viridis_c(option = "C")
# plot_gg(pp, width = 4, height = 4, scale = 300, multicore = TRUE)


```

# just another rayshader test :D

```{r}

count_by_decade_country_geoloc <- year_country_geoloc %>% 
    group_by(decade,country.1) %>% 
    summarise(count=n())


ufo_sightings_by_country_data_geoloc_rayshader <- count_by_decade_country_geoloc %>%
    filter(country.1%in%filter_countries_at_least_5_years) %>% 
    # mutate(country.1=fct_reorder2(country.1,decade,count)) %>%
    ggplot(aes(x=decade ,y=country.1,fill=log(count))) +
    geom_tile(color="black")+
    scale_fill_viridis_c(option = "C")+
    scale_x_continuous(breaks = seq(from = 1900,to=2015,by = 10),
                       labels = seq(from = 1900,to=2015,by = 10))+
    labs(title="UFO Sightings by decade by country\n( geolocation ) 1910 - 2014", 
         x="Decade",y="",
         fill="Log of\n# UFO Sightings",
         subtitle = "Where latitud & longitude are not NA,\nAND at least 5 reports in diferent years\nNA country are related to No Matching area points on Map",
         caption="#tidytuesday")+
    theme_light()+
    theme(axis.text.x = element_text(angle=90))

ggsave(ufo_sightings_by_country_data_geoloc_rayshader,filename = "ufo_sightings_by_country_data_imputed_crossing_for_ray_shader.png",width = 5,height = 5)
    


# plot_gg(ufo_sightings_by_country_data_geoloc_rayshader, 
#         width = 5, height = 5, scale = 300, multicore = TRUE)
# render_camera(fov = 70, zoom = 0.5, theta = 130, phi = 35)
# render_camera(fov = 70, zoom = 0.5, theta = 130, phi = 35)
# render_snapshot(clear = TRUE)

# http://blog.schochastics.net/post/mapping-nba-shot-locations/
plot_gg(ufo_sightings_by_country_data_geoloc_rayshader, width = 5, height = 5, multicore = TRUE, scale = 250, 
        zoom = 0.7, theta = 10, phi = 30, windowsize = c(800, 800))

phivec <- 20 + 70 * 1/(1 + exp(seq(-5, 10, length.out = 180)))
phivecfull <- c(phivec, rev(phivec))
thetavec <- 90 * sin(seq(0,359,length.out = 360) * pi/180) #+270
zoomvec <- 0.5 + 0.5 * 1/(1 + exp(seq(-5, 10, length.out = 180)))
zoomvecfull <- c(zoomvec, rev(zoomvec))

for(i in 116:360) {
  render_camera(theta = thetavec[i],phi = phivecfull[i],zoom = zoomvecfull[i])
  # render_snapshot(paste0("anim/frame", i, ".png"))
    render_snapshot(here::here("jas1_weeks","2019","2019-06-25",
                               "anim",paste0("frame", i, ".png")))
}

#Run this command in the command line using ffmpeg to stitch together a video:
#ffmpeg -framerate 60 -i frame%d.png -vcodec libx264 raymovie.mp4

#And run this command to convert the video to post to the web:
#ffmpeg -i raymovie.mp4 -pix_fmt yuv420p -profile:v baseline -level 3 -vf scale=-2:-2 rayweb.mp4


```

    



# idea , do prophet stuff to see how it evolves over time.

well not enough time today to do this.
will be next time :p 

```{r echo=FALSE,message=FALSE,warning=FALSE}
# library(prophet)
# 


```



#tweet: 

2019-06-25 #TidyTuesday #rstats UFO Sightings! 
Time lines and map! geolocation is amazing for getting missing countries :D
and yet another #rayshader , thanks @tylermorganwall for outstanding job !
also thanks: @schochastics for cool work & code :D

https://twitter.com/jspairani/status/1145536192605315072

https://twitter.com/jspairani/status/1145536194576617472

# communicate

well got stubborn with getting a rayshader done. 
also with getting the countries in a better way. 
thats all for today. 

would have been awsome to do something like : 
https://twitter.com/jakekaupp/status/1144063378543038465

as mentioned weeks like crossing population with events spoted ( meteors )

for the bivariate choroplet. 
also this for may be more relevant data: 

https://twitter.com/whipson3/status/1144063351313588226

( airports near ? may be planes ?)


may be next time :D 
also got to do the #datosdemiercoles :U!



