# 2018-09-11: first tidy tuesday: cat & dogs

## keep it simple:

### Objectives: 

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate

https://rviews.rstudio.com/2017/06/08/what-is-the-tidyverse/ 

( migrating from ETL + analysis)


### workflow:

1st iteration:

* get the git project: https://github.com/rfordatascience/tidytuesday
* add a new folder where your works will be placed, ex: 'jas1_weeks'
* inside it, add week folders inside with naming == data folder ; ex: '2018-09-11'
* add your rmd file for documenting your workflow, notes & stuff ; ex: 'tidytuesday/jas1_weeks/2018-09-11/readme.rmd'
* get to the specific challenge tasks

#### Understand the business

* to understand what to do , you need to check the challenge objectives;
* read the article wich is being analized; more information on the data folder of the week: tidytuesday/data/2018-09-11

https://www.washingtonpost.com/news/wonk/wp/2014/07/28/where-cats-are-more-popular-than-dogs-in-the-u-s-and-all-over-the-world/

> some css tricks for reading: 
> * F12 
> * display: none; 
> * overflow: auto;

* make notes / scripts , document what you do.

#### Define your goal to do:

as 1st timer on tidy tuesday

* show the ordered tables
* make the map, hex-tiled
* add the cat/dog histogram from: https://data.world/datanerd/cat-vs-dog-popularity-in-u-s
* update the objective as i acommplish task; ( new ideas, or at least document " future work" )


#### Import: git clone + git merge from the original

* import the data from the current dataset (ctrl + alt + i: shortcut for insert R chunks in rmd in rstudio)
* peek raw data to have an intuition,  ( always remember to check file size, because of larger datasets )
* refactor to make the proper read. (strings as factors= false)

```{r}
library(dplyr)

file_location <- '../../data/2018-09-11/cats_vs_dogs.csv'
raw_data <- read.csv(file_location,stringsAsFactors = FALSE)
glimpse( raw_data) 

```

#### Tidy: 


We can see, we got 49 observatitions, and 13 variables; and how are they structured.

```{r}
data <- raw_data %>% rename(id=X)
glimpse(data)
```

as 1st stransformation, X , its given the "id" name; as i think its more descriptive

we can see 2 variables do not have any sense to calculate stuff as they are categorical variables: state & id.

id for identification of observation
and state, as a "category" / "grouping"
( http://www.abs.gov.au/websitedbs/a3121120.nsf/home/statistical+language+-+what+are+variables )


#### Understand: 

##### Transform

lets take a look to some detail to all the variables; as im not quite happy with default "summary" just google tidiyverse summary; after digging little,  and found skimr from ropensci: https://ropensci.org/blog/2017/07/11/skimr/


```{r}
library(skimr)#install.packages('skimr')
skim(data)
```

thats an oustanding summary; it tells you a lot , first of all , no missing values on any variable.
and a mini histogram from em.

to take a look at distributions, always found helpfull the box plot: 
for that we need to transform the dataset from wide to long.

```{r}

library(tidyr)

long_data <- gather(data,key=variable,value=value,-state,-id)
glimpse(long_data)
```


##### Visuailze
```{r}


library(ggplot2)
ggplot(long_data,aes(y=value,x=variable,fill=variable)) + 
    geom_boxplot() + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1))

```


As seen some variables are not on same range of values: the avg & the percent.

##### Transform
```{r}
library(stringi)
library(stringr)
percent_data <- long_data%>% filter(str_detect(variable,"percent_") )
avg_data <- long_data%>% filter(str_detect(variable,"avg_") )
other_data <- long_data %>% filter( !str_detect(variable,"percent_")) %>% filter( !str_detect(variable,"avg_") )

# just to check the filter its ok
percent_data %>% group_by(variable) %>% tally() %>% select(variable)
avg_data %>% group_by(variable) %>% tally()  %>% select(variable)
other_data %>% group_by(variable) %>% tally()  %>% select(variable)

```

##### Visuailze

```{r}


ggplot(percent_data,aes(y=value,x=variable,fill=variable)) + 
    geom_boxplot() + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplot(avg_data,aes(y=value,x=variable,fill=variable)) + 
    geom_boxplot() + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplot(other_data,aes(y=value,x=variable,fill=variable)) + 
    geom_boxplot() + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1))

```


watching the variables can be said a lot of stuff; and can build some rations on them. 

as time assigned here, i'll skip this analysis for later / future.




# Outputs: 

For closing this first stage of tidytuesday , ill focus on the derivables

## show the ordered tables


before checking order, we need to see if data is consistent; not only numbers, but for the domain being analized.

in this case quantity of states, and no repeating; so we count the ocurrences, 
filter with N will be greather than 1

```{r}
data %>% count(state) %>% filter(n > 1)
```

luckily no repetitions.

Columns needed: 

Rank 	State 	Cats 	Dogs 	Ratio, dogs to cats Ratio, cats to dogs



```{r}
process_data <- data %>% 
    select(state,cat_population,dog_population) %>% 
    mutate(ratio_dog_cat = dog_population/cat_population ) %>%
    mutate(ratio_cat_dog = cat_population/dog_population )
```

### Top 10 dog-loving states

```{r}
process_data %>% top_n(10,wt = ratio_dog_cat) %>% arrange(desc(ratio_dog_cat))

```

### Top 10 cat-loving states

```{r}
process_data %>% top_n(10,wt = ratio_cat_dog) %>% arrange(desc(ratio_cat_dog))

```


## Cat/dog histogram from: 

https://data.world/datanerd/cat-vs-dog-popularity-in-u-s

needed the percent difference: dogs - cats

usefull links:

* http://www.sthda.com/english/wiki/ggplot2-title-main-axis-and-legend-titles
* http://www.sthda.com/english/wiki/ggplot2-essentials
* http://www.sthda.com/english/wiki/ggplot2-axis-ticks-a-guide-to-customize-tick-marks-and-labels
* http://www.sthda.com/english/wiki/ggplot2-legend-easy-steps-to-change-the-position-and-the-appearance-of-a-graph-legend-in-r-software


```{r}



process_data_hist_all <- data %>%
    mutate(percent_diff = percent_dog_owners-percent_cat_owners) %>%
    mutate(animal=if_else(percent_diff >0,"Dog","Cat"))

process_data_hist <- process_data_hist_all  %>% 
    select(state,percent_diff,animal) %>% 
    arrange(desc(percent_diff))

# glimpse(process_data_hist)

# process_data_hist %>% arrange(desc(percent_diff)) 

# ggplot(process_data_hist,aes(x=state))+geom_bar(stat = id)

ggplot(data = process_data_hist, 
       mapping = aes(x = reorder(state, percent_diff), fill = animal, 
                     y = ifelse(animal == "Cat", percent_diff, percent_diff))) +
  geom_bar(stat = "identity", # for the value
           width=0.4 ) +#, # for thiner bars
          # position = position_dodge(width=0.5)) +
  scale_y_continuous(labels = abs, limits = max(process_data_hist$percent_diff) * c(-1,1)) +
  labs(y = "Percent Difference") +
  coord_flip() +
    ggtitle("Dog Vs Cat orientation by State") +
    ylab("Percent Difference") + 
    xlab("State") +
    theme(axis.text.y = element_text( color="#993333", size=9,margin = margin(2, 2, 2, 2)))
    # couldnt make the label axis space ok.

```



## Map, hex-tiled

utils:

https://www.r-graph-gallery.com/328-hexbin-map-of-the-usa/

https://stackoverflow.com/questions/37912418/how-does-geom-map-map-id-function-work

http://ryanhafen.com/blog/geofacet

```{r}


library(geojsonio) #install.packages("geojsonio")
spdf <- geojson_read("us_states_hexgrid.geojson",  what = "sp")
 
# I need to 'fortify' the data to be able to show it with ggplot2 (we need a data frame format)
library(broom)
spdf@data = spdf@data %>% mutate(google_name = gsub(" \\(United States\\)", "", google_name))
spdf_fortified <- tidy(spdf, region = "google_name")


library(rgeos) #install.packages("rgeos")
centers <- cbind.data.frame(data.frame(gCentroid(spdf, byid=TRUE), id=spdf@data$iso3166_2))

library(mapproj) #install.packages("mapproj")


# Now I can plot this shape easily as described before:
# ggplot() +
#   geom_polygon(data = spdf_fortified, aes( x = long, y = lat, group = group), fill="skyblue", color="white") +
#   geom_text(data=centers, aes(x=x, y=y, label=id)) +
#   theme_void() +
#   coord_map()


# percent_data_long <- process_data_hist %>% spread(key = variable,value=value)
# glimpse(percent_data_long)
 
# Merge geospatial and numerical information
spdf_fortified_cat_dogs  <-  spdf_fortified %>% 
    left_join(. , process_data_hist, by=c("id"="state"))

# glimpse(spdf_fortified_cat_dogs)

spdf_fortified_cat_dogs_percent <- spdf_fortified_cat_dogs 
glimpse(spdf_fortified_cat_dogs)


# Make a first chloropleth map
ggplot() +
  geom_polygon(data = spdf_fortified_cat_dogs_percent,
               aes(fill =  percent_diff, x = long, y = lat, group = group)) +
  geom_text(data=centers,
            aes(x=x, y=y, label=id), color="white", size=3, alpha=0.6) +
  theme_void() +
  coord_map() + 
  ggtitle( "Dogs Vs Cats, orientation By state" ) +
theme(
    legend.position = "top",
    text = element_text(color = "#ffffff"))

# # Prepare binning
# spdf_fortified$bin = cut( spdf_fortified$percent_diff , 
#                           breaks=c(seq(5,10), Inf), labels=c("5-6", "6-7", "7-8", "8-9", "9-10", "10+" ), include.lowest = TRUE )
#  
# # Prepare a color scale coming from the viridis color palette
# library(viridis)
# my_palette=rev(magma(8))[c(-1,-8)]
#  
# # plot
# ggplot() +
#   geom_polygon(data = spdf_fortified, 
#                aes(fill = bin, x = long, y = lat, group = group) , size=0, alpha=0.9) +
#   geom_text(data=centers, 
#             aes(x=x, y=y, label=id), color="white", size=3, alpha=0.6) +
#   theme_void() +
#   scale_fill_manual( 
#     values=my_palette, 
#     name="Wedding per 1000 people in 2015", 
#     guide = guide_legend( keyheight = unit(3, units = "mm"), keywidth=unit(12, units = "mm"), label.position = "bottom", title.position = 'top', nrow=1) 
#   ) +
#   ggtitle( "A map of marriage rates, state by state" ) +
#   theme(
#     legend.position = c(0.5, 0.9),
#     text = element_text(color = "#22211d"),
#     plot.background = element_rect(fill = "#f5f5f2", color = NA), 
#     panel.background = element_rect(fill = "#f5f5f2", color = NA), 
#     legend.background = element_rect(fill = "#f5f5f2", color = NA),
#     plot.title = element_text(size= 22, hjust=0.5, color = "#4e4d47", margin = margin(b = -0.1, t = 0.4, l = 2, unit = "cm")),
#   )
#  

```


Enough for today ! 

# future stuff: 

- apply to other maps
- playe more with ggplot 
- analyze deeper
- make a better report, showing / not showing some stuff & comments.