---
title: "Tidy tuesday challenge: Week 2019-09-03 moore's law
author: "julio"
date: "2019-09-08"
output: html_document
---

# Tidy tuesday challenge: Week 2019-09-03 moore's law

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

just explore and plot something =) 

### Data:
 
 Week Week 2019-09-03 moore's law
 
### objectives:

explore & do something

## details:


## import data
```{r echo=FALSE,message=FALSE,warning=FALSE}
library(magrittr) # para el %T>%
library(tidyverse)
# library(sf)
library(dplyr)
library(stringr)#;
# library(rebus)#; install.packages('rebus')
# library(tidytext)
# library(prophet)


# install.packages("Rcpp")
# remotes::install_github("tylermorganwall/rayshader")
# library(rayshader)
library(lubridate)
library(ggforce)
library(ggrepel)

library(tidyquant)
# library(arules)#install.packages('arules')
# library(arulesViz)#install.packages('arulesViz')
# library(igraph)

```


```{r echo=FALSE,message=FALSE,warning=FALSE}
cpu <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-09-03/cpu.csv")

gpu <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-09-03/gpu.csv")

ram <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-09-03/ram.csv")

```

# explore data

```{r echo=FALSE,message=FALSE,warning=FALSE}
cpu %>% head()
gpu %>% head()
ram %>% head()


```

```{r echo=FALSE,message=FALSE,warning=FALSE}

cpu %>% glimpse()
gpu %>% glimpse()
ram %>% glimpse()


```

```{r echo=FALSE,message=FALSE,warning=FALSE}

cpu %>% skimr::skim()
gpu %>% skimr::skim()
ram %>% skimr::skim()

```


ummm as 3 datasets its quite ... not happy ... ill join all in one :p 
procesor, transistor_count, date_of_introduction


```{r echo=FALSE,message=FALSE,warning=FALSE}
all_one <-  cpu %>% 
    select(processor,transistor_count,date_of_introduction) %>% 
    mutate(type="cpu") %>% 
    union_all( (gpu %>% 
                    select(processor,transistor_count,date_of_introduction)%>% 
                    mutate(type="gpu"))) %>% 
    union_all((ram %>% 
                   rename(processor=chip_name) %>% 
                   select(processor,transistor_count,date_of_introduction)%>% 
                   mutate(type="ram"))) %>% 
    rename(name=processor)


all_one_complete <- all_one %>% filter(complete.cases(all_one))

skimr::skim(all_one_complete)
```


simple plot: 
transistors by type over time: 

```{r echo=FALSE,message=FALSE,warning=FALSE}
data_plot <- all_one_complete %>%  
    group_by(type,date_of_introduction) %>% 
    summarise(total_transistors=sum(transistor_count),
              max_transistors=max(transistor_count),
              min_transistors=min(transistor_count))
    
plot_out <- data_plot %>% 
    ggplot(aes(x=date_of_introduction,
               y=log(max_transistors),

               # fill=type
               )) +
    geom_line(aes(colour=type))+
    geom_candlestick(aes(open = log(min_transistors),
                         high = log(max_transistors),
                         low = log(min_transistors),
                         close = log(max_transistors)),alpha=0.5)+
    # geom_boxplot()+
    facet_grid(type~.,scales = "free_y")+
    scale_x_continuous(breaks = seq(from=1960,to=2020, by=5) )+
    theme_light()+
    theme(legend.position = "none",
          strip.text.y = element_text(angle = 360),
          axis.text.x = element_text(angle = 90))+
    labs(title="Transistors over time ( log transistors )",
         subtitle="line: max transistors in year\nbars: candlesticks over transistors",
         x="",y="",
         caption="#TidyTuesday")+
    geom_vline(xintercept = seq(1960,2020,10),linetype=2)
ggsave(plot = plot_out,filename = "transistors_over_time_log.png",width = 7,height = 5)

```


## items as absolute values



#tweet: 

2019-09-03 #TidyTuesday #rstats Moore's Law ! transistors over time :D
experimenting with candlesticks in other context :p 


# communicate

just a simple one

# resources: 
- candlesticks: https://cran.r-project.org/web/packages/tidyquant/vignettes/TQ04-charting-with-tidyquant.html#candlestick-chart