
---
title: "Tidy tuesday challenge: Week 2019-07-22 Womens World Cup"
author: "julio"
date: "2019-06-25"
output: html_document
---

# Tidy tuesday challenge: Week 2019-07-16 r4ds_members

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

### Data:
 
r4ds_members
 
### objectives:

check the data and do something.

## details:


## import data
```{r echo=FALSE,message=FALSE,warning=FALSE}
library(magrittr) # para el %T>%
library(tidyverse)
# library(sf)
library(dplyr)
library(stringr)#;
# library(rebus)#; install.packages('rebus')
# library(tidytext)
library(prophet)


# install.packages("Rcpp")
# remotes::install_github("tylermorganwall/rayshader")
# library(rayshader)
library(lubridate)
library(ggforce)
library(ggrepel)

```


```{r echo=FALSE,message=FALSE,warning=FALSE}
r4ds_members <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-16/r4ds_members.csv")
```

# explore data


```{r echo=FALSE,message=FALSE,warning=FALSE}
r4ds_members %>% head()

```

```{r echo=FALSE,message=FALSE,warning=FALSE}
glimpse(r4ds_members)

```

```{r echo=FALSE,message=FALSE,warning=FALSE}
r4ds_members %>% skimr::skim()


```


# time line of messages_posted
```{r echo=FALSE,message=FALSE,warning=FALSE}
# range(train_data$ds)
train_data <- r4ds_members %>% 
    select(date,messages_posted) %>% 
    rename(ds=date,y=messages_posted)

prophet_model <- prophet(yearly.seasonality = TRUE,
                   weekly.seasonality = TRUE,
                   daily.seasonality = TRUE,fit =FALSE )

prophet_model <- add_country_holidays(prophet_model, country_name = 'US')

prophet_model <- fit.prophet(prophet_model, train_data)

# future:
from_future <- lubridate::as_date("2019-07-01")
to_future <- lubridate::as_date("2019-09-01")
future_sequence <- seq(from=from_future,
                   to = to_future,
                   by=1)# by 1 day


past_and_future_dates <- c(train_data %>% pull(ds),future_sequence)

df_future <- data.frame(ds= past_and_future_dates)

to_september_prediction <- predict(prophet_model,df_future)


prophet_plot_components(prophet_model, to_september_prediction)

plot(prophet_model, to_september_prediction)


train_data_2 <- train_data %>% 
    mutate(ts="original") %>% 
    rename(value=y) %>% 
    mutate(ds=as.Date(ds))


prediction_plot_data <- to_september_prediction %>% 
    select(ds,trend,yhat) %>%
    rename(predicted=yhat) %>% 
    gather(ts,value,-ds) %>%
    mutate(ds=as.Date(ds)) %>% 
    union_all(train_data_2) 



prediction_plot <- prediction_plot_data %>% 
    ggplot(aes(x=as.Date(ds), y=value,color=ts))+
    geom_line(alpha=0.8)+
    geom_vline(xintercept = max(train_data_2$ds),linetype=2,color="red")+
    ggrepel::geom_label_repel(
    data = prediction_plot_data %>% 
        filter(ds==max(train_data_2$ds)) %>% 
        filter(ts=="original"),
    aes(label=paste0("date:",ds,"\nlast value:",ceiling(value))),
    size=3,show.legend = FALSE,
    nudge_y = -30000,
    nudge_x = -100)+
    geom_vline(xintercept = lubridate::as_date(c("2019-07-15",
                                              "2019-07-30",
                                              "2019-08-15",
                                              "2019-08-30")
                                              ),linetype=2)+
    ggrepel::geom_label_repel(
        data = prediction_plot_data %>% 
            filter(ds%in%lubridate::as_date(c("2019-07-15",
                                              "2019-07-30",
                                              "2019-08-15",
                                              "2019-08-30"))) %>% 
            filter(ts=="predicted"),
        aes(label=paste0("date:",ds,"\npredicted:",ceiling(value))),
        size=3,show.legend = FALSE,nudge_y = -40000)+
    scale_x_date(date_breaks = "15 days")+
    theme_light()+
    theme(axis.text.x=element_text(angle=90))+
    facet_zoom(  xlim = c(from_future-weeks(2), to_future))+
    labs(title="R4DS - Slack Community, Messages posted Predicted",
             subtitle=paste0("from ",from_future," till ",to_future),
         caption="#TidyTuesday",
         x="",y="# messages",colour="Time Series")
   prediction_plot
   
  ggsave(plot = prediction_plot,filename = "prediction.png",width = 10,height = 8)
   
```

#tweet: 

2019-07-16 #TidyTuesday #rstats R4ds Slack ! 
Prediction of messages for the following months, 
highlighted each 15 days from 15/7 with line & label
and facet_zoom in the zone of interest.
Enjoy :D


# communicate

not the pretiest plot but at least done some modeling & stuff :D







