---
title: "week 2019-01-01 rtweet #rstats #tidytuesday"
author: "_jas"
date: 20190101
output: html_document
---

# Tidy tuesday challenge: Week 2019-01-01

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

### Data:

this week data is twitter data on #rstats and #tidytuesday  hashtags


### objectives:

- get amount per week
- check for rstats & tidytuesday at same time
- may be: check for "lagged submits" ( as submissions not made on tuesday of that week)
- may be: newcomers / constancy / one submit.
- may be: "world propagation"?


## import data

```{r echo=FALSE,message=FALSE,warning=FALSE}
library(here)
library(readr)
library(dplyr)
library(tidyr)
library(scales)
library(ggplot2)
library(forcats)#install.packages('forcats')
library(lubridate)
library(visdat)#install.packages('visdat')
library(stringr)
```



```{r echo=FALSE,message=FALSE,warning=FALSE}
tidytuesday_tweets_file <- here::here("data","2019","2019-01-01","tidytuesday_tweets.rds")
tidytuesday_tweets_raw <- readRDS(tidytuesday_tweets_file)

# set lubridate start on tuesday
options(lubridate.week.start = 1)

dplyr::glimpse(tidytuesday_tweets_raw)

```

# tweets by week, week days / retweets / favorites per week / words per week / users per week

```{r echo=FALSE,message=FALSE,warning=FALSE}
group_week <- tidytuesday_tweets_raw %>% 
    mutate(week_n = lubridate::week(created_at)) %>% 
    filter(week_n!=51) %>% 
    group_by(week_n) %>% 
    summarise(week_n_total = n(),
              week_min=min(created_at),
              week_max=max(created_at),
              sum_retweets_week=sum(retweet_count),
              sum_favourites_week=sum(favorite_count),
              acum_words = paste0(text,collapse = "|"),
              acum_users = paste0(unique(screen_name),collapse=";"),
              unique_users_count=length(unlist(str_split(acum_users,";"))))


gathered_gw <- group_week %>% 
    gather(key = "key",value = "value", -week_n) %>% 
    mutate(value=as.numeric(value))

glimpse(gathered_gw)

gathered_gw %>% 
    filter(key %in% 
               c("week_n_total","sum_retweets_week","sum_favourites_week","unique_users_count")) %>%
    # filter(key %in% c("week_n_total","sum_retweets_week","acum_users_count")) %>% 
    ggplot(aes(week_n,value,colour =key)) +
    geom_line() +
    geom_point()+
    theme_light()+
    xlim(1,52 ) +
    labs(title="Tidytuesday hastag metrics by week",
         x="Week",
         y="Amount", 
         subtitle = "week 51 discarded",
         colour ="metric")
```

# user importance by weeks participated

```{r echo=FALSE,message=FALSE,warning=FALSE}
group_users <- tidytuesday_tweets_raw %>% 
    mutate(week_n = lubridate::week(created_at)) %>% 
    mutate(week_day = lubridate::wday(created_at)) %>% 
    # select(created_at,week_day,week_n)
    mutate(is_tuesday = week_day==2) %>% # options(lubridate.week.start = 1)
    group_by(screen_name,user_id) %>% 
    summarise(count_tweets = n() ,
              tuesdays=sum(is_tuesday),
              not_tuesday=sum(!is_tuesday),
              ratio_tuesdays= tuesdays / count_tweets,
              weeks_in=paste0(unique(week_n),collapse=";"),
              weeks_in_splitted=str_split(weeks_in,";"),
              weeks_count=length(unlist(weeks_in_splitted)),
              week_min=min(week_n),
              week_max=max(week_n),
              week_gap=week_max-week_min, # difference between 1st week and last week
              tweets_week_gap_ratio=if_else(week_gap
                                            !=0,
                                            count_tweets/week_gap,0), # ratio tweests per wek gap
              tweets_week_participate = count_tweets / weeks_count,# ratio weighted by week count
              date_min=min(created_at),
              date_max=max(created_at)) %>% 
    arrange(desc(weeks_count))
glimpse(group_users)
group_users %>% filter(screen_name=='jspairani')

group_users%>% 
    arrange(desc(weeks_count)) %>% 
    head(20) %>% arrange(desc(ratio_tuesdays))


group_users%>% 
    arrange(desc(weeks_count)) %>% 
    head(20) %>% arrange(desc(ratio_tuesdays)) 


group_users_gather <- group_users %>% 
        head(20) %>% 
        ungroup() %>% 
        mutate(screen_name=fct_reorder(screen_name,count_tweets)) %>% 
        select(-user_id,-weeks_in_splitted,-weeks_in,-date_min,-date_max) %>% 
        gather(key="key",
               value="value",
               -screen_name)

group_users_gather %>% 
    filter(key%in%c("tuesdays","not_tuesday")) %>% 
    ggplot(aes(screen_name,value,fill=key))+
    geom_col()+
    coord_flip()+
    labs(title="Top 20 by amount Tweets in tidy tuesday",
         subtitle = "week 51 discarded",
         x="",
         y="#amount tweets",
         fill="posted on")


```


# user participation by week

```{r echo=FALSE,message=FALSE,warning=FALSE}
group_users_gather2 <- group_users %>% 
        ungroup() %>% 
        mutate(screen_name=fct_reorder(screen_name,count_tweets)) %>% 
        select(-user_id,-weeks_in_splitted,-weeks_in,-date_min,-date_max) %>% 
        gather(key="key",
               value="value",
               -screen_name)

spread2 <- group_users_gather2 %>% 
    filter(key%in%c("weeks_count","count_tweets","tweets_week_participate")) %>% 
    spread(key,value) 

threshold <- 3


# it really doesnt matter to this business. 
# as everyone will post 1 with the tags, except the administrators :P
spread2 %>% 
    ggplot(aes(x=weeks_count,
               y=count_tweets,
               size=tweets_week_participate,
               label=paste0("@",screen_name)))+
    geom_point(alpha=0.5)+
    ggrepel::geom_text_repel(data=subset(spread2,tweets_week_participate > threshold) ,

                    hjust=5,
                    x_nudge=50,
                    # direction = "x",
                    size=3,
                    segment.size = 0.4,
                    segment.color="blue",
                    color="blue"
                    )+ 
    labs(title="Participated weeks vs tweet amounts in tidy tuesday",
         subtitle = "week 51 discarded",
         x="#participated in N weeks",
         y="#amount tweets",
         size="participation ratio ( tweet/week )") +
    theme_light()
    # theme(label.)


# still want to show something with weeks of participation 

spread2 %>% 
    arrange(desc(weeks_count)) %>% 
    head(20) %>% 
    mutate(screen_name=fct_reorder(screen_name,weeks_count)) %>% 
    ggplot(aes(screen_name,weeks_count,fill=screen_name))+
    geom_col()+
    coord_flip()+
    labs(title="Top 20 by week participation in tidy tuesday",
         x="",
         y="#amount of weeks participated",
         fill="posted on")+
    theme_light()+
    theme(legend.position = "none")

```

# slow users: 

- idea was not on tuesday , and hoyw mucha mount of delay.

```{r echo=FALSE,message=FALSE,warning=FALSE}
# group_weeks_user <- tidytuesday_tweets_raw %>% 
#     mutate(week_n = lubridate::week(created_at)) %>% 
#     
#     group_by(week_n) %>% 
#     summarise(week_n_total = n(),
#               week_min=min(created_at)+ days(1),# starts pm tuesday
#               week_max=max(created_at)+ days(1),
#               sum_retweets_week=sum(retweet_count),
#               sum_favourites_week=sum(favorite_count),
#               acum_words = paste0(text,collapse = "|"),
#               acum_users = paste0(unique(screen_name),collapse=";"))
```