---
title: "readme.rmd"
author: "_jas"
date: 2018-11-06
output: 
    html_document
---


# Tidy tuesday challenge 2018-11-06: 


### Objectives: 

* keep it simple,
* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate

### news

this week a particular usefull cheatsheet came out:

https://www.business-science.io/r-cheatsheet.html

https://www.business-science.io/assets/2018-11-04-data-science-with-R-cheatsheet/Data_Science_With_R_Workflow.png

### add the tmp folder

it is not on github but you should download the temp data folder
in this time the shape file from: https://eerscmap.usgs.gov/uswtdb/data/




### this week objective

- clean data
- ask some question
- merge with spatial map


# proposed questions:


* How do newer installations compare to older turbines?

* Which states/regions have the most wind farms?

* Spread of wind-turbines over time?

* Where is the most missing data?


```{r echo=FALSE,message=FALSE,warning=FALSE}

library(here)
library(readr)
library(tibble)
library(dplyr)
library(tidyr)
library(stringr)
library(stringi)#; install.packages("stringi")
library(sf)
library(ggplot2)
library(visdat)
library(naniar)
library(rpart)
library(rpart.plot)
library(plotly)
library(ggridges)
library(forcats)
library(DataExplorer)#; install.packages("DataExplorer")

```

## Import

```{r echo=FALSE,message=FALSE,warning=FALSE}
wind_raw <- readr::read_csv(here("data","2018-11-06","us_wind.csv"),
                            na = c("missing",-9999,"n/a"))
```

## Importing shape

got issues importing the shape file :

Error in CPL_read_ogr(dsn, layer, as.character(options), quiet, type,  : 
  Open failed.

```{r echo=FALSE,message=FALSE,warning=FALSE}
# wind_shape_raw <- sf::st_read(here("jas1_weeks","2018-11-06","tmp","uswtdbSHP","uswtdbSHP.shp"))
```

As time goes, ill go for the think i know how to do:

## Where is the most missing data?

```{r echo=FALSE,message=FALSE,warning=FALSE}
# wind_shape_raw <- sf::st_read(here("jas1_weeks","2018-11-06","tmp","uswtdbSHP","uswtdbSHP.shp"))

# visdat::vis_dat(wind_raw)
# visdat::vis_miss(wind_raw)

# glimpse(wind_raw)

wind_miss_prop <- wind_raw %>% naniar::add_prop_miss() %>% mutate(have_miss= (prop_miss_all== 0) )

cant <- wind_miss_prop %>% count() %>% pull(n)

complete <- wind_miss_prop %>% arrange(desc( prop_miss_all)) %>% 
    filter( prop_miss_all== 0) %>% 
    count() %>% pull(n)

miss <- wind_miss_prop %>% arrange(desc( prop_miss_all)) %>% 
    filter( prop_miss_all > 0) %>% 
    count() %>% pull(n)


wind_miss_by_state <- wind_miss_prop %>% 
    group_by(t_state,have_miss) %>% 
    count() 


wind_miss_prop_by_state <- wind_miss_prop %>% 
    group_by(t_state,have_miss) %>% 
    count() %>% 
    spread(have_miss,n,fill = 0) %>% 
    rename(miss='TRUE',complete='FALSE') %>% 
    mutate(miss_percent=round(miss/complete*100,2))

skimr::skim(wind_raw)
plot_miss <- naniar::gg_miss_var(wind_raw) + 
    labs(title="Wind turbines dataset amount of missing values",
         subtitle="complete: 20.26%, missing 79.73%",y = "# Missings")

plotly::ggplotly(plot_miss)


    


wind_miss_by_state %>% 
    arrange(desc(n)) %>% 
    ungroup() %>% 
    mutate(t_state=fct_reorder(t_state,n)) %>% 
    ggplot(aes(x=t_state,y=n,fill=have_miss)) +
    geom_col() +
    labs(x="",y="# cases",title="Missing Values by State")+
    coord_flip()



# 
# 100 * miss / cant
# 100 * complete / cant
```
## lets give other package a chance: Data explorer

```{r echo=FALSE,message=FALSE,warning=FALSE}
DataExplorer::create_report(wind_raw)
```

```{r echo=FALSE,message=FALSE,warning=FALSE}
# Geospatial data available in the geojson format!
library(geojsonio)
spdf <- geojson_read("https://raw.githubusercontent.com/gregoiredavid/france-geojson/master/communes.geojson",  what = "sp")
 
# Since it is a bit to much data, I select only a subset of it:
spdf = spdf[ substr(spdf@data$code,1,2)  %in% c("06", "83", "13", "30", "34", "11", "66") , ]
 
# I need to fortify the data AND keep trace of the commune code! (Takes 2 minutes)
library(broom)
spdf_fortified <- tidy(spdf, region = "code")
 
# Now I can plot this shape easily as described before:
ggplot() +
  geom_polygon(data = spdf_fortified, aes( x = long, y = lat, group = group)) +
  theme_void() +
  coord_map()

```



```{r echo=FALSE,message=FALSE,warning=FALSE}
wind_raw %>% 
    group_by(p_year,t_state,t_county) %>% 
    summarise(n_turbines=n()) %>% 
    arrange(desc(n_turbines)) %>% 
    filter(is.na(p_year))

wind_raw %>% 
    group_by(p_year,t_state,t_county) %>% 
    summarise(n_turbines=n()) %>% 
    arrange(desc(n_turbines)) %>% 
    filter(is.na(p_year))

```

Looks like there are some missing years. ( 62 )
```{r echo=FALSE,message=FALSE,warning=FALSE}
wind_raw %>%  filter(complete.cases(wind_raw)) %>%  count()
wind_raw %>%  filter(!complete.cases(wind_raw)) %>%  count()
```

```{r echo=FALSE,message=FALSE,warning=FALSE}


wind_raw %>% 
    group_by(p_year,t_state,t_county) %>% 
    summarise(n_turbines=n()) %>% 
    arrange(desc(n_turbines)) %>% 
    filter(p_year>2000) %>% 
    ggplot(aes(x=n_turbines, y=as.character(p_year)))+
        geom_density_ridges(alpha = 0.2)+
    theme_light() +
    labs(x="# turbines",
         y="", 
         title="Amount of turbines per year")

wind_raw %>% mutate(p_name=fct_lump(p_name,7))%>% count(p_name) 
wind_raw %>% count(p_name) %>% arrange(desc(n))
wind_raw %>% count(t_state) %>% arrange(desc(n))

wind_raw %>% 
    group_by(p_year,t_state,t_county) %>% 
    summarise(n_turbines=n()) %>% 
    arrange(desc(n_turbines)) %>% 
    filter(p_year<2000) %>% 
    ggplot(aes(x=n_turbines, y=as.character(p_year)))+
        geom_density_ridges(alpha = 0.2)+
    theme_light() +
    labs(x="# turbines",
         y="", 
         title="Amount of turbines per year")


```


# communicate

## half done

map things ... always got some issues with them :/ ; generally run out of time or got some issues.

have to work on that  :/ 

also have some issues with git ignore patterns , as i download the map data ( shp & geojson ) 
but i do not want to commit to repository. 

* solved

was related to rstudio thingie, y clicked staged , then updated the git ignore, and gotta unclick staged after that the file disapeaerd from the list.

## inspirations


## findings

well this week was kinda accidental ... so started with missings then time gone out. 

after that as was not happy with result, i started peeking others stuff on twitter & youtube.

- new tools that save a lot of time on missing data & initial exploring: 

the DataExplorer package.
https://twitter.com/dataandme/status/1059868547097264129

also take attention to natural segmented data as turbines(t_) Vs other as they might be usefull to make a subdataset and insights on them. ( as drob said)

## future works:

- use gganimate as proposed by @drob
https://www.youtube.com/watch?v=O1oDIQV6VKU
https://twitter.com/drob/status/1060191453874401280


- interesting tweets: 

super clean map + waffle plots
https://twitter.com/harrocyranka/status/1059810939929804800
https://github.com/hrbrmstr/waffle

cluster analysis + multiple maps:
https://twitter.com/TannerKoomar/status/1059868181488119808

gif + 
https://twitter.com/EvaMaeRey/status/1049797083060813824 

pretty barchart + highlighted bars:
https://twitter.com/ak88_est/status/1059927053716385792 

turbine characteristics + segmented by image source
https://twitter.com/DataVizJohansen/status/1059829569170104320

map wishlist:

https://twitter.com/k_j_norris/status/1060196410291875840
https://www.r-graph-gallery.com/327-chloropleth-map-from-geojson-with-ggplot2/
https://twitter.com/m_cadek/status/1060150590687318017
