---
title: "Tidy tuesday challenge: Week 2019-09-17 National Park Visits
author: "julio"
date: "2019-09-23"
output: html_document
---

# Tidy tuesday challenge: Week 2019-09-17 National Park Visits

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

just explore and plot something =) 

### Data:
 
 Week Week 2019-09-17 National Park Visits
 
### objectives:

explore & do something

## details:


## import data
```{r echo=FALSE,message=FALSE,warning=FALSE}
# library(magrittr)
library(tidyverse)
# library(ggrepel)
library(skimr)

library(here)
# library(dplyr)
# library(tidyr)
# library(readxl)
# library(janitor)
# library(stringr)
# library(lubridate)
library(geojsonsf)

library(geojsonio); #install.packages("geojsonio")

library(grid)
library(spatialEco)#; install.packages("spatialEco")
library(sp)
library(sf)# install.packages('sf')
# install.packages("Rcpp")
# remotes::install_github("tylermorganwall/rayshader")
# library(rayshader)

library(transformr)# install.packages('transformr')
library(gganimate)# install.packag
```


```{r echo=FALSE,message=FALSE,warning=FALSE}
park_visits <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-09-17/national_parks.csv")
# state_pop <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-09-17/state_pop.csv")
# gas_price <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-09-17/gas_price.csv")

```

too much data i'll just focus on parks

# explore data

```{r echo=FALSE,message=FALSE,warning=FALSE}
# now_data %>% purrr::map(now_data,head)
park_visits %>% head()


```

```{r echo=FALSE,message=FALSE,warning=FALSE}

park_visits %>% glimpse()



```

```{r echo=FALSE,message=FALSE,warning=FALSE}

park_visits %>% skimr::skim()


```

```{r echo=FALSE,message=FALSE,warning=FALSE}
DataExplorer::plot_missing(park_visits)

```

getting year/unitname/visitors/gnis_id

year for animation, unitname for name of park , visitors to see something, and gnis id to plot it.
1st gotta get the geojson. renamted to  usa_national_parks.geojson (got froM: https://data.world/codefordc/national-parks)

ok code not worked: 
just got some washington dc parks. after some cleaning. 
used: http://geojsonlint.com/ to validate parks
used: https://mapster.me/right-hand-rule-geojson-fixer/ to fix geojson.
used: https://gis.stackexchange.com/questions/259944/polygons-and-multipolygons-should-follow-the-right-hand-rule to ge tinformed about the righthand rule error

used: https://observablehq.com/@erincaughey/national-parks-geojson to get a working url with geojson

ok the geojson to sf not working with this geojson


ill try with leaflet ... 

::facepalm:: was just i got the path wrong :/

```{r echo=FALSE,message=FALSE,warning=FALSE}
 
# park_visits %>%  count(year)
usa_national_parks_geojson_file <- here::here("jas1_weeks","2019","2019-09-17","usa_national_parks.geojson")

usa_national_parks_original_geojson_file <- here::here("jas1_weeks","2019","2019-09-17","usa_national_parks_original.geojson")

# Error in rcpp_geojson_to_sf(geojson, expand_geometries) : Invalid JSON
usa_national_parks_geojson <- geojsonsf::geojson_sf(usa_national_parks_geojson_file) %>% 
    janitor::clean_names()
# 
# usa_national_parks_geojson_orig <- geojsonsf::geojson_sf(usa_national_parks_original_geojson_file)

# ok can join by unit name
usa_national_parks_geojson_sans_geom <- usa_national_parks_geojson %>% 
    sf::st_set_geometry(NULL)

simple_visits <- park_visits %>% 
    filter(year!="Total") %>% 
    mutate(year=as.integer(year)) %>% 
    select(year,unit_code,unit_name,visitors)

simple_visits_total_time <- simple_visits %>% 
    group_by(unit_name,unit_code) %>% 
    summarise(period_max=max(year,na.rm = TRUE),
              period_min=min(year,na.rm = TRUE),
              period_range=paste0(period_min,"-",period_max),
              total_visitors=sum(visitors,na.rm = TRUE))


simple_visits_total_time_total <- park_visits %>% 
    filter(year=="Total") %>% 
    select(unit_code,unit_name,visitors) %>% 
    group_by(unit_name,unit_code) %>% 
    summarise(period_max=max(year,na.rm = TRUE),
              period_min=min(year,na.rm = TRUE),
              period_range=paste0(period_min,"-",period_max),
              total_visitors=sum(visitors,na.rm = TRUE))

simple_visits_total_time_total %>% count(unit_code) %>%  filter(n>1)
usa_national_parks_geojson_sans_geom %>% count(unitco ) %>%  filter(n>1)

not_in_map <- usa_national_parks_geojson_sans_geom %>% 
    anti_join(simple_visits_total_time,by = c("unit_code"="unit_code")) %>% 
    select(unit_code,unit_code)

not_in_data <- simple_visits_total_time %>% 
    anti_join(usa_national_parks_geojson_sans_geom,
              by = c("unit_code"="UNIT_CODE")) %>% 
    select(unit_code,unit_name)

usa_national_parks_geojson %>% 
    left_join(simple_visits_total_time,by = c("UNIT_CODE"="unit_code"))

```

# simple stuff

ok, well got too much time out on this . ill go simple. 

was thinking on doing some plot with color acord to type of park 


ok, xml's also dont work
```{r echo=FALSE,message=FALSE,warning=FALSE}
fac_levels <- c("60+", "40+ - 60", "20+ - 40", "10+ - 20","5+ - 10","0 - 5") 
park_type_totals <- park_visits %>% 
    filter(year=="Total") %>% 
    group_by(unit_type) %>% 
    summarise(quantity=n(),visitors_total=sum(visitors,na.rm = TRUE),has_nas=any(is.na(visitors))) %>% 
    mutate(park_quant_bin=case_when(quantity> 60 ~ "60+",
                                    quantity> 40 & quantity<= 60 ~ "40+ - 60",
                                    quantity> 20 & quantity<= 40 ~ "20+ - 40",
                                    quantity> 10 & quantity<= 20 ~ "10+ - 20",
                                    quantity> 5 & quantity<= 10 ~ "5+ - 10",
                                    quantity<= 5 ~ "0 - 5")) %>% 
    mutate(park_quant_bin=factor(park_quant_bin,levels=fac_levels) ) %>% 
    mutate(unit_type=fct_reorder(unit_type,visitors_total))

    

    # factor(c("60+", "40+ - 60", "20+ - 40", "10+ - 20","5+ - 10","0 - 5"), 
    #    levels = c("60+", "40+ - 60", "20+ - 40", "10+ - 20","5+ - 10","0 - 5"))
```
```{r echo=FALSE,message=FALSE,warning=FALSE}    
    
park_type_c <- park_type_totals %>% 
    mutate(unit_type=fct_reorder(unit_type,visitors_total)) %>% 
    ggplot(aes(x=unit_type,y = log(visitors_total),fill=quantity))+
    geom_col()+
    coord_flip()+
    theme_light()+
    labs(title="USA Type of National Parks",
         subtitle = "visitors in log scale",
         x="",y="log visitors", fill="Park Quantity")
```
```{r echo=FALSE,message=FALSE,warning=FALSE}

plot_bkg <- "#3b95c0" # "#8e5017"
text_color_all <- "#f7e7c5"  #"#dff3f8" #"#a9d9e8" # "#7dcdc2"
text_color_all_with_nas <- "#fe9a99"  #"#dff3f8" #"#a9d9e8" # "#7dcdc2"

text_colours_vector <- if_else(park_type_totals$has_nas,text_color_all_with_nas,text_color_all)


# t1 <- textGrob(expression("visitors in log scale; " * phantom(bold("park with NA in COLOR"))),
#                x = 0.5, y = 1.1, gp = gpar(col = "black"))
# t2 <- textGrob(expression(phantom("visitors in log scale; ") * bold("park with NA in COLOR")),
#                x = 0.5, y = 1.1, gp = gpar(col = "#fe9a99"))

park_type_d <- park_type_totals %>% 
    ggplot(aes(x=unit_type,
               y = log(visitors_total),
               fill=park_quant_bin))+
    geom_col()+
    # geom_point(aes(color=has_nas))+
    scale_fill_brewer(palette = "BuGn",direction = -1)+
    coord_flip()+
    theme_light()+
    # annotation_custom(grobTree(t1, t2))+
    theme(text = element_text(colour = text_color_all),
          panel.grid=element_line(colour="#c18038"),
          axis.text.y = element_text(colour = text_colours_vector),
          axis.text.x = element_text(colour = text_color_all),
          legend.background = element_rect(fill = plot_bkg),
          legend.key = element_rect(fill ="#e1c083"),
          plot.background = element_rect(fill =plot_bkg),###8e5017
          panel.background = element_rect(fill ="#e1c083"))+
    labs(title="USA Type of National Parks",
         subtitle = "visitors in log scale; park with NA in Pink",
         x="",y="log visitors", fill="Park Quantity",
         caption="#TidyTuesday")

# park_type_d
ggsave(filename = "usa_national_park_types_d.png",plot = park_type_d,width = 7,height = 6)

```


## items as absolute values



#tweet: 

2019-09-23 #TidyTuesday #rstats National Park Visits !
Some simple plot with custom colors brougth from several scales of color brewer.

code: https://github.com/jas1/tidytuesday/tree/master/jas1_weeks/2019/2019-09-17


# communicate

too much time experimenting :p 
with geojson and some tiem with ggplot.

here tweet: https://twitter.com/jspairani/status/1175980540576620544

# resources: 

ok code not worked: 
just got some washington dc parks. after some cleaning. 
used: http://geojsonlint.com/ to validate parks
used: https://mapster.me/right-hand-rule-geojson-fixer/ to fix geojson.
used: https://gis.stackexchange.com/questions/259944/polygons-and-multipolygons-should-follow-the-right-hand-rule to ge tinformed about the righthand rule error
used: https://observablehq.com/@erincaughey/national-parks-geojson to get a working url with geojson
used: https://gist.githubusercontent.com/erincaughey/2f221501645757e28b715c4063e87595/raw/a90be1b434b1a8cdf71c2abc3373ca63987e2d23/nps-geo-boundary.json to get the geojson.

