---
title: "Week 20181002"
author: "_jas"
date: "`r format(Sys.Date())`"
output: html_document
---

# Tidy tuesday challenge: Week 2018-10-02


## keep it simple:

### Objectives: 

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate

https://rviews.rstudio.com/2017/06/08/what-is-the-tidyverse/ 

# Workflow:

## Import

```{r echo=FALSE,message=FALSE,warning=FALSE}
library(here)
library(tibble)
library(dplyr)
library(tidyr)
library(skimr)#install.packages('skimr')
```

## Tidy

```{r echo=FALSE,message=FALSE,warning=FALSE}
us_births_file_location <- here('data','2018-10-02','us_births_2000-2014.csv')
us_births <- read.csv(us_births_file_location,stringsAsFactors = FALSE)

us_births %>% glimpse() %>% skim()
```

looks like no missing data

## Understand ( iteration 1 )

### Transform

Adding a date field

```{r echo=FALSE,message=FALSE,warning=FALSE}
library(lubridate)
library(stringr)
us_births_2 <-  us_births %>% 
    mutate(date=ymd(paste0(
        year,
        str_pad(month,2,'left','0'),
        str_pad(date_of_month,2,'left','0'))))

us_births_2 %>% glimpse() %>% skim()
```

### Visualize

```{r echo=FALSE,message=FALSE,warning=FALSE}
library(ggplot2)

date_of_month_13 <- us_births_2 %>% 
    filter(date_of_month==13) %>% 
    pull(date)

us_births_2 %>%
    ggplot(aes(x=date,y=births)) +
    geom_line() + 
    geom_vline(xintercept = date_of_month_13, colour="red") +
    theme(axis.text.x = element_text(
        angle = 90, 
        hjust = 1))

```



### Model

--

## Communicate

* gotta review ggplot & faceting. 
year / month vline , may be joy division like plots.
