---
title: "Week 20181002"
author: "_jas"
date: "`r format(Sys.Date())`"
output: html_document
---

# Tidy tuesday challenge: Week 2018-10-02


## keep it simple:

### Objectives: 

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate

https://rviews.rstudio.com/2017/06/08/what-is-the-tidyverse/ 

# Workflow:

## Import

```{r echo=FALSE,message=FALSE,warning=FALSE}
library(here)
library(readr)
library(tibble)
library(dplyr)
library(tidyr)
library(skimr)#install.packages('skimr')
```

## Tidy

```{r echo=FALSE,message=FALSE,warning=FALSE}
us_births_file_location <- here('data','2018-10-02','us_births_2000-2014.csv')
us_births <- readr:::read_csv(us_births_file_location)

us_births %>% glimpse() %>% skim()
```

looks like no missing data

## Understand ( iteration 1 )

### Transform

Adding a date field

```{r echo=FALSE,message=FALSE,warning=FALSE}
library(lubridate)
library(stringr)
us_births_2 <-  us_births %>% 
    mutate(date= lubridate:::make_date(year,month,date_of_month)) %>% 
    mutate(date_of_month_str=as.factor(date_of_month)) %>% 
    mutate(day_of_week_lbl=wday(date, label=TRUE)) %>% 
    mutate(is_friday_13=if_else(date_of_month==13 & day_of_week==6,TRUE,FALSE))

us_births_2 %>% glimpse() %>% skim()
```

Just 25 friday 13 over 5454 days

### Visualize

```{r echo=FALSE,message=FALSE,warning=FALSE}
library(ggplot2)
library(ggridges)#; install.packages("ggridges")

date_of_month_13 <- us_births_2 %>% 
    filter(is_friday_13) %>% 
    pull(date)

```

To get an idea of how much friday 13th we have in all dataset ( 25 / 5454 )

```{r echo=FALSE,message=FALSE,warning=FALSE}

us_births_2 %>%
    ggplot(aes(x=date,y=births)) +
    geom_line() + 
    geom_vline(xintercept = date_of_month_13, colour="red") +
    theme(axis.text.x = element_text(
        angle = 90, 
        hjust = 1))
```


To see distributions over time in each month day:


```{r echo=FALSE,message=FALSE,warning=FALSE}

us_births_2 %>% 
    ggplot(
    aes(x = births, 
        y = date_of_month_str, 
        height = ..density..,fill=is_friday_13)) +
  geom_density_ridges(stat = "density",alpha = 0.5)
```

To see distributions form another point of view:

```{r echo=FALSE,message=FALSE,warning=FALSE}
us_births_2 %>% ggplot( aes(y = births, 
        x = date_of_month_str, fill=is_friday_13)) + 
    geom_boxplot()

```

### Model

--

## Understand ( iteration 2 )

will be try to put networks point of view on each data set. 
sometimes can be difficult ... but thats the challenge :p 

this time can be: 

a bipartite graph of:

date of month - day of week

and in the relation the value of births in this case.


### Transform

making the bipartite graph , i need the edgelist.

```{r echo=FALSE,message=FALSE,warning=FALSE}
library(igraph)
library(RColorBrewer)
library(visNetwork)

edgelist <- us_births_2 %>% 
    mutate(day_of_week_lbl_2 = case_when(
        day_of_week == 1 ~ "Sunday",
        day_of_week == 2 ~ "Monday",
        day_of_week == 3 ~ "Tuesday",
        day_of_week == 4 ~ "Wednesday",
        day_of_week == 5 ~ "Thursday",
        day_of_week == 6 ~ "Friday",
        day_of_week == 7 ~ "Saturday"
    )) %>%  
    group_by(date_of_month,day_of_week_lbl_2,day_of_week) %>% 
    summarise(births_mean=mean(births),
              births_amount=sum(births),
              month_week_day_count=n(),
              scaled_mean=sc) 

edgelist %>% arrange(births_mean)

graph_transform <- igraph:::graph_from_data_frame(edgelist,directed = FALSE)
type_day_of_week <- edgelist %>% pull(day_of_week_lbl_2) 

igraph:::V(graph_transform)$type <- igraph:::V(graph_transform)$name %in% type_day_of_week

igraph:::V(graph_transform)$color <- if_else(igraph:::V(graph_transform)$type,"#fb9a99","#aadda8")

# igraph:::is.bipartite(graph_transform)

# range(igraph:::E(graph_transform)$births_mean)
# hist(igraph:::E(graph_transform)$births_mean)
# edge_attr_names(graph_transform)
bin_3_colors <- brewer.pal(3,"BrBG")

edge_color <- data.frame(bmean=igraph:::E(graph_transform)$births_mean) %>% as_tibble()
edge_color <- edge_color %>% 
    mutate(color = case_when(
        bmean < 9500 ~ bin_3_colors[1],
        bmean >= 9500  & bmean < 11500  ~ bin_3_colors[2],
        bmean >= 11500  ~ bin_3_colors[3]
) ) %>%  
    mutate(title=paste0( "Mean: ",bmean))

igraph:::E(graph_transform)$color <- edge_color$color
igraph:::E(graph_transform)$title <- edge_color$title




visNetwork::visIgraph(graph_transform) %>% 
    visNetwork::visIgraphLayout(layout = "layout_as_bipartite") 

```


### Visualize

### Model

--

## Communicate

As can be seen , comparing to other days the fryday 13th birth hypotesis is valid.

### future work: 

