
---
title: "Tidy tuesday challenge: Week 2019-07-23 FAA Wildlife Strike Database"
author: "julio"
date: "2019-06-25"
output: html_document
---

# Tidy tuesday challenge: Week 2019-07-23 FAA Wildlife Strike Database

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

### Data:
 
2019-07-23 FAA Wildlife Strike Database
 
### objectives:

check the data and do ranking

## details:


## import data
```{r echo=FALSE,message=FALSE,warning=FALSE}
library(magrittr) # para el %T>%
library(tidyverse)
# library(sf)
library(dplyr)
library(stringr)#;
# library(rebus)#; install.packages('rebus')
# library(tidytext)
library(prophet)


# install.packages("Rcpp")
# remotes::install_github("tylermorganwall/rayshader")
# library(rayshader)
library(lubridate)
library(ggforce)
library(ggrepel)

```


```{r echo=FALSE,message=FALSE,warning=FALSE}
wildlife_impacts <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-23/wildlife_impacts.csv")
```

# explore data


```{r echo=FALSE,message=FALSE,warning=FALSE}
wildlife_impacts %>% head()

```

```{r echo=FALSE,message=FALSE,warning=FALSE}
wildlife_impacts %>% glimpse()

```

```{r echo=FALSE,message=FALSE,warning=FALSE}
wildlife_impacts %>% skimr::skim()


```


# time line of messages_posted
```{r echo=FALSE,message=FALSE,warning=FALSE}
# wildlife_impacts %>% incident_year,airport

        metric_sym <-rlang::sym("airport")
        ranking <- seq(from=1,to = 5,by = 1)
        ret <- wildlife_impacts %>% 
            count(!!!metric_sym) %>% 
            arrange(desc(5)) %>% 
            head(cant) %>% 
            mutate(rank=ranking) %>% 
            mutate(metric_value= !!metric_sym)

metric_str <- "airport"#"airport" , "species"
top_n_num <- 5
resu <- wildlife_impacts %>% 
nest(-incident_year) %>% 


    mutate(top_n_year=purrr::map(data,
                                    n_top=top_n_num,
                                    metric=metric_str,
                                    .f=function(d,metric,n_top){
        metric_sym <- rlang::sym(metric)
        ranking <- seq(from=1,to = n_top,by = 1)
        ret <- d %>% 
            filter(airport!="UNKNOWN") %>% 
            count(!!!metric_sym) %>% 
            arrange(desc(n)) %>% 
            head(n_top) %>% 
            mutate(rank=ranking)
        ret
    })) %>% unnest(top_n_year)

glimpse(resu)
# metrica_sym <-rlang::sym(metrica_s)
# plot con ranking
plot_out <- resu %>% 
    mutate(airport=fct_reorder(airport,rank,.desc = TRUE)) %>% 
    mutate(rank=as.factor(rank)) %>% 
    ggplot(aes(x=as.factor(incident_year),y=airport,color=rank,group=rank,label=n))+
    geom_point()+
    geom_line(show.legend = FALSE)+
    theme_light()+
    theme(axis.text.x = element_text(angle = 90, hjust = 1),
          axis.text.y = element_text(size = 6))+
    labs(x="",y="", 
         title="Top 5 USA Airports with Wildlife Strikes",
         subtitle="From 1990 to 2018, Airport: Unknown removed.",
         caption="#TidyTuesday")

plot_out

ggsave(filename = "top5_airports.png",plot = plot_out)

```

#tweet: 

2019-07-23 #TidyTuesday #rstats FAA Wildlife Strike Database ! 
Top 5 Ranking across the years and how switched places.
Unknown airport removed.
Enjoy :D


# communicate

ranking recicled & adapted from other stuff done in weekend :)








