---
title: "Tidy tuesday challenge: Week 2019-02-19  PhDs Awarded by Field"
author: "julio"
date: "24 de febrero de 2019"
output: html_document
---

# Tidy tuesday challenge: Week 2019-02-19 PhDs Awarded by Field

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

### Data:

this week data its related to PhDs Awarded by Field

### objectives:

- other different challenge: a more technical stuff: do the stand alone shiny app using electron.

## details:

this week my personal challenge is different from the other DS challenges, this is more technical may be more DE challenge... 

Some times making a dashboard do not have sense unless your public can use it, and some times the " default steps " are not enough. 

thats why im trying to use this one; when you need your users to get a " usable dashboard" and cannot host on shinyapps, or an internal shiny server, or other internet shiny stuff.

you need old school desktop app. using that implies that youll need the data inside the app. 
( download the data and put it inside your app.)

In this case i'll do the simple " embbed the data in the app " case. 
Just to take in account: Some times you just cant do that, and you need to connect to a DB or the internet. In that cases you should evaluate connection permissions, firewalls, network acceses, proxies, proper urls , proper ports, and may be some other stuff; and all that specific stuff got a lot of complexities themselves.

about the dashboard take a look at the tweet.
A ton of this effort has been done by other people, you can check the resources at the end.

I like hierarchical data its always awsome to visualize and play with it; ill use a guy shiny app of the week so i can really focus on the electron side of the thing.

## import data

```{r echo=FALSE,message=FALSE,warning=FALSE}
library(dplyr)
# library(Rcpp) #; install.packages("Rcpp")
library(skimr) #; install.packages("skimr")
# library(circlepackeR) #; devtools::install_github("jeromefroe/circlepackeR")
# library(data.tree)  #; install.packages("data.tree")
```



```{r echo=FALSE,message=FALSE,warning=FALSE}
phd_file <- here::here("data","2019","2019-02-19","phd_by_field.csv")
phd_raw <- readr::read_csv(phd_file)

phd_processed <- phd_raw



```


```{r echo=FALSE,message=FALSE,warning=FALSE}
glimpse(phd_raw) 


```

```{r echo=FALSE,message=FALSE,warning=FALSE}
skimr::skim(phd_raw) 


```

## tidy

always wante to try circle pack, so there we go: prepare data: deprecated -- shifted to the shiny app of the guy as i wan to have a functional deployable that i can give other people.
not just my pc.

```{r echo=FALSE,message=FALSE,warning=FALSE}
# library(circlepackeR) #; devtools::install_github("jeromefroe/circlepackeR")
# 
# phd_processed <- phd_raw %>% filter(year==2008) %>% select(-year) %>%  
#         mutate(path_string = paste("phds", broad_field, major_field, field, sep = "/"))
# # sample: http://shiny.rstudio.com/gallery/submitbutton-demo.html
# # UI
# 
# # library(data.tree)
# # phd_processed_circler <- data.tree::as.Node(phd_processed)
# 
# 
# phd_processed %>% count(field) %>%  filter(n>1)
# 
# phd_processed %>% filter(field=="Environmental toxicologyc")
# 
# circlepackeR::circlepackeR(phd_raw)
# 
# 
# 
# phd_processed2 <- phd_raw %>% filter(year==2008) %>% select(-year) %>% 
#     mutate(from=) %>% mutate( to="")
# 
# 
# # Libraries
# library(ggraph)
# library(igraph)
# library(tidyverse)
# library(viridis)
#  
# # We need a data frame giving a hierarchical structure. Let's consider the flare dataset:
# edges=flare$edges
# vertices = flare$vertices
# mygraph <- graph_from_data_frame( edges, vertices=vertices )
#  
# # Control the size of each circle: (use the size column of the vertices data frame)
# # png("~/Dropbox/R_GG/R_GRAPH/#314_custom_circle_packing1.png", height = 480, width=480)
# ggraph(mygraph, layout = 'circlepack', weight="size") + 
#   geom_node_circle() +
#   theme_void()


```

## visualize

as i want interactive for the shiny app im using package: circlepackeR: canceled.
using the guy shiny app as ive made some stuff with that library before should not be complex.

```{r echo=FALSE,message=FALSE,warning=FALSE}
# library(circlepackeR)
# library(data.tree)
# library(treemap)#;install.packages("treemap"); install.packages("httpuv"); install.packages("mime")
# 
# data(GNI2014)
# head(GNI2014)
# 
# GNI2014$pathString <- paste("world", 
#                             GNI2014$continent, 
#                             GNI2014$country, 
#                             sep = "/")
# population <- as.Node(GNI2014)
# 
# circlepackeR(population, size = "population", color_min = "hsl(56,80%,80%)", 
#              color_max = "hsl(341,30%,40%)")


```


## communicate


### general view:

started wanting to analize something simple like a circle pack , as it needed some preparation and wanted focus on the electron app , just switched to this twit app: https://twitter.com/theotheredgar/status/1097955862235279360

that final app resulted on ~250MB directory, zipped with 7z , becomes: 182MB

For practical reasons i wont be uploading the electron app result here, as its quite heavy.

### summary: 

you install git , you install nodejs ( that comes with npm), then you can make an electron app. 
inside that thing you can embedd other aplications, like ... an R-portable; thats one reason of heavy MB.
Then for your specific app, depending on the app " fanciness", you probably use several libraries.
each library , use another bunch of MB. 

and well the electron stuff itself.


Reasons why you might want to do this: 

https://www.youtube.com/watch?v=ARrbbviGvjc

the other usefull video is: 
https://www.youtube.com/embed/O56WR-yQFC0

### full story:

i suggest going the videos 1st, then start coding because you might lose a ton of time searching for stuff that is there. 

Also the video its on old version of R, and old version of Electron. There are not much differencies. 
I spend some time looking the new version empty app, vs the repo app, and looking at the structure of files.

before trying the demo app, i installed all the packages needed for my app on the R-portable instance inside the electron application. ( like installing shiny and dependencies )

the installation was not done through the " install.packages " but to unzipping the libraries on the R-portable library folder: "../R-Portable/App/R-Portable/library"
for example the shiny folder will be: 

"electron-quick-start/R-Portable/App/R-Portable/library/shiny"

After i get a demo app going in electron, i switch to the twit app. 
(demo app: https://shiny.rstudio.com/gallery/submitbutton-demo.html)


1st, tryed as a simple shiny app app.r  ( changed the flex dashboard format, as i dont like it much.)
after saw it working, put it on the electron application. 
this can be seen on my repo as this week app.r

then tryed to run the application as i've done with the demo app; i knew that it wont work as some dependencies might be miss , so some debugging dependencies was done in a really empirical way. 
( run app > check error > go download the zip > install on the library of R-portable)

worked on that till i have a really nasty problem with a library, that it wasnt found the "vI" 
as seen that the problem was really ... "hairy" : https://community.rstudio.com/t/not-able-to-load-tidyverse-object-vi-not-found/10243/7

I just stepped aside from that and made a little change, using data tables instead of Formatted table.
after that life was easier and got my "more than basic" application running.

now to the packaging.
the specific line is on this video: https://www.youtube.com/embed/O56WR-yQFC0

After a while got my "uber folder" whit ton of stuff and my dear "exe" ( doing this on windows )

zipping the folder was next, 
and finally , test the thing on other computer, as I had none near, used a virtual machine that do not have R installed nor nodejs.

and it : worked like charm :D (check the screenshot :D 20190224_tidy_tuesday_electron_app.png )




### to close ...:

- heavy output ( big folder to giv to other ppl)
- can be usefull on some circumstances
- tech knowladge is definitly a plus here; will get you quicker to the places.
- can contact on twitter or github issue if need help


### useful resources:

- https://twitter.com/theotheredgar/status/1097955862235279360
- https://github.com/ksasso/useR_electron_meet_shiny
- https://www.youtube.com/watch?v=ARrbbviGvjc
- https://www.youtube.com/embed/O56WR-yQFC0
- https://github.com/ksasso/Electron_ShinyApp_Deployment
- https://sourceforge.net/projects/rportable/
