---
title: "Tidy tuesday challenge: Week 2019-05-21 plastic waste"
author: "julio"
date: "21 - 5 - 2019"
output: html_document
---

# Tidy tuesday challenge: Week 2019-05-21 plastic waste

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

### Data:

Global Plastic Waste our world in data

### objectives:

- explore & do a quick one , short time this week

## details:


## import data
```{r echo=FALSE,message=FALSE,warning=FALSE}
library(dplyr)
library(tidyr)
library(skimr)
library(readr) 
library(ggplot2)
library(scales)
library(forcats)
library(lubridate)
library(gganimate)
library(animation)

library(stringr)#;
library(rebus)#; install.packages('rebus')
library(tidytext)
# 
# library(rworldxtra) # para datos iso3, segun gabo
# library(sf)

# devtools::install_github('rensa/ggflags')
library(ggflags) # https://github.com/rensa/ggflags

# library(widyr) # for pairwise core
# library(ggraph) # for graph drawing
# library(igraph)  # for graph managing

# install.packages("waffle")
# devtools::install_github("hrbrmstr/waffle")
library(waffle)
library(countrycode)
```


```{r echo=FALSE,message=FALSE,warning=FALSE}
coast_vs_waste <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-05-21/coastal-population-vs-mismanaged-plastic.csv")

coast_vs_waste <- coast_vs_waste %>% janitor::clean_names()
```

# explore data

```{r echo=FALSE,message=FALSE,warning=FALSE}
glimpse(coast_vs_waste)

```


```{r echo=FALSE,message=FALSE,warning=FALSE}
coast_vs_waste %>% skimr::skim()

# nobel_winner_all_pubs$
```

# Plastic waste vs Coast population in South America

```{r echo=FALSE,message=FALSE,warning=FALSE}

# ?geom_waffle
# coast_vs_waste %>% 
#   filter(year >= 2010) %>% 
#   count(year, status) -> storms_df



countries_continent <- countrycode::codelist %>%  
    select(continent,iso3c,iso2c,country.name.en) %>% 
    mutate(iso2c=str_to_lower(iso2c))

dataset_to_plot <- coast_vs_waste %>% 
    left_join(countries_continent,by=c("code"="iso3c")) %>% 
    filter(!is.na(continent)) %>% 
    filter(iso2c %in% c('ar','br','cl','co','ec','gf','sr','gy','py','pe','uy','ve')) %>% 
    filter(year>=1950) %>% 
    group_by(continent,code,iso2c) %>% 
    summarise(log_total_coastal_population=log(sum(coastal_population,na.rm = TRUE)+1),
              log_total_mismanaged_plastic_waste_tonnes=log(sum(mismanaged_plastic_waste_tonnes,na.rm = TRUE)+1)) %>% 
    filter(log_total_coastal_population!=0)

# countries_continent %>% filter(iso2c=="gf")
flag_data <- dataset_to_plot %>% select(iso2c,log_total_coastal_population,log_total_mismanaged_plastic_waste_tonnes)

plot_plastic_coast_population <- dataset_to_plot %>% 
    ggplot(aes(x=log_total_coastal_population,
               y=log_total_mismanaged_plastic_waste_tonnes
               #color=continent))+
    ))+
    # geom_hex(alpha=0.5) +#+
    scale_fill_viridis_c()+
    # geom_tile()
    geom_point()+
    geom_text(data = flag_data,aes(label = iso2c),nudge_y = 0.3,alpha=0.75,colour="#6c6c6c")+
    geom_flag(data = flag_data,aes(country = iso2c), alpha=0.5)+
    theme_light()+
    labs(title="Plastic waste vs Coast population in South America",
         subtitle="log scaled, total sum 1950-2013",
         caption="#TidyTuesday",
         y="Total mismanaged plastic waste(log)",
         x="Total Coast population(log)")
    # facet_grid(.~continent)

ggsave(plot_plastic_coast_population,filename = "plastic_coast_pop.png")

```
# Waffles with publications

```{r}


nobel_arts_waffle_df_2 <- nobel_winner_all_pubs %>% 
    mutate(decade=floor(pub_year/10)*10) %>% 
    count(category,decade) %>% 
    filter(complete.cases(.)) %>% arrange(decade)

# nobel_arts_waffle_df <- nobel_winner_all_pubs %>% 
#     count(category) %>% 
#     mutate(n=n/1000)

# dim(nobel_winner_all_pubs %>% count(category))    
nobel_arts_waffle_df_2 %>% 

    ggplot(aes(x=decade,colour=category,fill=category,y=n, values = n)) + 
    geom_jitter()+
    facet_grid(category~.)+
    ggthemes::scale_fill_tableau(name=NULL)

```

# # Plastic waste vs Coast line km in South America

```{r}
#https://en.wikipedia.org/wiki/List_of_countries_by_length_of_coastline

 # Brazil 	16 	7,491 	9 	33,379 	8,459,417 	0.886 	3.95
 # Argentina 	24 	4,989 	30 	8,397 	2,736,690 	1.82 	3.07
 # Uruguay 	85 	660 	98 	1,096 	175,015 	3.77 	6.26
 # Chile 	19 	6,435 	5 	78,563 	743,812 	8.65 	106
 # Peru 	48 	2,414 	55 	3,362 	1,279,996 	1.89 	2.63
 # Venezuela 	39 	2,800 	37 	6,762 	882,050 	3.17 	7.67
 # Colombia 	36 	3,208 	41 	5,875 	1,038,700 	3.09 	5.66
 # Ecuador 	51 	2,237 	47 	4,597 	276,841 	8.08 	16.6
 # Suriname 	107 	386 	114 	620 	156,000 	2.47 	3.97
 # Guyana 	98 	459 	95 	1,154 	196,849 	2.33 	5.86
 # French Guiana 			— 	763
# World Resources Institute values only + country & iso2 for joining
wri_coast_table <- tibble::tribble(
           ~country, ~iso2, ~coast_km,
           "Brazil",  "br",     33379,
        "Argentina",  "ar",      8397,
          "Uruguay",  "uy",      1096,
            "Chile",  "cl",     78563,
             "Peru",  "pe",      3362,
        "Venezuela",  "ve",      6762,
         "Colombia",  "co",      5875,
          "Ecuador",  "ec",      4597,
         "Suriname",  "sr",       620,
           "Guyana",  "gy",      1154,
    "French Guiana",  "gf",       763
    )
data_plot_2 <- dataset_to_plot %>% left_join(wri_coast_table,by = c("iso2c"="iso2"))
flag_data_2 <- data_plot_2 %>% select(iso2c,coast_km,log_total_coastal_population,log_total_mismanaged_plastic_waste_tonnes)

plot_plastic_coast_km <- data_plot_2 %>% 
    ggplot(aes(x=log(coast_km),
               y=log_total_mismanaged_plastic_waste_tonnes
               #color=continent))+
    ))+
    # geom_hex(alpha=0.5) +#+
    scale_fill_viridis_c()+
    # geom_tile()
    geom_point()+
    geom_text(data = flag_data_2,aes(label = iso2c),nudge_y = 0.3,alpha=0.75,colour="#6c6c6c")+
    geom_flag(data = flag_data_2,aes(country = iso2c))+
    theme_light()+
    labs(title="Plastic waste vs Coast Km in South America",
         subtitle="log scaled, total sum 1950-2013",
         caption="#TidyTuesday",
         y="Total mismanaged plastic waste(log)",
         x="Total Coast Km (log)")
ggsave(plot_plastic_coast_km,filename = "plastic_coast_km.png")

```

```{r}

nobel_arts_waffle_df <- nobel_winner_all_pubs %>% 
    mutate(decade=floor(pub_year/10)*10) %>% 
    count(category,decade) %>% 
    filter(complete.cases(.)) %>% arrange(decade)

nobel_arts_waffle_df %>% 
    filter(decade <1900 )%>% 
ggplot( aes(fill = category, values = n)) + 
  geom_waffle(color = "white", size = .25, n_rows = 10, flip = TRUE) +
  facet_wrap(~decade, nrow = 1, strip.position = "bottom") +
  scale_x_discrete() + 
  scale_y_continuous(labels = function(x) x * 10, # make this multiplyer the same as n_rows
                     expand = c(0,0)) +
  ggthemes::scale_fill_colorblind(name=NULL) +
  coord_equal() +
  labs(
    title = "Nobel prizes - articles - before 1900",
    subtitle = "#tidytuesday 2019-05-14",
    x = "",
    y = "#"
  ) +
  theme_minimal() +
  theme(panel.grid = element_blank(), axis.ticks.y = element_line()) +
  guides(fill = guide_legend(reverse = TRUE))
    
```

```{r}

# install.packages("ggbeeswarm")
# library("ggbeeswarm")
# 
# nobel_arts_waffle_df %>% 
#     # filter(decade >=1900  )%>% 
#     filter(decade <1950  )%>% 
#     ggplot( aes( x = decade, y = n, fill = category))+
#         geom_beeswarm(aes(color = category))
nobel_arts_waffle_df %>% 
    ggplot( aes( x = decade, y = n, fill = category))+
    geom_col()+
    scale_x_discrete() 


nobel_arts_waffle_df %>% count(category)

# 
# nobel_arts_waffle_df %>% 
#     filter(decade >=1900  )%>% 
#     filter(decade <1950  )%>% 
# ggplot( aes(x=fill = category, values = n)) + 
#   # geom_waffle(color = "white", size = .25, n_rows = 10, flip = TRUE) +
# geom_point()
#   facet_wrap(~decade, nrow = 1, strip.position = "bottom") +
#   scale_x_discrete() + 
#   scale_y_continuous(labels = function(x) x * 10, # make this multiplyer the same as n_rows
#                      expand = c(0,0)) +
#   ggthemes::scale_fill_tableau(name=NULL) +
#   coord_equal() +
#   labs(
#     title = "Nobel prizes 1900 - 1950",
#     subtitle = "#tidytuesday 20190514",
#     x = "",
#     y = "#"
#   ) +
#   theme_minimal() +
#   theme(panel.grid = element_blank(), axis.ticks.y = element_line()) +
#   guides(fill = guide_legend(reverse = TRUE))
```


# communicate

well not best week not best times ... just keep rowing.x2



