---
title: "Tidy tuesday challenge: Week 2019-05-14 Nobel prize"
author: "julio"
date: "19 - 5 - 2019"
output: html_document
---

# Tidy tuesday challenge: Week 2019-05-14 Nobel Prize

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

### Data:

UNESCO Student Teacher Ratio

### objectives:

- explore and decide

## details:


## import data
```{r echo=FALSE,message=FALSE,warning=FALSE}
library(dplyr)
library(tidyr)
library(skimr)
library(readr) 
library(ggplot2)
library(scales)
library(forcats)
library(lubridate)
library(gganimate)
library(animation)

library(stringr)#;
library(rebus)#; install.packages('rebus')
library(tidytext)
# 
# library(rworldxtra) # para datos iso3, segun gabo
# library(sf)

# devtools::install_github('rensa/ggflags')
# library(ggflags) # https://github.com/rensa/ggflags

# library(widyr) # for pairwise core
# library(ggraph) # for graph drawing
# library(igraph)  # for graph managing

```


```{r echo=FALSE,message=FALSE,warning=FALSE}
nobel_winners <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-05-14/nobel_winners.csv")
nobel_winner_all_pubs <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-05-14/nobel_winner_all_pubs.csv")
```

# explore data

```{r echo=FALSE,message=FALSE,warning=FALSE}
glimpse(nobel_winners)
glimpse(nobel_winner_all_pubs)

# bird_collisions %>% count(family,genus,species,sort = TRUE)

```


```{r echo=FALSE,message=FALSE,warning=FALSE}
nobel_winners %>% skimr::skim()

# nobel_winner_all_pubs$
```

# switched to WAffles ! :D  trying default
```{r echo=FALSE,message=FALSE,warning=FALSE}
# install.packages("waffle")
# devtools::install_github("hrbrmstr/waffle")
library(waffle)
# ?geom_waffle
storms %>% 
  filter(year >= 2010) %>% 
  count(year, status) -> storms_df

# dim(storms_df)

ggplot(storms_df, aes(fill = status, values = n)) + 
  geom_waffle(color = "white", size = .25, n_rows = 10, flip = TRUE) +
  facet_wrap(~year, nrow = 1, strip.position = "bottom") +
  scale_x_discrete() + 
  scale_y_continuous(labels = function(x) x * 10, # make this multiplyer the same as n_rows
                     expand = c(0,0)) +
  ggthemes::scale_fill_tableau(name=NULL) +
  coord_equal() +
  labs(
    title = "Faceted Waffle Bar Chart",
    subtitle = "{dplyr} storms data",
    x = "Year",
    y = "Count"
  ) +
  theme_minimal(base_family = "Roboto Condensed") +
  theme(panel.grid = element_blank(), axis.ticks.y = element_line()) +
  guides(fill = guide_legend(reverse = TRUE))

```
# Waffles with publications

```{r}


nobel_arts_waffle_df_2 <- nobel_winner_all_pubs %>% 
    mutate(decade=floor(pub_year/10)*10) %>% 
    count(category,decade) %>% 
    filter(complete.cases(.)) %>% arrange(decade)

# nobel_arts_waffle_df <- nobel_winner_all_pubs %>% 
#     count(category) %>% 
#     mutate(n=n/1000)

# dim(nobel_winner_all_pubs %>% count(category))    
nobel_arts_waffle_df_2 %>% 

    ggplot(aes(x=decade,colour=category,fill=category,y=n, values = n)) + 
    geom_jitter()+
    facet_grid(category~.)+
    ggthemes::scale_fill_tableau(name=NULL)

```

```{r}

nobel_wins_waffle_df <- nobel_winners %>% 
    mutate(decade=floor(prize_year/10)*10) %>% 
    count(category,decade) %>% 
    mutate(tenths=n/10)

rows_multiplier <- 10

waffle_save <- nobel_wins_waffle_df %>% 
ggplot( aes(fill = category, values = n)) + 
  geom_waffle(color = "white", size = .25, n_rows = rows_multiplier, flip = FALSE) +
      facet_wrap(~decade, nrow = 2, strip.position = "bottom") +

  scale_x_discrete() + 
  scale_y_continuous(labels = function(x) x * rows_multiplier, # make this multiplyer the same as n_rows
                     expand = c(0,0)) +
  # ggthemes::scale_fill_(name=NULL) +
    scale_fill_brewer(type="qual",palette = "Accent")+
  coord_equal() +
  labs(
    title = "Nobel Prizes Winners by category and decade",
    caption = "#tidytuesday 2019-05-14",
    x = "",
    y = "#"
  ) +
  theme_minimal() +
    # coord_flip()+
  theme(panel.grid = element_blank(), axis.ticks.y = element_line()) +
  guides(fill = guide_legend(reverse = TRUE))

# ggsave(filename = waffle_save,plot="waffle_winners_categry_date.png")
    ggsave(plot = waffle_save,filename = "waffle_winners_categry_date.png")

```

```{r}

nobel_arts_waffle_df <- nobel_winner_all_pubs %>% 
    mutate(decade=floor(pub_year/10)*10) %>% 
    count(category,decade) %>% 
    filter(complete.cases(.)) %>% arrange(decade)

nobel_arts_waffle_df %>% 
    filter(decade <1900 )%>% 
ggplot( aes(fill = category, values = n)) + 
  geom_waffle(color = "white", size = .25, n_rows = 10, flip = TRUE) +
  facet_wrap(~decade, nrow = 1, strip.position = "bottom") +
  scale_x_discrete() + 
  scale_y_continuous(labels = function(x) x * 10, # make this multiplyer the same as n_rows
                     expand = c(0,0)) +
  ggthemes::scale_fill_colorblind(name=NULL) +
  coord_equal() +
  labs(
    title = "Nobel prizes - articles - before 1900",
    subtitle = "#tidytuesday 2019-05-14",
    x = "",
    y = "#"
  ) +
  theme_minimal() +
  theme(panel.grid = element_blank(), axis.ticks.y = element_line()) +
  guides(fill = guide_legend(reverse = TRUE))
    
```

```{r}

# install.packages("ggbeeswarm")
# library("ggbeeswarm")
# 
# nobel_arts_waffle_df %>% 
#     # filter(decade >=1900  )%>% 
#     filter(decade <1950  )%>% 
#     ggplot( aes( x = decade, y = n, fill = category))+
#         geom_beeswarm(aes(color = category))
nobel_arts_waffle_df %>% 
    ggplot( aes( x = decade, y = n, fill = category))+
    geom_col()+
    scale_x_discrete() 


nobel_arts_waffle_df %>% count(category)

# 
# nobel_arts_waffle_df %>% 
#     filter(decade >=1900  )%>% 
#     filter(decade <1950  )%>% 
# ggplot( aes(x=fill = category, values = n)) + 
#   # geom_waffle(color = "white", size = .25, n_rows = 10, flip = TRUE) +
# geom_point()
#   facet_wrap(~decade, nrow = 1, strip.position = "bottom") +
#   scale_x_discrete() + 
#   scale_y_continuous(labels = function(x) x * 10, # make this multiplyer the same as n_rows
#                      expand = c(0,0)) +
#   ggthemes::scale_fill_tableau(name=NULL) +
#   coord_equal() +
#   labs(
#     title = "Nobel prizes 1900 - 1950",
#     subtitle = "#tidytuesday 20190514",
#     x = "",
#     y = "#"
#   ) +
#   theme_minimal() +
#   theme(panel.grid = element_blank(), axis.ticks.y = element_line()) +
#   guides(fill = guide_legend(reverse = TRUE))
```


# communicate

well not best week not best times ... just keep rowing.x2



