---
title: "Tidy tuesday challenge: Week 2019-10-08 PowerLifts!
author: "julio"
date: "2019-10-07"
output: html_document
---

# Tidy tuesday challenge: Week 2019-10-08 powerlifts !

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

just explore and plot something =) 

### Data:
 
 Week Week 2019-10-08 powerlifts!
 
### objectives:

explore & do something

## details:


## import data
```{r echo=FALSE,message=FALSE,warning=FALSE}
# library(magrittr)
library(tidyverse)
# library(ggrepel)
library(skimr)

library(here)
library(ggtext)#;devtools::install_github("clauswilke/ggtext")
# library(choroplethr)#;install.packages("choroplethr")
# library(choroplethrMaps)#;install.packages("choroplethrMaps") # ??choroplethrMaps

# library(dplyr)
# library(tidyr)
# library(readxl)
# library(janitor)
# library(stringr)
# library(lubridate)
# library(geojsonsf)

# library(geojsonio); #install.packages("geojsonio")

# library(grid)
# library(spatialEco)#; install.packages("spatialEco")
# library(sp)
# library(sf)# install.packages('sf')
# install.packages("Rcpp")
# remotes::install_github("tylermorganwall/rayshader")
# library(rayshader)

# library(transformr)# install.packages('transformr')
# library(gganimate)# install.packag
```


```{r echo=FALSE,message=FALSE,warning=FALSE}
ipf_lifts <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-10-08/ipf_lifts.csv")

```

too much data i'll just focus on parks

# explore data

```{r echo=FALSE,message=FALSE,warning=FALSE}
# now_data %>% purrr::map(now_data,head)
ipf_lifts %>% head()


```

```{r echo=FALSE,message=FALSE,warning=FALSE}

ipf_lifts %>% glimpse()



```

```{r echo=FALSE,message=FALSE,warning=FALSE}

ipf_lifts %>% skimr::skim()


```

```{r echo=FALSE,message=FALSE,warning=FALSE}
DataExplorer::plot_missing(ipf_lifts)

```

just simple sex, date, weight_class_kg, 

```{r echo=FALSE,message=FALSE,warning=FALSE}

ipf_lifts %>% 
    select(date,sex,weight_class_kg) %>% 
    ggplot(aes(x=date,y=weight_class_kg,color=sex))+
    geom_line()

```
ok not what i expected, so , ill re formule to , date, sex, amount of meets by decade


```{r echo=FALSE,message=FALSE,warning=FALSE}
ipf_processed <- ipf_lifts %>% 
    mutate(decade=as.integer(floor(lubridate::year(date)/10)*10)) %>% 
    mutate(meet_name=str_replace(meet_name,"Men's","")) %>% 
mutate(meet_name=str_replace(meet_name,"Women's","")) %>% 
mutate(meet_name=str_replace(meet_name,"  "," ")) %>% 
mutate(meet_name=str_trim(meet_name)) %>% 
mutate(meet_name=str_replace(meet_name,"6th ","")) %>%
mutate(meet_name=str_replace(meet_name,"3rd ","")) %>%
mutate(meet_name=str_replace(meet_name,"48th ","")) %>% 
mutate(meet_name=str_replace(meet_name,"Championships","Championship")) %>% 
    mutate(meet_name=str_replace(meet_name,"Student's","Students")) %>% 
    mutate(meet_name=str_replace(meet_name,"Students World Cup","World Students Cup"))
    
    

ipf_processed %>% 
    count(meet_name,sex,sort = TRUE) %>% 
    count(meet_name) %>% 
    arrange(meet_name)

#%>% 
    # filter(str_detect(meet_name,rebus::DGT))




```

male / female events proportion by decade

```{r echo=FALSE,message=FALSE,warning=FALSE}

total_participants_sex_by_decade <- ipf_processed %>% 
    count(decade,sex)

total_events_sex_by_decade <- ipf_processed %>% 
    distinct(date,decade,sex,meet_name) %>% 
    count(decade,sex,meet_name)

total_events_by_decade <-  ipf_processed %>% 
    distinct(date,decade,meet_name) %>% 
    count(decade,meet_name) %>% 
    rename(total_events=n)

total_events_proportions <- total_events_sex_by_decade %>% 
    left_join(total_events_by_decade) %>% 
    mutate(proportion=n/total_events) %>% 
    arrange(decade,meet_name)

top_event_by_decade <- ipf_processed %>% 
    nest(-decade,-sex) %>% 
    mutate(top_event=purrr::map(.x=data,.f=function(d){
        d %>% count(date,meet_name,sort=TRUE) %>% head(1)
    })) %>% 
    unnest(top_event) %>% 
    arrange(decade,sex) %>% 
    select(-data)

# top_event_by_decade$data[[1]] %>% count(date,meet_name,sort=TRUE) %>% head(1)



```
```{r echo=FALSE,message=FALSE,warning=FALSE}
# lubridate::now() - lubridate::days(365)
# annotations
plot_annotations <- top_event_by_decade %>% 
    mutate(label=paste0("**",str_replace_all(meet_name," ","<br>"),
                        "**<br>(",date,
                        ")<br> participants: ",n),
           # x=date-lubridate::days(365*5), #  1mont before will display the laabel
           x_minus= c(lubridate::days(365*5),#75, M
                      lubridate::days(365*5),#75,F
                      lubridate::days(365*5),#75,M
                      lubridate::days(365*5),#75,F
                      lubridate::days(365*5),#325,M
                      lubridate::days(365*5),#75,F
                      lubridate::days(365*2)*-1,#75,M
                      lubridate::days(365*5),#325,F
                      lubridate::days(365*8)),#325)M
           x=date-x_minus,
           # y=n # at the same Y
           # y=c(325,75,75,325,325,75,75,425,425)
           y=c(75,75,75,75,325,75,325,325,325)
           ) # text horizontal justifaction

# curves for arrows
plot_curves <- plot_annotations %>% 
    filter(sex=="M",decade%in%c(1990,2000)) %>% # only the ones that need arrows
      mutate(curve_delta=c(lubridate::days(365*2),
                           lubridate::days(365*3)),
             curve_x=x+curve_delta,
             curve_y=c(450,500),
             curve_x_end=date,curve_y_end=n,
             curvature=c(-0.3,-0.5))

             # (aes(x = date, y = y, 
             #         xend = x, yend = y),
             # arrow = arrow(length = unit(0.02, "npc")),
             # curvature = -0.3) +
```     


```{r echo=FALSE,message=FALSE,warning=FALSE}     

# amount of ppl attending to event.
plot_out <- ipf_processed %>% 
    count(date,sex) %>% 
    ggplot(aes(x=date,y=n,color=sex))+
    geom_line()+
    expand_limits(y = c(0,1200))+
    geom_point(data=top_event_by_decade,
               alpha=0.5,
               size=5)+
    geom_richtext(data = plot_annotations,
                 aes(label = str_wrap(label), x = x, y = y), 
                 hjust=0,
                 vjust = -0.2, size = 3,
                     # family = "IBM Plex Sans",
                 show.legend = FALSE,
                     fill = NA, label.color = NA) +
    geom_curve(data = plot_curves,
               aes(x = curve_x,
                   y = curve_y,
                   xend = curve_x_end,
                   yend = curve_y_end,
                   curvature=curvature),
             arrow = arrow(length = unit(0.02, "npc")))+
    facet_grid(sex~.)+
    theme_light()+
    theme(legend.position = "none")+
    labs(title="Participation of Powerlifts overtime",
         subtitle="highlighted top events of decades.",
         y="# amount of participants",
         x="", caption="#TidyTuesday")

# waffle::geom_waffle()

# ggsave("simple_choro.png",simple_choro,width = 7,height = 5)
# y_male=c(600,200,600,200,600)
# y_f=c(200,600,200,600)

# y_total <- c(600,200,200,600,600,200,200,600,600)

# c(600,200,200,600,600,200,200,600,600)-175
# print(c(600,200,200,600,600,200,200,600,600)-175)
# x_f=c(date-lubridate::days(365*5),
# date-lubridate::days(365*5),
# date-lubridate::days(365*5),
# date-lubridate::days(365*5),
# date-lubridate::days(365*15))
# x_total

# c(75,75,75,75,325,75,75,325,325)
# c(date-lubridate::days(365*15),#75, M
# date-lubridate::days(365*15),#75,F
# date-lubridate::days(365*15),#75,M
# date-lubridate::days(365*15),#75,F
# date-lubridate::days(365*15),#325,M
# date-lubridate::days(365*15),#75,F
# date-lubridate::days(365*5),#75,M
# date-lubridate::days(365*15),#325,F
# date-lubridate::days(365*20)),#325)M
plot_out
ggsave("powlift.png",plot_out,width = 7,height = 5)
```



#tweet: 

2019-10-08 #TidyTuesday #rstats Power Lift IPF !
Event participation over time, highlighted top events of each decade.
thanks @geokaramanis for sharing code ! helped me a lot to get started on annotations :D

code: https://github.com/jas1/tidytuesday/tree/master/jas1_weeks/2019/2019-10-08


# communicate

here tweet: https://twitter.com/jspairani/status/1175980540576620544

# resoutces: 

- https://github.com/gkaramanis/tidytuesday/blob/master/week-34/nuclear.R
- https://github.com/clauswilke/ggtext


