---
title: "Tidy tuesday challenge: Week 2019-04-16 better plots"
author: "julio"
date: "02-04-2019"
output: html_document
---

# Tidy tuesday challenge: Week 2019-04-16 better plots

keep it simple:

## Objectives: 

**general:**

* work on data, 
* practice, 
* get better on your workflow,
* get better on your skills: import, tidy , understand( transform, visualize,model ) , communicate


** this week **

### Data:

Better plots. 
https://medium.economist.com/mistakes-weve-drawn-a-few-8cdd8a42d368


### objectives:

- make some of the better plots, at least once that i like .

## details:


## import data
```{r echo=FALSE,message=FALSE,warning=FALSE}
library(dplyr)
library(tidyr)
library(skimr)
library(readr) 
library(ggplot2)
library(scales)
library(forcats)
library(lubridate)
library(gganimate)
library(animation)

```


```{r echo=FALSE,message=FALSE,warning=FALSE}
brexit <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-04-16/brexit.csv")
```

# explore data

```{r echo=FALSE,message=FALSE,warning=FALSE}
# 
color_red <- "#dd6e6d"
color_blue <- "#116ea1"

color_values_manual <- c("Wrong"=color_red,"Right"=color_blue)
colour_texts <- data.frame(x=ymd(c("2018-02-01","2018-02-01")),
                           y=c(43,46),
                           label=c("Right","Wrong"))

brexit_data <- brexit %>% 
    mutate(date=lubridate::dmy(date)) %>% 
    rename(Right=percent_responding_right,Wrong=percent_responding_wrong) %>% 
    gather(key="response_type",value = "percent",Right,Wrong)

bremorse_plot <- brexit_data  %>%  ggplot(aes(x=date,y=percent,colour=response_type))+
    geom_point(alpha=0.5)+
    geom_smooth(se = FALSE)+
    annotate("text", x = ymd("2018-02-01"), y = 47, label = "Wrong",
             colour=color_red,fontface =2)+
    annotate("text", x = ymd("2018-02-01"), y = 41, label = "Right",
             colour=color_blue,fontface =2)+
    # geom_text(data = colour_texts,aes(x=x,y=y,label=label))+
    scale_y_continuous(labels = scales::percent_format(scale=1),
                       position = "right")+
    # scale_x_continuous(breaks = breaks = c(2016,2017,2018))+
    scale_color_manual( values=color_values_manual)+
    theme_minimal()+
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90),
          plot.title=element_text(face="bold",colour = "#000000"),
          plot.subtitle=element_text(face="plain",colour = "#333333"),
          plot.caption =element_text(face="plain",colour = "#666666"))+
    labs(x="",y="",
         title="Bremorse",
         subtitle='"In hindsigth, do you think Britain was Rigth or Wrong\nto vote to leave EU"',
         caption = "Source NatCen Social Research")

ggsave(plot = bremorse_plot,filename = "20190418_bremorse.png")
        

# brexit_data %>% skimr::skim()

```

# usefull links: 

smooth line: 
https://ggplot2.tidyverse.org/reference/geom_smooth.html

text annotations: 
https://stackoverflow.com/questions/31568453/using-different-font-styles-in-annotate-ggplot2

percent: 
https://stackoverflow.com/questions/27433798/how-to-change-y-axis-range-to-percent-from-number-in-barplot-with-r

themes: 
https://ggplot2.tidyverse.org/reference/theme.html
